---
title: "RPA_Reweighter"
format: html
editor: source
project: execute-dir: project
---
##
## NOTES - need to run Scenario 1 and 4 with both the MAPC97 and MAPC101 config

# Need to run Scenario 3 for 2025 with the acs1721 headship rates instead of the scen2 rates
to do this need to make new target files with acs1721 rates on Scen3 population for 2025


## RPA Reweighter

This set of scripts runs the household reweighter to produce reweighted PUMS data that satisfies household targets set by the user.

# 1 - Set Scenario Name, knitr path, and root directory, load packages

```{r, setup, include=FALSE}

pacman::p_load(
  tidyverse,
  tidycensus,
  mapcdatakeys,
  janitor,
  readxl,
  rjson,
  stringr,
  data.table,
  here
)
# Set Scenario name

# scen <- 'Scenario_1'
# scen <- 'Scenario_2'
scen <- 'Scenario_3'
# scen <- 'Scenario_4'

# Set ACS / PUMS vintage
vintage <-  2021

#Headship rate Scenario
if (scen == 'Scenario_1'){
  hdshp <- 'acs1721'
}
if (scen != 'Scenario_1'){
  hdshp <- 'scen2'
}
# Set MAPC at 101 or 97 towns
# reb <- 'MAPC101'
reb <- 'MAPC97'

# List of RPA names
mpos <- unlist(c(mapcdatakeys::all_muni_data_keys |> select(mpo) |> unique()))
rpa.list <- mapcdatakeys::all_muni_data_keys |> select(rpa_acr) |> distinct(rpa_acr) |> pull()

# Paths
if (grepl('VM',Sys.info()[['nodename']])){
  base <- '//data-sync/Public/DataServices'
} else {
  base <- 'S:/Network Shares/K Drive/DataServices'
}

# Root Directory
root <- paste0(base,'/Projects/')

umdi.path <- paste0(root, 'Current_Projects/Housing/StatewideHousingPlan/04_Analysis/Data/Working/Regional_Control_Totals/UMDI/', scen, '/')

# if(scen %in% c('Scenario_2', 'Scenario_3', 'Scenario_4')){
# ctrltotal.path <- paste0(root, 'Current_Projects/Housing/StatewideHousingPlan/04_Analysis/Data/Working/Regional_Control_Totals/', scen, '/UPDATE/')
# } else { ctrltotal.path <- paste0(root, 'Current_Projects/Housing/StatewideHousingPlan/04_Analysis/Data/Working/Regional_Control_Totals/', scen, '/')
# }

# Set Shelter Households path
shelter_path <- paste0(
  root,
  'Current_Projects/Housing/StatewideHousingPlan/04_Analysis/Data/Final_Outputs/Shelter_Households/'
)

ctrltotal.path <- paste0(root, 'Current_Projects/Housing/StatewideHousingPlan/04_Analysis/Data/Working/Regional_Control_Totals/', scen, '/UPDATE/')

workpath <- paste0(root, 'Current_Projects/Housing/StatewideHousingPlan/04_Analysis/Data/Working/Reweighter/')

input_path <- paste0(workpath, scen, '/Input_Files_UPDATE/') 
output_path <- paste0(workpath, scen, '/Output_Files_UPDATE/') 

# Subfolders for RPA inputs and outputs
dir.create(input_path, showWarnings = F)
dir.create(output_path, showWarnings = F)

  for (m in mpos) {
    dir.create(paste0(input_path, m), showWarnings = F)
    dir.create(paste0(output_path, m), showWarnings = F)
  }

#Set knitr path
#knitr::opts_knit$set(root.dir = workpath)

## turn off scientific notation
options(scipen = 999)
#Set random seed
set.seed(351)
# Helper function
sun <- function(x){sort(unique(x))}
lun <- function(x){length(unique(x))}

```

# 2 - Subset PUMS data for each RPA

Formatted PUMS data is created by the *01_create_regional_control_totals.qmd* in section 6.2

# 2.1 - Scenario 1

```{r}
if(scen=='Scenario_1') {
  pums_data <- fread(paste0(workpath, 'PUMS_data/PUMS', vintage, '_formatted.csv')) |> 
    filter(TYPEHUGQ == "1") |>
    mutate(AgeCat = ageCAT6)
  
  mapc.pumas <- pums_data[rpa_acr == 'MAPC', unique(PUMA)]
  swm.pumas <- pums_data[rpa_acr != 'MAPC', .(rpa_acr, PUMA)] |>
    unique()
  # MAPC 101 subset
  fwrite(
    pums_data[PUMA %in% mapc.pumas],
    paste0(
      workpath,
      scen,
      '/Input_Files_UPDATE/MAPC/PUMS2021_',
      scen,
      '_MAPC.csv'
    )
  )
  message(paste0('MAPC 101 PUMS export complete '))
  # SWM join PUMA and export subsets
  for (m in unique(swm.pumas$rpa_acr)) {
    pums_data |>
      filter(rpa_acr == m) |>
      fwrite(
        paste0(
          workpath,
          scen,
          '/Input_Files_UPDATE/',
          m,
          '/PUMS2021_',
          scen,
          '_',
          m,
          '.csv'
        )
      )
    message(paste0(m, ' PUMS export complete '))
  }
}
```

# 2.2 - Scenario 2/3/4

```{r}
if(scen %in% c('Scenario_2', 'Scenario_3', 'Scenario_4')) {
  
  # We keep the base ACS 17-21 headship rates and PUMS subset from Scenario 1 for the year 2025, but applied to the Scenario 2/3/4 populations from UMDI. Adjusted headship rates and PUMS subsets are used only for year 2030 and after
  for (yr in 2025){
    pums_data <- fread(paste0(workpath, 'PUMS_data/PUMS', vintage, '_formatted.csv')) |> filter(TYPEHUGQ == "1") |>
    mutate(AgeCat = ageCAT6)
  for (m in mpos) {
    dir.create(paste0(input_path, m), showWarnings = F)
    dir.create(paste0(output_path, m), showWarnings = F)
  }
  mapc.pumas <- pums_data[rpa_acr == 'MAPC', unique(PUMA)]
  swm.pumas <- pums_data[rpa_acr != 'MAPC', .(rpa_acr, PUMA)] |>
    unique()
  # MAPC 101 subset
  fwrite(
    pums_data[PUMA %in% mapc.pumas],
    paste0(
      input_path,'MAPC/PUMS2021_',
      scen,
      '_MAPC_',
      yr,
      '.csv'
    )
  )
  message(paste0('MAPC 101 PUMS export complete ', yr))
  # SWM join PUMA and export subsets
  for (m in unique(swm.pumas$rpa_acr)) {
    pums_data |>
      filter(rpa_acr == m) |>
      fwrite(
        paste0(
          input_path,
          m,
          '/PUMS2021_',
          scen,
          '_',
          m,
          '_',
          yr,
          '.csv'
        )
      )
    message(paste0(m, ' PUMS export complete ', yr))
  }
  }
  
  for (yr in seq(2030, 2050, 5)) {
    x2 <- fread(paste0(workpath, 'PUMS_data/PUMA10_RPA_crosswalk.csv')) |> 
      mutate(PUMA = as.numeric(PUMA))
    pums_data <- fread(paste0(workpath, 'PUMS_data/PUMS2021_Scenario_2_', yr, '.csv')) |>
      left_join(x2, by = c('PUMA')) |>
      filter(TYPEHUGQ == "1") |>
      mutate(AgeCat = ageCAT6)
    pums_data[!is.na(SPORDER.adj), SPORDER := SPORDER.adj]
    pums_data[!is.na(HHtype.adj), HHtype := HHtype.adj]
    pums_data[!is.na(SERIALNO.adj), SERIALNO := SERIALNO.adj]

    mapc.pumas <- pums_data[rpa_acr == 'MAPC', unique(PUMA)]
    swm.pumas <- pums_data[rpa_acr != 'MAPC', .(rpa_acr, PUMA)] |>
      unique()
    
    # MAPC 101 subset
    fwrite(
      pums_data[PUMA %in% mapc.pumas],
      paste0(
        input_path,'MAPC/PUMS2021_',
        scen,
        '_MAPC_',
        yr,
        '.csv'
      )
    )
    message(paste0('MAPC 101 PUMS export complete ', yr))
    
    # SWM join PUMA and export subsets
    for (m in unique(swm.pumas$rpa_acr)) {
      pums_data |>
        filter(rpa_acr == m) |>
        fwrite(
          paste0(
            input_path,
            m,
            '/PUMS2021_',
            scen,
            '_',
            m,
            '_',
            yr,
            '.csv'
          )
        )
      message(paste0(m, ' PUMS export complete ', yr))
    }
  }
}
```

# 3 - Create RPA config JSON files for reweighter configuration

```{r}

# Template config file
infile1 <- paste0(workpath, 'Templates/reweighting_config_2021.json')

infile2 <- paste0(workpath, 'Templates/reweighting_config_2020baseyear.json')

for (m in mpos){
  outfile1 <- paste0(input_path, m, '/reweighting_config_2021_', m, '.json')
  outfile2 <- paste0(input_path, m, '/reweighting_config_2020baseyear_', m, '.json')
  cfig1 = readLines(infile1)
  cfig1 = gsub('PUMS2021', paste0('PUMS2021_', m), cfig1)
  writeLines(cfig1, outfile1)
  cfig2 = readLines(infile2)
  cfig2 = gsub('PUMS2021', paste0('PUMS2021_', m), cfig2)
  writeLines(cfig2, outfile2)
}

```

# 4 - Configure, set targets and run reweighter for each RPA and year

# 4.1 - Scenario 1

```{r}

#
source(here('scripts/reweighter/algo_hh.R'))

# Set Targets and run reweighter loop

# Different configuration for 2020 base year
if (scen == 'Scenario_1') {
  for (yr in c(2020)) {
    #for (mid in c('MAPC')) {
    for (mid in mpos) {
      # Set config files
      inpath <- paste0(input_path, mid, '/')
      
      source(here('scripts/reweighter/read_config_RPA_2020baseyear.R'))
      message(paste0(mid, ' configuration complete'))
      
      outpath <- paste0(output_path, mid, '/')
      age.target <- fread(
        paste0(
          ctrltotal.path,
          'age.targetfile.',
          reb,
          '.',
          scen,
          '.hdshp.',
          hdshp,
          '.y',
          yr,
          '.csv'
        )
      ) |>
        filter(RPA == mid) |>
        select(-RPA) |> setDT()
      setorder(age.target, ageCAT4)
      fwrite(age.target, paste0(inpath, '/age.csv'))
      
      size.target <- fread(
        paste0(
          ctrltotal.path,
          'size.targetfile.',
          reb,
          '.',
          scen,
          '.hdshp.',
          hdshp,
          '.y',
          yr,
          '.csv'
        )
      ) |>
        filter(RPA == mid) |>
        select(-RPA) |> setDT()
      setorder(size.target, size)
      fwrite(size.target, paste0(inpath, '/size.csv'))
      
      type.target <- fread(
        paste0(
          ctrltotal.path,
          'type.targetfile.',
          reb,
          '.',
          scen,
          '.hdshp.',
          hdshp,
          '.y',
          yr,
          '.csv'
        )
      ) |>
        filter(RPA == mid) |>
        select(-RPA) |> setDT()
      setorder(type.target, HHtype)
      fwrite(type.target, paste0(inpath, '/type.csv'))
      
      message(paste0(mid, ' targets set'))
      
      pums <- fread(paste0(inpath, 'PUMS', vintage, '_', mid, '.csv'))
      
      source(here('scripts/reweighter/Manage_Scenario_RPA.R'))
      message(paste0(mid, ' reweighter complete'))
      
      mess <- data.table(txt = paste0(mid, ' reweighter complete'))
      fwrite(mess, paste0(outpath, mid, '_', yr, '_complete.txt'))
    }
  }
  # for (yr in c(2025, 2030, 2035, 2040, 2045, 2050)) {
  #   for (mid in mpos) {
  #     # Set config
  #     inpath <- paste0(input_path, mid, '/')
  #     source(here('scripts/reweighter/read_config_RPA.R'))
  #     message(paste0(mid, ' configuration complete'))
  #     
  #     # Set targets
  #     outpath <- paste0(output_path, mid, '/')
  #     dir.create(outpath, showWarnings = F)
  #     template <- fread(paste0(workpath, 'Templates/AgeHHder_HHtype_template.csv')) |>
  #       select(-BASELINE, -TARGET)
  #     target <- fread(
  #       paste0(
  #         ctrltotal.path,
  #         'hhtype.targetfile.',
  #         reb,
  #         '.',
  #         scen,
  #         '.hdshp.',
  #         hdshp,
  #         '.y',
  #         yr,
  #         '.csv'
  #       )
  #     ) |>
  #       filter(RPA == mid) |>
  #       select(-RPA) |> right_join(template, by = c('AgeCat', 'HHtype')) |> setDT()
  #     target[AgeCat %in% c('x == 1', 'x == 2', 'x == 3'), `:=`(BASELINE =
  #                                                                0,
  #                                                              INTER = 0,
  #                                                              TARGET = 0)]
  #     target[is.na(BASELINE), `:=`(BASELINE = 0,
  #                                  INTER = 0,
  #                                  TARGET = 0)]
  #     setorder(target, AgeCat, HHtype)
  #     fwrite(target, paste0(inpath, '/AgeHHder_HHtype.csv'))
  #     
  #     template2 <- fread(paste0(workpath, 'Templates/age_tenure_template.csv'))
  #     target2 <- fread(
  #       paste0(
  #         ctrltotal.path,
  #         "tenure.targetfile.",
  #         reb,
  #         ".",
  #         scen,
  #         '.hdshp.',
  #         hdshp,
  #         ".y",
  #         yr,
  #         ".csv"
  #       )
  #     ) |>
  #       filter(RPA == mid) |>
  #       select(-RPA) |>
  #       right_join(template2, by = c('AgeCat', 'tenure')) |>
  #       setDT()
  #     
  #     target2[is.na(BASELINE), `:=`(BASELINE = 0,
  #                                   INTER = 0,
  #                                   TARGET = 0)]
  #     setorder(target2, AgeCat, tenure)
  #     fwrite(target2, paste0(inpath, '/age_tenure.csv'))
  #     
  #     message(paste0(mid, ' targets set'))
  #     
  #     pums <- fread(paste0(inpath, 'PUMS', vintage, '_', mid, '.csv'))
  #     if (scen == 'Scenario_2') {
  #       pums[, SPORDER := SPORDER.adj][, SERIALNO := SERIALNO.adj]
  #     }
  #     
  #     source(here('scripts/reweighter/Manage_Scenario_RPA.R'))
  #     message(paste0(mid, ' reweighter complete'))
  #     
  #     mess <- data.table(txt = paste0(mid, ' reweighter complete'))
  #     fwrite(mess, paste0(outpath, mid, '_', yr, '_complete.txt'))
  #   }
  # }
}

```

# 4.2 - Scenario 2/3/4

```{r}


source(here('scripts/reweighter/algo_hh.R'))

if (scen %in% c('Scenario_2', 'Scenario_3', 'Scenario_4')) {
  for (yr in c(2030, 2035, 2040, 2045, 2050)) {
  # for (yr in c(2025)) {
    for (mid in mpos) {
      infile1 <- paste0(workpath, 'Templates/reweighting_config_2021.json')
      outfile1 <- paste0(
        input_path,
        mid,
        '/reweighting_config_2021_',
        mid,
        '.json'
      )
      cfig1 = readLines(infile1)
      cfig1 = gsub('PUMS2021.csv',
                   paste0('PUMS2021_', scen, '_', mid, '_', yr, '.csv'),
                   cfig1)
      writeLines(cfig1, outfile1)
      # Set config
      inpath <- paste0(input_path, mid, '/')
      source(here('scripts/reweighter/read_config_RPA.R'))
      message(paste0(mid, ' configuration complete'))
      
      # Set targets
      outpath <- paste0(output_path, mid, '/')
      dir.create(outpath, showWarnings = F)
      template <- fread(paste0(workpath, 'Templates/AgeHHder_HHtype_template.csv')) |>
        select(-BASELINE, -TARGET)
      target <- fread(
        paste0(
          ctrltotal.path,
          'hhtype.targetfile.',
          reb,
          '.',
          scen,
          '.hdshp.',
          hdshp,
          '.y',
          yr,
          '.csv'
        )
      ) |>
        filter(RPA == mid) |>
        select(-RPA) |> right_join(template, by = c('AgeCat', 'HHtype')) |> setDT()
      target[AgeCat %in% c('x == 1', 'x == 2', 'x == 3'), `:=`(BASELINE =
                                                                 0,
                                                               INTER = 0,
                                                               TARGET = 0)]
      target[is.na(BASELINE), `:=`(BASELINE = 0,
                                   INTER = 0,
                                   TARGET = 0)]
      setorder(target, AgeCat, HHtype)
      fwrite(target, paste0(inpath, '/AgeHHder_HHtype.csv'))
      
      template2 <- fread(paste0(workpath, 'Templates/age_tenure_template.csv'))
      target2 <- fread(
        paste0(
          ctrltotal.path,
          "tenure.targetfile.",
          reb,
          ".",
          scen,
          '.hdshp.',
          hdshp,
          ".y",
          yr,
          ".csv"
        )
      ) |>
        filter(RPA == mid) |>
        select(-RPA) |>
        right_join(template2, by = c('AgeCat', 'tenure')) |>
        setDT()
      
      target2[is.na(BASELINE), `:=`(BASELINE = 0,
                                    INTER = 0,
                                    TARGET = 0)]
      setorder(target2, AgeCat, tenure)
      fwrite(target2, paste0(inpath, '/age_tenure.csv'))
      
      message(paste0(mid, ' targets set'))
      
      pums <- fread(paste0(inpath, 'PUMS2021_', scen, '_', mid, '_', yr, '.csv'))

      source(here('scripts/reweighter/Manage_Scenario_RPA.R'))
      message(paste0(mid, ' reweighter complete'))
      
      mess <- data.table(txt = paste0(mid, ' reweighter complete'))
      fwrite(mess, paste0(outpath, mid, '_', yr, '_complete.txt'))
    }
  }
}

```

# 5 - Create reweighted PUMS files and QC the reweighter Outputs

# 5.0 - Check that all years/RPAs successfully converged

```{r}
  
log.comp <- c()
for (mid in mpos) {
  outpath <- paste0(output_path, mid, '/')
  logs <- list.files(path = outpath, pattern =paste0('logfile_',scen,"_",reb))
  for (f in logs){
    df <- readLines(paste0(outpath,f))
    log.comp <- c(log.comp, if (length(
      grep("Maximum deviation from target stopping condition achieved", df)
    ) != 0) {
      print(paste0(f, ' success'))
    } else if (length(grep("Maximum number of iterations reached", df)) !=
               0) {
      print(paste0(f, ' failed to converge'))
    } else if (length(grep(
      "No change in weights stopping condition met", df
    )) != 0) {
      print(paste0(f, ' no change in weights'))
    })
  }
}
fwrite(as.data.table(log.comp),paste0(output_path,'/reweighter_convergence_check_',reb,'.txt'))




```

# 5.1 - Scenario 1

```{r}

comp.dt <- data.table()
if(scen=='Scenario_1'){
for (yr in 2020){
  for (mid in mpos) {
    inpath <- paste0(workpath, scen, '/Input_Files_UPDATE/', mid, '/')
    outpath <- paste0(workpath, scen, '/Output_Files_UPDATE/', mid, '/')
    # Summarize total households by mpo from target inputs to reweighter
    target <- fread(paste0(ctrltotal.path, 'type.targetfile.', reb, '.', scen, '.hdshp.', hdshp, '.y', yr, '.csv')) |>
      filter(RPA == mid) |> setDT()
    
    tsum <- target[, lapply(.SD, sum), RPA, .SDcols = 'TARGET']
    tsum[, TARGET := round(TARGET)][,year:=yr]
    setnames(tsum, 'RPA', 'mpo')
    setkey(tsum, mpo)
    
    # Calculate total households from the reweighter outputs
    pums <- fread(paste0(inpath, 'PUMS', vintage, '_', mid, '.csv'))
    setkey(pums, SERIALNO, SPORDER)

    x <- fread(paste0(outpath, 'weights_final_', scen, '_', reb, '_', mid, '_', yr, '.csv')) |> select(SERIALNO, SPORDER, new_WGTP) |> mutate(mpo = mid) |> setDT()
    setkey(x, SERIALNO, SPORDER)
    xp <- pums[x]
    fwrite(xp, paste0(outpath, '/reweighted_PUMS_', scen, '_', reb, '_', mid, '_', hdshp, '_', yr, '.csv'))
    rw <- xp[SPORDER == 1, lapply(.SD, sum), .(mpo), .SDcols = c('WGTP', 'new_WGTP')]
    rw[, rw_HH := round(new_WGTP)][, pums_baseline_hh := round(WGTP)]
    setkey(rw, mpo)
    comp <- tsum[rw]
    comp.out <- comp[,.(year,mpo,pums_baseline_hh, TARGET,rw_HH)]
    comp.out[,diff:= TARGET - rw_HH]
    fwrite(comp.out,paste0(outpath,'reweighter_QC_', scen, '_', reb, '_', mid, '_', hdshp, '_', yr,' .csv'))
    comp.dt <- rbind(comp.dt,comp.out)
  }
}
# for (yr in c(2025,2030,2035,2040,2045,2050)){
#   for (mid in mpos) {
#     inpath <- paste0(workpath, scen, '/Input_Files/', mid, '/')
#     outpath <- paste0(workpath, scen, '/Output_Files/', mid, '/')
#     # Summarize total households by mpo from target inputs to reweighter
#     target <- fread(paste0(ctrltotal.path, 'hhtype.targetfile.', reb, '.', scen, '.hdshp.', hdshp, '.y', yr, '.csv')) |>
#       filter(RPA == mid) |> setDT()
#     
#     tsum <- target[, lapply(.SD, sum), RPA, .SDcols = 'TARGET']
#     tsum[, TARGET := round(TARGET)][,year:=yr]
#     setnames(tsum, 'RPA', 'mpo')
#     setkey(tsum, mpo)
#     
#     # Calculate total households from the reweighter outputs
#     pums <- fread(paste0(inpath, 'PUMS', vintage, '_', mid, '.csv'))
#     setkey(pums, SERIALNO, SPORDER)
# 
#     x <- fread(paste0(outpath, 'weights_final_', scen, '_', reb,  '_', mid, '_', yr, '.csv')) |> select(SERIALNO, SPORDER, new_WGTP) |> mutate(mpo = mid) |> setDT()
#     setkey(x, SERIALNO, SPORDER)
#     xp <- pums[x]
#     fwrite(xp, paste0(outpath, '/reweighted_PUMS_', scen, '_', reb, '_', mid, '_', hdshp, '_', yr, '.csv'))
#     rw <- xp[SPORDER == 1, lapply(.SD, sum), .(mpo), .SDcols = c('WGTP', 'new_WGTP')]
#     rw[, rw_HH := round(new_WGTP)][, pums_baseline_hh := round(WGTP)]
#     setkey(rw, mpo)
#     comp <- tsum[rw]
#     comp.out <- comp[,.(year,mpo,pums_baseline_hh, TARGET,rw_HH)]
#     comp.out[,diff:= TARGET - rw_HH]
#     fwrite(comp.out,paste0(outpath,'reweighter_QC_', scen, '_', reb, '_', mid, '_', hdshp, '_', yr,' .csv'))
#     comp.dt <- rbind(comp.dt,comp.out)
# 
#   }
# }

fwrite(comp.dt,paste0(workpath, scen, '/Output_Files/reweighter_QC_all_RPAs_all_years_', scen,  '_', reb, '_', hdshp,' .csv'))

comp.dt[abs(diff)>1]
}
```

# 5.2 - Scenario 2/3/4

```{r}
c0 <- data.table()
comp.dt <- data.table()
if(scen %in% c('Scenario_2','Scenario_3','Scenario_4')){
for (yr in c(2020)){
  for (mid in mpos) {
    inpath <- paste0(input_path, mid, '/')
    outpath <- paste0(output_path, mid, '/')
    
    # Use Scenario 1 outputs for 2020 
    file.copy(paste0(workpath, 'Scenario_1/Output_Files_UPDATE/', mid, '/reweighted_PUMS_Scenario_1_', reb, '_', mid, '_acs1721_', yr, '.csv'),paste0(outpath, 'reweighted_PUMS_', scen, '_', reb, '_', mid, '_', hdshp, '_', yr, '.csv'), overwrite = T)
  }
}

if(reb=='MAPC101') {
  for (yr in c(2025)) {
    for (mid in mpos) {
      inpath <- paste0(input_path, mid, '/')
      outpath <- paste0(output_path, mid, '/')

      # Use unaffected outputs for 2025
      file.copy(
        paste0(
          workpath,
          scen,
          '/Output_Files/',
          mid,
          '/reweighted_PUMS_',
          scen,
          '_',
          reb,
          '_',
          mid,
          '_scen2_',
          yr,
          '.csv'
        ),
        paste0(
          outpath,
          '/reweighted_PUMS_',
          scen,
          '_',
          reb,
          '_',
          mid,
          '_',
          hdshp,
          '_',
          yr,
          '.csv'
        ),
        overwrite = T
      )
    }
  }
}

for (yr in c(2030,2035,2040,2045,2050)){
# for (yr in c(2025)){
  for (mid in mpos) {
    inpath <- paste0(input_path, mid, '/')
    outpath <- paste0(output_path, mid, '/')
    
    
    # Calculate total households from the reweighter outputs
  
      pums <- fread(paste0(inpath, 'PUMS', vintage, '_', scen, '_', mid, '_', yr, '.csv'))
    setkey(pums, SERIALNO, SPORDER)

    x <- fread(paste0(outpath, 'weights_final_', scen,  '_', reb, '_', mid, '_', yr, '.csv')) |> select(SERIALNO, SPORDER, new_WGTP) |> mutate(mpo = mid) |> setDT()
    setkey(x, SERIALNO, SPORDER)
    xp <- pums[x]
    fwrite(xp, paste0(outpath, 'reweighted_PUMS_', scen, '_', reb, '_', mid, '_', hdshp, '_', yr, '.csv'))
    rw <- xp[SPORDER == 1, lapply(.SD, sum), .(mpo), .SDcols = c('WGTP', 'new_WGTP')]
    rw[, rw_HH := round(new_WGTP)][, pums_baseline_hh := round(WGTP)]
    setkey(rw, mpo)
    
    r1 <- xp[SPORDER == 1, lapply(.SD, sum), .(mpo,AgeCat,HHtype), .SDcols = c('WGTP', 'new_WGTP')]
    r1[, new_WGTP := round(new_WGTP)][, WGTP := round(WGTP)]
    r1[,AgeCat:=paste0('x == ',AgeCat)]
    r1[,HHtype:=paste0('x == ',HHtype)]
    setkey(r1, mpo,AgeCat,HHtype)
    # Summarize total households by mpo from target inputs to reweighter
    target <- fread(paste0(ctrltotal.path, 'hhtype.targetfile.', reb, '.', scen, '.hdshp.', hdshp, '.y', yr, '.csv')) |>
      filter(RPA == mid) |> rename(mpo=RPA) |> setDT()
    setkey(target,mpo,AgeCat,HHtype)
    
    cc <- r1[target]
    cc[,diff:=new_WGTP-TARGET]
    cc[,year:=yr]
    
    
    tsum <- target[, lapply(.SD, sum), mpo, .SDcols = 'TARGET']
    tsum[, TARGET := round(TARGET)][,year:=yr]
    #setnames(tsum, 'RPA', 'mpo')
    setkey(tsum, mpo)
    
    comp <- tsum[rw]
    comp.out <- comp[,.(year,mpo,pums_baseline_hh, TARGET,rw_HH)]
    comp.out[,diff:= TARGET - rw_HH]
    fwrite(comp.out,paste0(outpath,'reweighter_QC_', scen, '_', reb, '_', mid, '_', hdshp, '_', yr,'.csv'))
    comp.dt <- rbind(comp.dt,comp.out)
    c0 <- rbind(c0,cc)
  }
}

fwrite(comp.dt,paste0(output_path,'reweighter_QC_all_RPAs_all_years_', scen, '_', reb,  '_', hdshp,'.csv'))
fwrite(c0,paste0(output_path,'reweighter_QC_all_RPAs_by_AgeCat_HHtype_all_years_', scen,  '_', reb, '_', hdshp,'.csv'))

comp.dt[abs(diff)>1]

}
```

# 6 - Create summary tables of reweighted households

```{r}

for (yr in c(2020,2025,2030,2035,2040,2045,2050)){
# for (yr in c(2020)){
  sw <- data.table()
  for (mid in mpos){
    inp <- paste0(output_path, mid, '/')
    dt <- fread(paste0(inp, 'reweighted_PUMS_', scen, '_', reb, '_', mid, '_', hdshp, '_', yr, '.csv'))
    dt[, year:=yr][,reb:=reb][,hdshp:=hdshp][,scen:=scen]
    sw <- rbind(sw,dt)
  }
 fwrite(sw,paste0(output_path,'/reweighted_PUMS_', scen, '_', reb, '_', hdshp, '_', yr, '.csv'))
 }
```

# 7 - OPTIONAL - Population Post-Processor

# 7.1 - Population difference calculation

```{r}
#| eval: false


# Set Year(s)
#
# Post-processing for year 2020 was done for Scenario 1 (both MAPC97 and MAPC101) and will be copied over from the Scenario 1 folder to the current scenario folder and renamed at the end of the 7.3 post-processor code block.

for (yr in c(2025,2030,2035,2040,2045,2050)){

# Load UMDI HHpop comparison data

umdi <- fread(paste0(ctrltotal.path,'pop.in.hhds.',reb,'_by.RPA.ageCAT6.csv'))

setnames(
  umdi,
  c(
    'mpo',
    'ageCAT6',
    'year',
    'pop',
    'umdi_hh_pop'
  )
)

umdi <-
  umdi[, lapply(.SD, sum, na.rm = T), .(year, mpo, ageCAT6), .SDcols = c('umdi_hh_pop')]
umdi[, umdi_hh_pop := round(umdi_hh_pop)]
umdi <- umdi[year >= 2020]
setkey(umdi, mpo, year, ageCAT6)

##  CALCULATE differences in HHpop by age between UMDI/Control Totals and reweighter Outputs

    dx <- fread(paste0(output_path,'/reweighted_PUMS_', scen, '_', reb, '_', hdshp, '_', yr, '.csv'))
    pums <- fread(paste0(workpath,'PUMS_data/PUMS2021_PINCP.csv'))
    if(!'PINCP' %in% names(dx)){
      dx <- dx |> left_join(pums, by=c('SERIALNO','SPORDER'))
    }
    dx[PINCP<0 | is.na(PINCP),PINCP:=0]
    dx[, c("WGT1", "WGT2") := tstrsplit(new_WGTP, ".", fixed=TRUE)]
    dx[,WGT1:=as.numeric(WGT1)]
    dx[is.na(WGT2),WGT2:=0]
    dx[,WGT2:=as.numeric(paste0('0.',WGT2))]
    
    dfrac <- dx[WGT2>0]
    dfrac[,new_WGTP:=WGT2]
    dfrac[,hid:=paste0(SERIALNO,'_',mpo,'_frac')]
    
    xr <- dx[rep(1:.N,WGT1)]
    xr[,new_WGTP:=1]
    setorder(xr,SERIALNO,SPORDER)
    xr[,ss:=seq_along(.I),.(SERIALNO,SPORDER)]
    xr[, hid := paste(SERIALNO, mpo, ss, sep = '_')]
    xr[,ss:=NULL]
    dt <- rbind(xr,dfrac)
    dt[,pop:=new_WGTP]
    dt[,pid:=paste0(hid,'_',SPORDER)]
    dt[,person_num:=SPORDER][,age:=AGEP]
    dt[,HINCP:=sum(PINCP),hid]
    saveRDS(dt,paste0(output_path,'expanded_reweighted_PUMS_',reb,'_',yr,'.rds'))
    uspop <- dt[,lapply(.SD,sum),.(mpo,year,ageCAT6),.SDcols='pop']
    uspop[,pop:=round(pop)]
    setkey(uspop, mpo, year, ageCAT6)
    adiff <- umdi[uspop]
    adiff[, diff := pop - umdi_hh_pop]
    setnames(adiff, 'pop', 'urbansim_hh_pop')
    fwrite(adiff, paste0(output_path,'umdi_pop_age_difference_',reb,'_',yr,'.csv'))
rm(xr,dt,dx,dfrac);gc()

message(paste(yr,'Complete'))

}


```

# 7.2 - Person duplication function

```{r}
#| eval: false

# Function that creates subsets of households to sample from for replicate/delete adjustments
cx <- function(x) {
  sub1 <- x[ageCAT6 == 1  & children > 1 & ageHHder>17]
  add1 <- x[ageCAT6 == 1  & children < 4 &
              adults > 1 & ageHHder %in% 20:60]
  sub2 <- x[ageCAT6 == 2  & children > 1 & ageHHder>17]
  add2 <- x[ageCAT6 == 2  & children < 4 &
              adults > 1 & ageHHder %in% 20:60]
  sub3 <- x[ageCAT6 == 3  & children > 1 & ageHHder>17]
  add3 <- x[ageCAT6 == 3  & children < 4 &
              adults > 1 & ageHHder %in% 20:60]
  sub4 <-
    x[(ageCAT6 == 4  &
         HHT == 3 &
         persons > 2 & person_num > 1) | (ageCAT6 == 4 &
                            HHT == 1 & children > 1 & person_num > 1 & ageHHder>17)]
  add4 <-
    x[(ageCAT6 == 4  &
         HHT == 1 &
         children < 4 &
         adults > 1 &
         ageHHder %in% 25:50 &
         person_num > 1) |
        (ageCAT6 == 4  & ageHHder < 44 & HHT == 3 & persons < 6 &
           person_num > 1)]
  
  sub5 <-
    x[(ageCAT6 == 5  &
         HHT == 3 &
         persons > 2 &
         person_num > 1) | (ageCAT6 == 5 &
                              HHT == 1 & adults >= 2 &
                              person_num > 1)]
  add5 <- x[ageCAT6 == 5  & HHT == 3 &
              persons >= 2  & persons<8 & ageHHder %in% 20:44]
  
  sub6 <-
    x[(ageCAT6 == 6  &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 6  & HHT == 1 & adults >= 2 &
           person_num > 1)]
  add6 <- x[ageCAT6 == 6  & HHT == 3 &
              persons >= 2  & persons<8 & ageHHder %in% 20:44]
  
  sub7 <-
    x[(ageCAT6 == 7  &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 7  & HHT == 1 & adults >= 2 &
           person_num > 1)]
  add7 <- x[ageCAT6 == 7  & HHT == 3 &
              persons >= 2  & persons<8 & ageHHder %in% 20:44]
  
  sub8 <-
    x[(ageCAT6 == 8  &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 8  & HHT == 1 & adults >= 2 &
           person_num > 1)]
  add8 <- x[ageCAT6 == 8  & HHT == 3 &
              persons >= 2  & persons<8 & ageHHder %in% 24:49]
  
  sub9 <-
    x[(ageCAT6 == 9  &
         HHT == 3 &
         persons > 2 &
         person_num > 1) | (ageCAT6 == 9 &
                              HHT == 1 & adults >= 2 &
                              person_num > 1)]
  add9 <- x[ageCAT6 == 9  & HHT == 3 &
              persons >= 2  & persons<8 & ageHHder %in% 30:54]
  
  sub10 <-
    x[(ageCAT6 == 10 &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 10 & HHT == 1 & adults >= 2 &
           person_num > 1)]
  add10 <- x[ageCAT6 == 10 &
               HHT == 3 & persons >= 2  & persons<8 & ageHHder %in% 40:59]
  
  sub11 <-
    x[(ageCAT6 == 11 &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 11 &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add11 <- x[ageCAT6 == 11 &
               HHT == 3 & persons >= 2  & persons<8 & ageHHder %in% 40:64]
  sub12 <-
    x[(ageCAT6 == 12 &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 12 &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add12 <- x[ageCAT6 == 12  &
               HHT == 3 & persons >= 2  & persons<8 & ageHHder %in% 45:69]
  sub13 <-
    x[(ageCAT6 == 13 &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 13  &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add13 <- x[ageCAT6 == 13  &
               HHT == 3 & persons >= 2 & ageHHder %in% 50:74]
  
  sub14 <-
    x[(ageCAT6 == 14  &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 14  &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add14 <- x[ageCAT6 == 14 &
               HHT > 1 & persons >= 2 & persons<8 ]
  sub15 <-
    x[(ageCAT6 == 15 &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 15  &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add15 <- x[ageCAT6 == 15 &
               HHT > 1 & persons >= 2 & persons<8 ]
  sub16 <-
    x[(ageCAT6 == 16 &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 16 &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add16 <- x[ageCAT6 == 16 &
               HHT > 1 & persons >= 2 & persons<8 ]
  sub17 <-
    x[(ageCAT6 == 17 &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 17 &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add17 <- x[ageCAT6 == 17 &
               HHT > 1 & persons >= 2 & persons<8 ]
  sub18 <-
    x[(ageCAT6 == 18 &
         HHT == 3 &
         persons > 2 &
         person_num > 1) |
        (ageCAT6 == 18 &
           HHT == 1 & adults >= 2 & person_num > 1)]
  add18 <- x[ageCAT6 == 18 &
               HHT > 1 & persons >= 2 & persons<8 ]
  
  add.sample <<-
    list(
      add1,
      add2,
      add3,
      add4,
      add5,
      add6,
      add7,
      add8,
      add9,
      add10,
      add11,
      add12,
      add13,
      add14,
      add15,
      add16,
      add17,
      add18
    )
  sub.sample <<-
    list(
      sub1,
      sub2,
      sub3,
      sub4,
      sub5,
      sub6,
      sub7,
      sub8,
      sub9,
      sub10,
      sub11,
      sub12,
      sub13,
      sub14,
      sub15,
      sub16,
      sub17,
      sub18
    )
  
}
```

# 7.3 - Post-processor 04 - run person duplication/subtraction

```{r}

# for (yr in c(2020)){
for (yr in c(2025,2030,2035,2040,2045,2050)){

flaglist <- data.table()

adiff <- fread(paste0(output_path,'umdi_pop_age_difference_',reb,'_',yr,'.csv'))
adiff[,diff:=round(diff)]


dt <- readRDS(paste0(output_path,'expanded_reweighted_PUMS_',reb,'_',yr,'.rds')) |> 
                mutate(
      child = case_when(AGEP < 18 ~ 1,
                        AGEP >= 18 ~ 0)) |>  
      setDT()
  dt[, children := sum(child), hid]
  dt[child == 0, adult := 1]
  dt[, adults := sum(adult,na.rm=T), hid]
  dt[ageCAT6 == 1, age1 := .N, hid]
  dt[ageCAT6 == 2, age2 := .N, hid]
  dt[ageCAT6 == 3, age3 := .N, hid]
  dt[ageCAT6 == 4, age4 := .N, hid]
  dt[ageCAT6 == 5, age5 := .N, hid]
  dt[ageCAT6 == 6, age6 := .N, hid]
  dt[ageCAT6 == 7, age7 := .N, hid]
  dt[ageCAT6 == 8, age8 := .N, hid]
  dt[ageCAT6 == 9, age9 := .N, hid]
  dt[ageCAT6 == 10, age10 := .N, hid]
  dt[ageCAT6 == 11, age11 := .N, hid]
  dt[ageCAT6 == 12, age12 := .N, hid]
  dt[ageCAT6 == 13, age13 := .N, hid]
  dt[ageCAT6 == 14, age14 := .N, hid]
  dt[ageCAT6 == 15, age15 := .N, hid]
  dt[ageCAT6 == 16, age16 := .N, hid]
  dt[ageCAT6 == 17, age17 := .N, hid]
  dt[ageCAT6 == 18, age18 := .N, hid]
  dt[, person_num:=SPORDER]
  dt[, sporder := SPORDER]
  dt[SPORDER == 1, ageHH := AGEP]
  dt[, ageHHder := max(ageHH, na.rm = T), hid]
  dt[, persons := max(SPORDER), hid]
  dt[SPORDER == 1, ht := HHtype]
  dt[, HHT := max(ht, na.rm = T), hid]
  
  
  addon <- data.table()
  delflag <- data.table()
  
  for (m in mpos) {
    swm <- dt[mpo == m]
    cx(swm)
    
    for (i in 1:3) {
      diff <- adiff[mpo == m & ageCAT6 == i & year == yr, diff]
      if (diff < 0) {
        sN <- abs(diff)
        m1 <- add.sample[[i]]
        if (lun(m1$pid) < sN) {
          m1 <- swm[ageCAT6 == i  & children < 4 &
                       adults >= 1 & ageHHder %in% 20:60]
        }
        if (lun(m1$pid) < sN) {
          m1 <- swm[ageCAT6 == i  & children <= 4]
        }
        if (lun(m1$pid) < sN) {
          m1 <- swm[ageCAT6 == i]
        }
        hhs <- sample(unique(m1$hid), min(lun(m1$hid), sN))
        if (lun(m1$hid) < sN) {
          flaglist <-
            c(flaglist, data.table(
              flag = paste0(
                'WARNING - Insufficient sample size for duplication ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i
              )
            ))
          message(
            paste0(
              'WARNING - Insufficient sample size for duplication ',
              m,
              ' ',
              yr,
              ' - ',
              'ageCAT == ',
              i
            )
          )
        }
        samp <- m1[hid %in% hhs & ageCAT6 == i]
        samp1 <- samp[get(paste0('age', i)) == 1]
        samp2 <-
          samp[get(paste0('age', i)) > 1, .SD[sample(.N, 1)], hid]
        addon <- rbind(addon, samp1, samp2)
      }
      if (diff > 0) {
        sN <- abs(diff)
        m1 <- sub.sample[[i]]
        m1 <- m1[!hid %in% delflag$hid]
        m1[,maxsp:=max(SPORDER),hid]
        m1 <- m1[SPORDER==maxsp]
        hhs <- sample(unique(m1$pid),  min(lun(m1$pid), sN))
        if (lun(m1$pid) < sN) {
          flaglist <-
            c(flaglist, data.table(
              flag = paste0(
                'WARNING - Insufficient sample size for duplication ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i,
                ' --- diff = ',
                diff,
                ' --- sample = ',
                dim(m1)[1]
              )
            ))
          message(
            paste0(
              'WARNING - Insufficient sample size for duplication ',
              m,
              ' ',
              yr,
              ' - ',
              'ageCAT == ',
              i,
              ' --- diff = ',
              diff,
              ' --- sample = ',
              dim(m1)[1]
            )
          )
        }
        samp <- m1[pid %in% hhs]
        samp[, delete := 1]
        delflag <- rbind(delflag, samp[, .(pid, hid, delete)])
      }
    }
    
    for (i in 4) {
      diff <- adiff[mpo == m & ageCAT6 == i & year == yr, diff]
      if (diff < 0) {
        sN <- abs(diff)
        m1 <- add.sample[[i]]
        hhs <- sample(unique(m1$hid), min(lun(m1$hid), sN))
        if (lun(m1$hid) < sN) {
          flaglist <-
            c(flaglist, data.table(
              flag = paste0(
                'WARNING - Insufficient sample size for duplication ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i
              )
            ))
          message(
            paste0(
              'WARNING - Insufficient sample size for duplication ',
              m,
              ' ',
              yr,
              ' - ',
              'ageCAT == ',
              i
            )
          )
        }
        samp1 <-
          m1[hid %in% hhs &
               HHT == 1 & ageCAT6 == i, .SD[sample(.N, 1)], hid]
        samp1[, age := sample(15:18, 1)]
        samp2 <-
          m1[hid %in% hhs &
               HHT == 3 &
               ageCAT6 == i, .SD[sample(.N, 1)], hid]
        samp2[, age := 19][, ageCAT6 := 4]
        addon <- rbind(addon, samp1, samp2)
      }
      if (diff > 0) {
        sN <- abs(diff)
        m1 <- sub.sample[[i]]
        m1 <- m1[!hid %in% delflag$hid]
        m1[,maxsp:=max(SPORDER),hid]
        m1 <- m1[SPORDER==maxsp]
        hhs <- sample(unique(m1$pid),  min(lun(m1$pid), sN))
        if (lun(m1$pid) < sN) {
          flaglist <-
            c(flaglist, data.table(
              flag = paste0(
                'WARNING - Insufficient sample size for duplication ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i,
                ' --- diff = ',
                diff,
                ' --- sample = ',
                dim(m1)[1]
              )
            ))
          message(
            paste0(
              'WARNING - Insufficient sample size for duplication ',
              m,
              ' ',
              yr,
              ' - ',
              'ageCAT == ',
              i,
              ' --- diff = ',
              diff,
              ' --- sample = ',
              dim(m1)[1]
            )
          )
        }
        samp <- m1[pid %in% hhs]
        samp[, delete := 1]
        delflag <- rbind(delflag, samp[, .(pid, hid, delete)])
      }
    }
    for (i in 5:10) {
      diff <- adiff[mpo == m & ageCAT6 == i & year == yr, diff]
      if (diff < 0) {
        sN <- abs(diff)
        if (m %in% c('CCC',
                     'FRCOG',
                     'MRPC',
                     'MVC',
                     'NPEDC',
                     'NMCOG',
                     'OCPC',
                     'PVPC')) {
          m1 <- swm[ageCAT6 == i & HHT == 3]
          if (lun(m1$pid) < sN) {
              m1 <- swm[ageCAT6 == i & HHT < 5]
            }
          hhs <- sample(unique(m1$pid), min(lun(m1$pid), sN))
          samp <- m1[pid %in% hhs]
          if (lun(m1$pid) < sN) {
            flaglist <-
              c(flaglist, data.table(
                flag = paste0(
                  'WARNING - Insufficient sample size for duplication ',
                  m,
                  ' ',
                  yr,
                  ' - ',
                  'ageCAT == ',
                  i,
                  ' --- diff = ',
                  diff,
                  ' --- sample = ',
                  dim(m1)[1]
                )
              ))
            message(
              paste0(
                'WARNING - Insufficient sample size for duplication ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i,
                ' --- diff = ',
                diff,
                ' --- sample = ',
                dim(m1)[1]
              )
            )
          }
        }
        if (!m %in% c('CCC',
                      'FRCOG',
                      'MRPC',
                      'MVC',
                      'NPEDC',
                      'NMCOG',
                      'OCPC',
                      'PVPC')) {
          m1 <- add.sample[[i]]
          hhs <- sample(unique(m1$pid), min(lun(m1$pid), sN))
          if (lun(m1$pid) < sN) {
            flaglist <-
              c(flaglist, data.table(
                flag = paste0(
                  'WARNING - Insufficient sample size for duplication ',
                  m,
                  ' ',
                  yr,
                  ' - ',
                  'ageCAT == ',
                  i,
                  ' --- diff = ',
                  diff,
                  ' --- sample = ',
                  dim(m1)[1]
                )
              ))
            message(
              paste0(
                'WARNING - Insufficient sample size for duplication ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i,
                ' --- diff = ',
                diff,
                ' --- sample = ',
                dim(m1)[1]
              )
            )
          }
          samp <- m1[pid %in% hhs]
        }
        addon <- rbind(addon, samp)
      }
        if (diff > 0) {
          sN <- abs(diff)
          m1 <- sub.sample[[i]]
          m1 <- m1[!hid %in% delflag$hid]
          m1[,maxsp:=max(SPORDER),hid]
          m1 <- m1[SPORDER==maxsp]
          hhs <- sample(unique(m1$pid),  min(lun(m1$pid), sN))
          if (lun(m1$pid) < sN) {
            flaglist <-
              c(flaglist, data.table(
                flag = paste0(
                  'WARNING - Insufficient sample size for duplication ',
                  m,
                  ' ',
                  yr,
                  ' - ',
                  'ageCAT == ',
                  i,
                  ' --- diff = ',
                  diff,
                  ' --- sample = ',
                  dim(m1)[1]
                )
              ))
            message(
              paste0(
                'WARNING - Insufficient sample size for duplication ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i,
                ' --- diff = ',
                diff,
                ' --- sample = ',
                dim(m1)[1]
              )
            )
          }
          samp <- m1[pid %in% hhs]
          samp[, delete := 1]
          delflag <- rbind(delflag, samp[, .(pid, hid, delete)])
        }
      }
      
      
      for (i in 11:18) {
        diff <- adiff[mpo == m & ageCAT6 == i & year == yr, diff]
        if (diff < 0) {
          sN <- abs(diff)
          m1 <- add.sample[[i]]
          hhs <- sample(unique(m1$pid), min(lun(m1$pid), sN))
          if (lun(m1$pid) < sN) {
            m1 <- swm[ageCAT6 == i & HHT == 3]
          }
          if (lun(m1$pid) < sN) {
            m1 <- swm[ageCAT6 == i & HHT < 5]
          }
          if (lun(m1$pid) < sN) {
            m1 <- swm[ageCAT6 == i]
          }
          if (lun(m1$pid) < sN) {
            flaglist <-
              c(flaglist, data.table(
                flag = paste0(
                  'WARNING - Insufficient sample size for duplication ',
                  m,
                  ' ',
                  yr,
                  ' - ',
                  'ageCAT == ',
                  i,
                  ' --- diff = ',
                  diff,
                  ' --- sample = ',
                  dim(m1)[1]
                )
              ))
            message(
              paste0(
                'WARNING - Insufficient sample size for duplication ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i,
                ' --- diff = ',
                diff,
                ' --- sample = ',
                dim(m1)[1]
              )
            )
          }
          samp <- m1[pid %in% hhs]
          addon <- rbind(addon, samp)
        }
        if (diff > 0) {
          sN <- abs(diff)
          m1 <- sub.sample[[i]]
          m1 <- m1[!hid %in% delflag$hid]
          m1[,maxsp:=max(SPORDER),hid]
          m1 <- m1[SPORDER==maxsp]
          # if (lun(m1$pid) < sN){
          #   m1 <- swm[ageCAT6 == i & person_num > 1]
          # }
          hhs <- sample(unique(m1$pid),  min(lun(m1$pid), sN))
          if (lun(m1$pid) < sN) {
            flaglist <-
              c(flaglist, data.table(
                flag = paste0(
                  'WARNING - Insufficient sample size for duplication ',
                  m,
                  ' ',
                  yr,
                  ' - ',
                  'ageCAT == ',
                  i,
                  ' --- diff = ',
                  diff,
                  ' --- sample = ',
                  dim(m1)[1]
                )
              ))
            message(
              paste0(
                'WARNING - Insufficient sample size for duplication ',
                m,
                ' ',
                yr,
                ' - ',
                'ageCAT == ',
                i,
                ' --- diff = ',
                diff,
                ' --- sample = ',
                dim(m1)[1]
              )
            )
          }
          samp <- m1[pid %in% hhs]
          samp[, delete := 1]
          delflag <- rbind(delflag, samp[, .(pid, hid, delete)])
        }
      }
        message(paste0(m, ' complete ', yr))
        rm(add.sample,sub.sample)
    }
  
    # Join IDs of persons to delete back to main data.table and then delete those rows
    setkey(delflag, pid)
    setkey(dt, pid)
    dt2 <- left_join(dt,delflag,by=c('pid','hid')) |> setDT()
    
    dt2 <- dt2[is.na(delete)]
    
    # Join person records to append to data.table
    addon[, addflag := 1]
    dt3 <- rbind(dt2, addon, fill = T)
    
    # Recalculate person and household characteristics
    setorder(dt3, hid, person_num)
    dt3[, person_num := 1:.N, hid]
    dt3[, sporder := person_num]
    dt3[, SPORDER := person_num]
    dt3[, pid := paste0(hid, '_', person_num)]
    dt3[, persons := .N, hid]
    dt3[, workers := sum(worker), hid]
    dt3[PINCP<0, PINCP:=0]
    dt3[, pinc2021:= as.numeric(ADJINC) * PINCP]
    dt3[, HINCP.adj.2021USD := sum(pinc2021), hid]
    dt3[, HHsize:=max(sporder),hid]
    dt3[, age:=AGEP]
    
    dt3 <- dt3 %>% mutate(
      child = case_when(age < 18 ~ 1,
                        age >= 18 ~ 0)) |>  
      setDT()
    # Count children and adults in each household
    dt3[, children_in_HH := sum(child, na.rm = T), hid]
    dt3[, children := sum(child, na.rm = T), hid]
    dt3[child == 0, adult := 1][is.na(adult), adult := 0]
    dt3[, adults := sum(adult, na.rm = T), hid]
    dt3[,size2:=HHsize]
    dt3[size2>7,size2:=7]
    
    # Persons by 5-yr age categories
    dt3[age < 5 , ageCAT6 := 1]
    dt3[age %in% 5:9 , ageCAT6 := 2]
    dt3[age %in% 10:14 , ageCAT6 := 3]
    dt3[age %in% 15:19 , ageCAT6 := 4]
    dt3[age %in% 20:24 , ageCAT6 := 5]
    dt3[age %in% 25:29 , ageCAT6 := 6]
    dt3[age %in% 30:34 , ageCAT6 := 7]
    dt3[age %in% 35:39 , ageCAT6 := 8]
    dt3[age %in% 40:44 , ageCAT6 := 9]
    dt3[age %in% 45:49 , ageCAT6 := 10]
    dt3[age %in% 50:54 , ageCAT6 := 11]
    dt3[age %in% 55:59 , ageCAT6 := 12]
    dt3[age %in% 60:64 , ageCAT6 := 13]
    dt3[age %in% 65:69 , ageCAT6 := 14]
    dt3[age %in% 70:74 , ageCAT6 := 15]
    dt3[age %in% 75:79 , ageCAT6 := 16]
    dt3[age %in% 80:84 , ageCAT6 := 17]
    dt3[age >= 85 , ageCAT6 := 18]
    dt3[,HHtype:=HHT]
    dt3[,WGTP:=new_WGTP]
    
    dt3 <- dt3 |> mutate(
      HHtype.chk = case_when(
      children >= 1 & persons > 1 ~ 1,
      children == 0 & persons > 1 ~ 3,
      persons == 1 ~ 5
    )
    ) |> setDT()
    
    bad <- dt3[HHT!=HHtype.chk]
    
    
    cf <- dt3[,lapply(.SD,sum),ageCAT6,.SDcols='new_WGTP']
    setnames(cf,'new_WGTP','pop_in_HH')
    cf[,year:=yr][,pop_in_HH:=round(pop_in_HH)]
    setkey(cf,ageCAT6)
    ad <- adiff[year==yr,lapply(.SD,sum),ageCAT6,.SDcols = c('umdi_hh_pop','urbansim_hh_pop')]
    setkey(ad,ageCAT6)
    
    cc <- ad[cf]
    cc[,diff:=pop_in_HH-umdi_hh_pop]
    
    ct <- dt3[,lapply(.SD,sum),.(ageCAT6,HHT),.SDcols='new_WGTP']
    setnames(ct,'new_WGTP','pop_in_HH')
    ct[,year:=yr][,pop_in_HH:=round(pop_in_HH)]
    setorder(ct,ageCAT6,HHT)
    
    mf <- dt3[,lapply(.SD,sum),.(mpo,ageCAT6),.SDcols='new_WGTP']
    setnames(mf,'new_WGTP','pop_in_HH')
    mf[,year:=yr][,pop_in_HH:=round(pop_in_HH)]
    setkey(mf,mpo,ageCAT6)
    md <- adiff[year==yr,lapply(.SD,sum),.(mpo,ageCAT6),.SDcols = c('umdi_hh_pop','urbansim_hh_pop')]
    setkey(md,mpo,ageCAT6)
    
  
    mc <- md[mf]
    mc[,diff:=pop_in_HH-umdi_hh_pop]
    setnames(mc,c('mpo','ageCAT6','umdi_pop_in_HH','uncorrected_pop_in_HH','corrected_pop_in_HH','year','corrected_minus_umdi'))
    
    # if (reb=='MAPC101'){
    #   setnames(dt3,'mpo','rpa_acr')
    #   setnames(mc,'mpo','rpa_acr')
    # }
    
    fwrite(mc,paste0(output_path,'corrected_pop_vs_UMDI_by_ageCAT6_HHtype_',reb,'_',scen,'_',yr,'.csv'))
    
    
    saveRDS(
      dt3,
      paste0(output_path,
        'processor_corrected_PUMS_',reb,'_',scen,'_',yr,'.rds'
      )
    )
message(paste(yr,'Complete'))
}

# Use processor-corrected outputs from Scenario 1 for 2020 for all Scenarios
# copy file and rename
file.copy(
  paste0(
    workpath,
    'Scenario_1/Output_Files_UPDATE/processor_corrected_PUMS_',
    reb,
    '_Scenario_1_2020.rds'
  ),
  paste0(
    output_path,
    'processor_corrected_PUMS_',
    reb,
    '_',
    scen,
    '_2020.rds'
  )
)

```


# 7.4 - Load Shelter population for SHoP Scenarios 2 and 3

```{r}


if (scen %in% c('Scenario_2', 'Scenario_3')) {
  # Read in the shelter household sample
  
  shelter.pums_cleaned <- fread(paste0(shelter_path, "cleaned.shelter.pums.sample_run.1.csv")) |>
    mutate(
      # Align SERIALNO variables
      SERIALNO = SERIALNO.adj,
      hid = SERIALNO.adj,
      HHsize = household.size,
      PUMA = as.numeric(PUMA),
      HISP = as.numeric(HISP),
      HHLDRHISP = as.numeric(HHLDRHISP),
      HHT2 = as.numeric(HHT2),
      # Creates a group designation based on household income.
      Income_Group = case_when(
        HINCP.adj <= 35000 ~ "<=$35,000",
        HINCP.adj > 35000 & HINCP.adj <= 75000 ~ "$35,001 - $75,000",
        HINCP.adj > 75000 &
          HINCP.adj <= 125000 ~ "$75,001 - $125,000",
        HINCP.adj > 125000 &
          HINCP.adj <= 225000 ~ "$125,001 - $225,000",
        HINCP.adj > 225000 ~ ">$225,000",
        .default = "Unknown"
      ),
      # Creates a group designation based on the age of householder.
      Age_Group = case_when(
        ageCAT6 == 1 ~ "00-04",
        ageCAT6 == 2 ~ "05-09",
        ageCAT6 == 3 ~ "10-14",
        ageCAT6 == 4 ~ "15-19",
        ageCAT6 == 5 ~ "20-24",
        ageCAT6 == 6 ~ "25-29",
        ageCAT6 == 7 ~ "30-34",
        ageCAT6 == 8 ~ "35-39",
        ageCAT6 == 9 ~ "40-44",
        ageCAT6 == 10 ~ "45-49",
        ageCAT6 == 11 ~ "50-54",
        ageCAT6 == 12 ~ "55-59",
        ageCAT6 == 13 ~ "60-64",
        ageCAT6 == 14 ~ "65-69",
        ageCAT6 == 15 ~ "70-74",
        ageCAT6 == 16 ~ "75-79",
        ageCAT6 == 17 ~ "80-84",
        ageCAT6 == 18 ~ "85 plus",
        .default = "Unknown"
      ),
      # Categorizing worker in household categories:
      WRKHH = case_when(
        Worker_Total == 0 ~ 0,
        Worker_Total == 1 ~ 1,
        Worker_Total == 2 ~ 2,
        Worker_Total == 3 ~ 3,
        Worker_Total == 4 ~ 4,
        Worker_Total == 5 ~ 5,
        Worker_Total == 6 ~ 6,
        Worker_Total == 7 ~ 7,
        Worker_Total == 8 ~ 8,
        Worker_Total == 9 ~ 9,
        Worker_Total == 10 ~ 10,
      ),
      # Creates a group designation based on the number of workers in a household.
      WRKHH_Group = case_when(
        Worker_Total == 0 ~ "No Workers",
        Worker_Total == 1 ~ "One Worker",
        Worker_Total == 2 ~ "Two Workers",
        Worker_Total >= 3 ~ "Three or More Workers",
        .default = "Unknown"
      ),
      # Creates a group designation based on the number of children in a household.
      CHILDHH_Group = case_when(
        Child_Total == 0 ~ "No Children",
        Child_Total == 1 ~ "One Child",
        Child_Total == 2 ~ "Two Children",
        Child_Total == 3 ~ "Three Children",
        Child_Total >= 4 ~ "Four or More Children",
        .default = "Unknown"
      ),
      # Creates a group designation based on the number of people in a household.
      Size_Group = case_when(
        NP == 1 ~ "One Person",
        NP == 2 ~ "Two person",
        NP == 3 ~ "Three Person",
        NP == 4 ~ "Four Person",
        NP == 5 ~ "Five Person",
        NP == 6 ~ "Six Person",
        NP == 7 ~ "Seven Person",
        NP == 8 ~ "Eight Person",
        NP == 9 ~ "Nine Person",
        NP == 10 ~ "Ten Person",
        NP == 11 ~ "Eleven Person",
        NP == 12 ~ "Twelve Person",
        NP == 13 ~ "Thirteen Person",
        NP == 14 ~ "Fourteen Person",
        NP >= 15 ~ "Fifteen Persons or More Persons",
        .default = "Unknown"
      ),
      # Creates a group designation based on the type of household.
      HHtype_Group = case_when(
        HHtype == 1 ~ "Household w Children",
        HHtype == 3 ~ "Household w 2 plus Adults wo Children",
        HHtype == 5 ~ "Single Person Household",
        .default = "Unknown"
      ),
      SHELT = 1
    )
  
  if (reb == 'MAPC97') {
    setnames(shelter.pums_cleaned, 'rpa_acr', 'mpo')
  }
}


# Need hid, PINCP, HHsize
```
# 7.5 - Combine with pop corrected reweighter outputs
```{r}

# Set a list of years corresponding to the projection years considered

years <- c(2030, 2035, 2040, 2045, 2050)
for (yr in years) {
  tmp.df <- readRDS(paste0(
    output_path,
    "processor_corrected_PUMS_",
    reb,
    "_",
    scen,
    "_",
    yr,
    ".rds"
  )) |>
    mutate(SPORDER = sporder) |>
    setDT()
  tmp.df[pinc2021 < 0, pinc2021 := 0]
  tmp.df[, HINCP.adj.2021USD := sum(pinc2021), hid]
  
  if (reb == 'MAPC97') {
    tmp.df[, rpa_acr := NULL]
  }
  if (reb == 'MAPC101') {
    tmp.df[, mpo := NULL]
  }
  
  rwt.pums.df.join <- tmp.df |>
    mutate(
      # Creates a group designation based on the number of children in a household.
      CHILDHH_Group = case_when(
        children_in_HH == 0 ~ "No Children",
        children_in_HH == 1 ~ "One Child",
        children_in_HH == 2 ~ "Two Children",
        children_in_HH == 3 ~ "Three Children",
        children_in_HH >= 4 ~ "Four or More Children",
        .default = "Unknown"
      ),
      # Creates a group designation based on household income.
      Income_Group = case_when(
        HINCP.adj.2021USD <= 35000 ~ "<=$35,000",
        HINCP.adj.2021USD > 35000 &
          HINCP.adj.2021USD <= 75000 ~ "$35,001 - $75,000",
        HINCP.adj.2021USD > 75000 &
          HINCP.adj.2021USD <= 125000 ~ "$75,001 - $125,000",
        HINCP.adj.2021USD > 125000 &
          HINCP.adj.2021USD <= 225000 ~ "$125,001 - $225,000",
        HINCP.adj.2021USD > 225000 ~ ">$225,000",
        .default = "Unknown"
      ),
      # Creates a group designation based on the age of householder.
      Age_Group = case_when(
        ageCAT6 == 1 ~ "00-04",
        ageCAT6 == 2 ~ "05-09",
        ageCAT6 == 3 ~ "10-14",
        ageCAT6 == 4 ~ "15-19",
        ageCAT6 == 5 ~ "20-24",
        ageCAT6 == 6 ~ "25-29",
        ageCAT6 == 7 ~ "30-34",
        ageCAT6 == 8 ~ "35-39",
        ageCAT6 == 9 ~ "40-44",
        ageCAT6 == 10 ~ "45-49",
        ageCAT6 == 11 ~ "50-54",
        ageCAT6 == 12 ~ "55-59",
        ageCAT6 == 13 ~ "60-64",
        ageCAT6 == 14 ~ "65-69",
        ageCAT6 == 15 ~ "70-74",
        ageCAT6 == 16 ~ "75-79",
        ageCAT6 == 17 ~ "80-84",
        ageCAT6 == 18 ~ "85 plus",
        .default = "Unknown"
      ),
      # Creates a group designation based on the type of household.
      HHtype_Group = case_when(
        HHtype == 1 ~ "Household w Children",
        HHtype == 3 ~ "Household w 2 plus Adults wo Children",
        HHtype == 5 ~ "Single Person Household",
        .default = "Unknown"
      ),
      # Creates a group designation based on the number of people in a household.
      Size_Group = case_when(
        HHsize == 1 ~ "One Person",
        HHsize == 2 ~ "Two person",
        HHsize == 3 ~ "Three Person",
        HHsize == 4 ~ "Four Person",
        HHsize == 5 ~ "Five Person",
        HHsize == 6 ~ "Six Person",
        HHsize == 7 ~ "Seven Person",
        HHsize == 8 ~ "Eight Person",
        HHsize == 9 ~ "Nine Person",
        HHsize == 10 ~ "Ten Person",
        HHsize == 11 ~ "Eleven Person",
        HHsize == 12 ~ "Twelve Person",
        HHsize == 13 ~ "Thirteen Person",
        HHsize == 14 ~ "Fourteen Person",
        HHsize >= 15 ~ "Fifteen or More Persons",
        .default = "Unknown"
      ),
      # Creates a group designation based on the tenure of the household.
      TEN_Group = case_when(tenure == 1 ~ "Own", tenure == 2 ~ "Rent", .default = "Unknown"),
      # Categorizing worker categories:
      # (0) no workers
      # (1) one worker
      # (2) two workers
      # (3) three or more workers
      WRKHH = case_when(
        workers == 0 ~ 0,
        workers == 1 ~ 1,
        workers == 2 ~ 2,
        workers == 3 ~ 3,
        workers == 4 ~ 4,
        workers == 5 ~ 5,
        workers == 6 ~ 6,
        workers == 7 ~ 7,
        workers == 8 ~ 8,
        workers == 9 ~ 9,
        workers >= 10 ~ 10,
      ),
      # Creates a group designation based on the number of workers in a household.
      WRKHH_Group = case_when(
        workers == 0 ~ "No Workers",
        workers == 1 ~ "One Worker",
        workers == 2 ~ "Two Workers",
        workers >= 3 ~ "Three or More Workers",
        .default = "Unknown"
      ),
      SHELT = 0
    )
  
  rm(tmp.df)
  gc()
  
  
  rwt.pums.df.join <- rbind(rwt.pums.df.join, shelter.pums_cleaned, fill =
                              T) |>
    mutate(year=yr) |> 
    setDT()
  
  rwt.pums.df.join[is.na(children_in_HH), children_in_HH := Child_Total]
  
  saveRDS(
    rwt.pums.df.join,
    paste0(
      output_path,
      "processor_corrected_PUMS_",
      reb,
      "_",
      scen,
      "_with_ShelterPop_",
      yr,
      ".rds"
    )
  )
  print(paste0(yr, ' Complete'))
}



```



# QA of MAPC97 vs MAPC101

```{r}

vs <- data.table(mpo=mpos)

for (yr in c(2030,2035,2040,2045,2050)){
  x <- fread(paste0(output_path,'reweighted_PUMS_',scen,'_MAPC97_',hdshp,'_',yr,'.csv'))
  # x <- readRDS(paste0(output_path,'processor_corrected_PUMS_MAPC97_',scen,'_',yr,'.rds'))
  hh <- x[SPORDER==1, round(sum(new_WGTP)), mpo] 
  setnames(hh,c('mpo',paste0('hh_MAPC97_',yr)))
  vs <- left_join(vs,hh,by='mpo')
  rm(x);gc()
}

for (yr in c(2030,2035,2040,2045,2050)){
  x <- fread(paste0(output_path,'reweighted_PUMS_',scen,'_MAPC101_',hdshp,'_',yr,'.csv'))
  # x <- readRDS(paste0(output_path,'processor_corrected_PUMS_MAPC101_',scen,'_',yr,'.rds'))
  hh <- x[SPORDER==1, round(sum(new_WGTP)), mpo] 
  setnames(hh,c('mpo',paste0('hh_MAPC101_',yr)))
  vs <- left_join(vs,hh,by='mpo')
  rm(x);gc()
}

vs <- vs |> 
  mutate(diff2030 = hh_MAPC101_2030 - hh_MAPC97_2030,
         diff2035 = hh_MAPC101_2035 - hh_MAPC97_2035,
         diff2040 = hh_MAPC101_2040 - hh_MAPC97_2040,
         diff2045 = hh_MAPC101_2045 - hh_MAPC97_2045,
         diff2050 = hh_MAPC101_2050 - hh_MAPC97_2050
         ) |> 
  setDT()

for (yr in c(2030, 2035, 2040, 2045, 2050)) {
  x <- readRDS(
    paste0(
      root,"Current_Projects/Projections/Projections_2028/Data/01_Regional_Control_Totals/SHoP_Scenarios/Reweighter_Outputs/",scen,"/bad_MAPC101/processor_corrected_PUMS_",scen,"_",
      yr,
      ".rds"
    ))
    hh <- x[SPORDER == 1, round(sum(new_WGTP)), mpo]
    setnames(hh, c('mpo', paste0('oldSHoP_hh_MAPC101_', yr)))
    vs <- left_join(vs, hh, by = 'mpo')
    rm(x)
    gc()
}

vs <- vs |> 
  mutate(old_diff2030 = hh_MAPC101_2030 - oldSHoP_hh_MAPC101_2030,
         old_diff2035 = hh_MAPC101_2035 - oldSHoP_hh_MAPC101_2035,
         old_diff2040 = hh_MAPC101_2040 - oldSHoP_hh_MAPC101_2040,
         old_diff2045 = hh_MAPC101_2045 - oldSHoP_hh_MAPC101_2045,
         old_diff2050 = hh_MAPC101_2050 - oldSHoP_hh_MAPC101_2050
         ) |> 
  setDT()
```