---
title: "03_Reweighter.Output.Summaries"
format: html
editor: visual
---

# 0.0 Environment Set Up

# 0.1 Filepath Set Up

```{r}
#===============================================================================
#TODO: 
# 1. Create a way to loop through:
# - MAPC vintages (101 or 97)
# 2. Create output table function 
# 3. Create way to loop through Scenarios for Output tables
# 4. QA/QC of Tables based on reweighter inputs
# 3. Writing to workbook
# 4. Data Visualizations
#===============================================================================

### Set Scenario Names ###

#Population Scenario

# scen <- 'Scenario_1'
# scen <- 'Scenario_2'
scen <- 'Scenario_3'
# scen <- 'Scenario_4'

#Headship rate Scenario
if (scen == 'Scenario_1'){
  hdshp <- 'acs1721'
}
if (scen != 'Scenario_1'){
  hdshp <- 'scen2'
}

reb <- 'MAPC101'
# reb <- 'MAPC97'


#Use when Rproject is open
#root <- '../../../../Data/Working/Regional_Control_Totals/'

# Base root. Toggle between S: and K: drive depedning on mapped paths on your computer
# base <- "K:/DataServices/Projects/Current_Projects/"
base <- "S:/Network Shares/K Drive/DataServices/Projects/Current_Projects/"

# General K drive filepath
root <- paste0(base, "Housing/StatewideHousingPlan/04_Analysis/Data/Working/Regional_Control_Totals/")

# Set output filepath
# output_path <- paste0(base, "Housing/StatewideHousingPlan/04_Analysis/Data/Final_Outputs/", scen, "/UPDATE_w_shelter2/")

output_path <- paste0(base, "Housing/StatewideHousingPlan/04_Analysis/Data/Final_Outputs/", scen, "/For_Bob_20260420/")
dir.create(output_path, showWarnings = F)

# Set Shelter Households path
shelter_path <- paste0(
  base,
  "Housing/StatewideHousingPlan/04_Analysis/Data/Final_Outputs/Shelter_Households/"
)

# Corrected Reweighter files path
rwt_path <- paste0(
  base,
  "Housing/StatewideHousingPlan/04_Analysis/Data/Working/Reweighter/",
  scen,
  '/Output_Files_UPDATE/'
)

# Set knitr path
knitr::opts_knit$set(root.dir = root)
```

# 0.2 Set Package Dependencies + Set Environment Variables

```{r}
library(tidycensus)
library(tidyverse)
library(data.table)
library(reticulate)
library(mapcdatakeys)
library(janitor)
library(readxl)

##get rid of scientific notation
options(scipen = 999)
#Set random seed
set.seed(351)

# Create RPA list
rpa.list <- mapcdatakeys::all_muni_data_keys |> select(rpa_acr) |> distinct(rpa_acr) |> pull()

```

# Load Output Data

# 1.1 Load Outputs from the Reweighter

```{r results = 'hide'}
# Set a list of years corresponding to the projection years considered

years <- c(2030,2035,2040,2045,2050)
# years <- c(2025, 2035)

read_rwt.files <- function(scenario, year) {
  readRDS(
    paste0(
      rwt_path,
      "processor_corrected_PUMS_",
      reb,
      "_",
      scenario,
      "_",
      year,
      ".rds"
    )
  ) 
}

rwt.pums.df <- data.table()

for (j in years) {
  # Reads in the reweighted PUMS data by year.
  tmp.rwt <- read_rwt.files(scenario = scen, year = j) |>
    select(
      SERIALNO,
      sporder,
      mpo,
      year,
      HHsize,
      child,
      workers,
      ageCAT6,
      HHtype,
      HINCP.adj.2021USD,
      tenure,
      children_in_HH,
      hid,
      pinc2021,
      new_WGTP
    ) |>
    mutate(SPORDER = sporder) |>
    setDT()
  tmp.rwt[pinc2021 < 0, pinc2021 := 0]
  tmp.rwt[, HINCP.adj.2021USD := sum(pinc2021), hid]
  
  rwt.pums.df <- rbind(rwt.pums.df, tmp.rwt)
  rm(tmp.rwt)
  gc()
  message(paste('Year', j, 'Loaded'))
}

rwt.pums.df.join <- rwt.pums.df |>
  mutate(
    # Creates a group designation based on the number of children in a household.
    CHILDHH_Group = case_when(
      children_in_HH == 0 ~ "No Children",
      children_in_HH == 1 ~ "One Child",
      children_in_HH == 2 ~ "Two Children",
      children_in_HH == 3 ~ "Three Children",
      children_in_HH >= 4 ~ "Four or More Children",
      .default = "Unknown"
    ),
    # Creates a group designation based on household income.
    Income_Group = case_when(
      HINCP.adj.2021USD <= 35000 ~ "<=$35,000",
      HINCP.adj.2021USD > 35000 &
        HINCP.adj.2021USD <= 75000 ~ "$35,001 - $75,000",
      HINCP.adj.2021USD > 75000 &
        HINCP.adj.2021USD <= 125000 ~ "$75,001 - $125,000",
      HINCP.adj.2021USD > 125000 &
        HINCP.adj.2021USD <= 225000 ~ "$125,001 - $225,000",
      HINCP.adj.2021USD > 225000 ~ ">$225,000",
      .default = "Unknown"
    ),
    # Creates a group designation based on the age of householder.
    Age_Group = case_when(
      ageCAT6 == 1 ~ "00-04",
      ageCAT6 == 2 ~ "05-09",
      ageCAT6 == 3 ~ "10-14",
      ageCAT6 == 4 ~ "15-19",
      ageCAT6 == 5 ~ "20-24",
      ageCAT6 == 6 ~ "25-29",
      ageCAT6 == 7 ~ "30-34",
      ageCAT6 == 8 ~ "35-39",
      ageCAT6 == 9 ~ "40-44",
      ageCAT6 == 10 ~ "45-49",
      ageCAT6 == 11 ~ "50-54",
      ageCAT6 == 12 ~ "55-59",
      ageCAT6 == 13 ~ "60-64",
      ageCAT6 == 14 ~ "65-69",
      ageCAT6 == 15 ~ "70-74",
      ageCAT6 == 16 ~ "75-79",
      ageCAT6 == 17 ~ "80-84",
      ageCAT6 == 18 ~ "85 plus",
      .default = "Unknown"
    ),
    # Creates a group designation based on the type of household.
    HHtype_Group = case_when(
      HHtype == 1 ~ "Household w Children",
      HHtype == 3 ~ "Household w 2 plus Adults wo Children",
      HHtype == 5 ~ "Single Person Household",
      .default = "Unknown"
    ),
    # Creates a group designation based on the number of people in a household.
    Size_Group = case_when(
      HHsize == 1 ~ "One Person",
      HHsize == 2 ~ "Two person",
      HHsize == 3 ~ "Three Person",
      HHsize == 4 ~ "Four Person",
      HHsize == 5 ~ "Five Person",
      HHsize == 6 ~ "Six Person",
      HHsize == 7 ~ "Seven Person",
      HHsize == 8 ~ "Eight Person",
      HHsize == 9 ~ "Nine Person",
      HHsize == 10 ~ "Ten Person",
      HHsize == 11 ~ "Eleven Person",
      HHsize == 12 ~ "Twelve Person",
      HHsize == 13 ~ "Thirteen Person",
      HHsize == 14 ~ "Fourteen Person",
      HHsize >= 15 ~ "Fifteen or More Persons",
      .default = "Unknown"
    ),
    # Creates a group designation based on the tenure of the household.
    TEN_Group = case_when(tenure == 1 ~ "Own", tenure == 2 ~ "Rent", .default = "Unknown"),
    # Categorizing worker categories:
    # (0) no workers
    # (1) one worker
    # (2) two workers
    # (3) three or more workers
    WRKHH = case_when(
      workers == 0 ~ 0,
      workers == 1 ~ 1,
      workers == 2 ~ 2,
      workers == 3 ~ 3,
      workers == 4 ~ 4,
      workers == 5 ~ 5,
      workers == 6 ~ 6,
      workers == 7 ~ 7,
      workers == 8 ~ 8,
      workers == 9 ~ 9,
      workers >= 10 ~ 10,
    ),
    # Creates a group designation based on the number of workers in a household.
    WRKHH_Group = case_when(
      workers == 0 ~ "No Workers",
      workers == 1 ~ "One Worker",
      workers == 2 ~ "Two Workers",
      workers >= 3 ~ "Three or More Workers",
      .default = "Unknown"
    ),
    SHELT = 0
  )


if (reb=='MAPC101'){
      setnames(rwt.pums.df.join,'mpo','rpa_acr')
}

rm(rwt.pums.df);gc()




```

# 1.2 Load Shelter Households

```{r}

if (scen %in% c('Scenario_2','Scenario_3')){
# Read in the shelter household sample
shelter.pums <- read_csv(paste0(shelter_path, "cleaned.shelter.pums.sample_run.1.csv"))

shelter.pums_cleaned <- shelter.pums |> 
  mutate(
    # Align SERIALNO variables
    SERIALNO = SERIALNO.adj,
    PUMA = as.numeric(PUMA),
    HISP = as.numeric(HISP),
    HHLDRHISP = as.numeric(HHLDRHISP),
    HHT2 = as.numeric(HHT2),
    # Creates a group designation based on household income.
    Income_Group = case_when(
      HINCP.adj <= 35000 ~ "<=$35,000",
      HINCP.adj > 35000 & HINCP.adj <= 75000 ~ "$35,001 - $75,000",
      HINCP.adj > 75000 & HINCP.adj <= 125000 ~ "$75,001 - $125,000",
      HINCP.adj > 125000 & HINCP.adj <= 225000 ~ "$125,001 - $225,000",
      HINCP.adj > 225000 ~ ">$225,000",
      .default = "Unknown"
    ),
    # Creates a group designation based on the age of householder.
    Age_Group = case_when(
      ageCAT6 == 1 ~ "00-04",
      ageCAT6 == 2 ~ "05-09",
      ageCAT6 == 3 ~ "10-14",
      ageCAT6 == 4 ~ "15-19",
      ageCAT6 == 5 ~ "20-24",
      ageCAT6 == 6 ~ "25-29",
      ageCAT6 == 7 ~ "30-34",
      ageCAT6 == 8 ~ "35-39",
      ageCAT6 == 9 ~ "40-44",
      ageCAT6 == 10 ~ "45-49",
      ageCAT6 == 11 ~ "50-54",
      ageCAT6 == 12 ~ "55-59",
      ageCAT6 == 13 ~ "60-64",
      ageCAT6 == 14 ~ "65-69",
      ageCAT6 == 15 ~ "70-74",
      ageCAT6 == 16 ~ "75-79",
      ageCAT6 == 17 ~ "80-84",
      ageCAT6 == 18 ~ "85 plus",
      .default = "Unknown"
    ),
    # Categorizing worker in household categories:
    WRKHH = case_when(
      Worker_Total == 0 ~ 0,
      Worker_Total == 1 ~ 1,
      Worker_Total == 2 ~ 2,
      Worker_Total == 3 ~ 3,
      Worker_Total == 4 ~ 4,
      Worker_Total == 5 ~ 5,
      Worker_Total == 6 ~ 6,
      Worker_Total == 7 ~ 7,
      Worker_Total == 8 ~ 8,
      Worker_Total == 9 ~ 9,
      Worker_Total == 10 ~ 10,
    ),
    # Creates a group designation based on the number of workers in a household.
    WRKHH_Group = case_when(
      Worker_Total == 0 ~ "No Workers",
      Worker_Total == 1 ~ "One Worker",
      Worker_Total == 2 ~ "Two Workers",
      Worker_Total >= 3 ~ "Three or More Workers",
      .default = "Unknown"
    ),
    # Creates a group designation based on the number of children in a household.
    CHILDHH_Group = case_when(
      Child_Total == 0 ~ "No Children",
      Child_Total == 1 ~ "One Child",
      Child_Total == 2 ~ "Two Children",
      Child_Total == 3 ~ "Three Children",
      Child_Total >= 4 ~ "Four or More Children",
      .default = "Unknown"
    ),
    # Creates a group designation based on the number of people in a household.
    Size_Group = case_when(
      NP == 1 ~ "One Person",
      NP == 2 ~ "Two person",
      NP == 3 ~ "Three Person",
      NP == 4 ~ "Four Person",
      NP == 5 ~ "Five Person",
      NP == 6 ~ "Six Person",
      NP == 7 ~ "Seven Person",
      NP == 8 ~ "Eight Person",
      NP == 9 ~ "Nine Person",
      NP == 10 ~ "Ten Person",
      NP == 11 ~ "Eleven Person",
      NP == 12 ~ "Twelve Person",
      NP == 13 ~ "Thirteen Person",
      NP == 14 ~ "Fourteen Person",
      NP >= 15 ~ "Fifteen Persons or More Persons",
      .default = "Unknown"
    ),
    # Creates a group designation based on the type of household.
    HHtype_Group = case_when(
      HHtype == 1 ~ "Household w Children",
      HHtype == 3 ~ "Household w 2 plus Adults wo Children",
      HHtype == 5 ~ "Single Person Household",
      .default = "Unknown"
    ),
    SHELT = 1
  )

rm(shelter.pums)
gc()

# Creates a reference table for the total number of households in each year to compare to subsequent tables as a QA check.
tbl.0 <- shelter.pums_cleaned |> 
  filter(TYPEHUGQ == 1 & SPORDER == 1) |> 
  group_by(rpa_acr) |> 
  summarise(hhds = sum(new_WGTP)) |> 
  ungroup()

print(sum(tbl.0$hhds))

# .1 UrbanSim Household Income Groupings
tbl.1 <- shelter.pums_cleaned |> 
  filter(TYPEHUGQ == 1 & SPORDER == 1) |> 
  group_by(
    rpa_acr,
    Income_Group
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup()

# .2 5-year Age Groupings (ageCAT6)
tbl.2 <- shelter.pums_cleaned |> 
  filter(TYPEHUGQ == 1 & SPORDER == 1) |> 
  group_by(
    rpa_acr,
    Age_Group
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup()

# .3 Household Type
tbl.3 <- shelter.pums_cleaned  |> 
  filter(TYPEHUGQ == 1 & SPORDER == 1) |> 
  group_by(
    rpa_acr,
    HHtype_Group
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup()

# .4 Laborforce
# Note: A bit of a roundabout way of calculating labor force. Its based on the number households with a specific number of workers in them. Households with two workers are multiplied by 2 to get the total number of people in that household in the labor force. "Workers" are people in the LF + military service members (a negligible amount of people in MA).
tbl.4 <- shelter.pums_cleaned  |> 
  filter(TYPEHUGQ == 1 & SPORDER == 1) |> 
  group_by(
    rpa_acr,
    WRKHH
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup() |> 
  mutate(
    lf = WRKHH*hhds
  ) |> 
  group_by(rpa_acr) |> 
  summarise(
    lf = sum(lf)
  ) |> 
  ungroup()
}

```

# 1.3 Add Shelter HH to Reweighter Output

```{r}

if (scen %in% c('Scenario_2','Scenario_3')){

shelter.pums_cleaned.2030 <- shelter.pums_cleaned |> mutate(year = 2030)
shelter.pums_cleaned.2035 <- shelter.pums_cleaned |> mutate(year = 2035)
shelter.pums_cleaned.2040 <- shelter.pums_cleaned |> mutate(year = 2040)
shelter.pums_cleaned.2045 <- shelter.pums_cleaned |> mutate(year = 2045)
shelter.pums_cleaned.2050 <- shelter.pums_cleaned |> mutate(year = 2050)

shelter.pums_cleaned.year <- bind_rows(
  shelter.pums_cleaned.2030,
  shelter.pums_cleaned.2035,
  shelter.pums_cleaned.2040,
  shelter.pums_cleaned.2045,
  shelter.pums_cleaned.2050
)

rm(
  shelter.pums_cleaned.2030,
  shelter.pums_cleaned.2035,
  shelter.pums_cleaned.2040,
  shelter.pums_cleaned.2045,
  shelter.pums_cleaned.2050
)

if (reb == 'MAPC97') {
  setnames(shelter.pums_cleaned.year, 'rpa_acr', 'mpo')
  rwt.pums.df.join <- bind_rows(rwt.pums.df.join, shelter.pums_cleaned.year) |>
    setDT()
}

if (reb == 'MAPC101') {
  rwt.pums.df.join <- bind_rows(rwt.pums.df.join, shelter.pums_cleaned.year) |> 
    setDT()
}

rwt.pums.df.join[is.na(children_in_HH),children_in_HH:=Child_Total]

rm(shelter.pums_cleaned.year)
gc()
}

```

# 2.0 Univariate Output Tables For UrbanSim

-   .1 UrbanSim Household Income Groupings
-   .2 5-year Age Groupings (ageCAT6)
-   .3 Household Type
-   .4 Laborforce in Households
-   .5 Number of Children in a Household
-   .6 Size of Household
-   .7 Tenure
-   .8 Number of Workers in a Household

```{r}
#===============================================================================
# Creates a reference table for the total number of households in each year to compare to subsequent tables as a QA check.
tbl.0 <- rwt.pums.df.join |> 
  #filter(TYPEHUGQ == 1 & SPORDER == 1) |>
  filter(SPORDER == 1) |>
  group_by(
    year
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup()

gc()

#===============================================================================
# .1 UrbanSim Household Income Groupings
if (reb == 'MAPC97') {
tbl.1 <- rwt.pums.df.join |> 
  #filter(TYPEHUGQ == 1 & SPORDER == 1) |> 
  filter(SPORDER == 1) |> 
  group_by(
    mpo,
    year,
    Income_Group
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  )
} else {
  tbl.1 <- rwt.pums.df.join |> 
  filter(SPORDER == 1) |> 
  group_by(
    rpa_acr,
    year,
    Income_Group
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup()
}


tbl.1.test <- tbl.1 |> 
  group_by(
    year
  ) |> 
  summarise(
    hhds = sum(hhds)
  ) |> 
  ungroup()

if (round(sum(tbl.1$hhds)) != round(sum(tbl.1.test$hhds))){
  
  print("STOP! Check table!")
  
} else {
  
  print("Proceed.")
  
}

#Write Output tables to K Drive
write.csv(
  tbl.1,
  paste0(output_path,"/output.tbl_",scen,"_",reb,"_Households.by.IncomeGroups.fix.csv"),
  row.names = FALSE
)
gc()
#===============================================================================
# .2 5-year Age Groupings (ageCAT6)
if (reb == 'MAPC97') {
tbl.2 <- rwt.pums.df.join |> 
  filter(SPORDER == 1) |>
  group_by(
    mpo,
    year,
    Age_Group
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup()
} else {
  tbl.2 <- rwt.pums.df.join |> 
  filter(SPORDER == 1) |>
  group_by(
    rpa_acr,
    year,
    Age_Group
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup()
}

tbl.2.test <- tbl.2 |> 
  group_by(
    year
  ) |> 
  summarise(
    hhds = sum(hhds)
  ) |> 
  ungroup()

if (round(sum(tbl.2$hhds)) != round(sum(tbl.2.test$hhds))){
  
  print("STOP! Check table!")
  
} else {
  
  print("Proceed.")
  
}

#Write Output Tables to K Drive
write.csv(
  tbl.2,
  paste0(output_path,"/output.tbl_",scen,"_",reb,"_Households.by.AgeGroups.csv"),
  row.names = FALSE
)

gc()

#===============================================================================
# .3 Household Type
if (reb == 'MAPC97') {
tbl.3 <- rwt.pums.df.join  |>
  filter(SPORDER == 1) |> 
  group_by(
    mpo,
    year,
    HHtype_Group
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup()
} else {
  tbl.3 <- rwt.pums.df.join  |>
  filter(SPORDER == 1) |> 
  group_by(
    rpa_acr,
    year,
    HHtype_Group
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup()
}

tbl.3.test <- tbl.3 |> 
  group_by(
    year
  ) |> 
  summarise(
    hhds = sum(hhds)
  ) |> 
  ungroup()

if (round(sum(tbl.3$hhds)) != round(sum(tbl.3.test$hhds))){
  
  print("STOP! Check table!")
  
} else {
  
  print("Proceed.")
  
}

# Write output tables to K drive
write.csv(
  tbl.3,
  paste0(output_path,"output.tbl_",scen,"_",reb,"_Households.by.HHtypeGroups1.csv"),
  row.names = FALSE
)
gc()
#===============================================================================
# .4 Laborforce
# Note: A bit of a roundabout way of calculating labor force. Its based on the number households with a specific number of workers in them. Households with two workers are multiplied by 2 to get the total number of people in that household in the labor force. "Workers" are people in the LF + military service members (a negligible amount of people in MA).
if (reb=='MAPC97'){
tbl.4 <- rwt.pums.df.join  |> 
  filter(SPORDER == 1) |> 
  group_by(
    mpo,
    year,
    WRKHH
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup() |> 
  mutate(
    lf = WRKHH*hhds
  ) |> 
  group_by(year, mpo) |> 
  summarise(
    lf = sum(lf)
  ) |> 
  ungroup()

} else {
  filter(SPORDER == 1) |> 
  group_by(
    rpa_acr,
    year,
    WRKHH
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup() |> 
  mutate(
    lf = WRKHH*hhds
  ) |> 
  group_by(year, rpa_acr) |> 
  summarise(
    lf = sum(lf)
  ) |> 
  ungroup()
}
# Write output tables to K drive
write.csv(
  tbl.4,
  paste0(output_path,"/output.tbl_",scen,"_",reb,"_Laborforce.in.Households.csv"),
  row.names = FALSE
)

#===============================================================================
# .5 Number of Children in a Household

if (reb=='MAPC97'){
tbl.5 <- rwt.pums.df.join |>
  filter(SPORDER == 1) |>  
  group_by(
    mpo,
    year,
    CHILDHH_Group
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup()
} else {
  tbl.5 <- rwt.pums.df.join |>
  filter(SPORDER == 1) |>  
  group_by(
    mrpa_acr,
    year,
    CHILDHH_Group
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup()
}

tbl.5.test <- tbl.5 |> 
  group_by(
    year
  ) |> 
  summarise(
    hhds = sum(hhds)
  ) |> 
  ungroup()

if (round(sum(tbl.5$hhds)) != round(sum(tbl.5.test$hhds))){
  
  print("STOP! Check table!")
  
} else {
  
  print("Proceed.")
  
}

# Alternate version no grouping
setDT(tbl.5)

if (reb=='MAPC97'){
tbl.5b <- rwt.pums.df.join |>
  filter(SPORDER == 1) |>  
  group_by(
    mpo,
    year,
    children_in_HH,
    CHILDHH_Group
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup() |> 
  left_join(tbl.5,by=c('mpo','year','CHILDHH_Group')) |> 
  mutate(diff = hhds.x-hhds.y) |> 
  setDT()

tbl.5b[diff!=0,unique(CHILDHH_Group)]
tbl.5b[CHILDHH_Group=='Four or More Children',c4:=sum(hhds.x),.(mpo,year)]
tbl.5b[,diff2:=c4-hhds.y]
tbl.5b[!is.na(diff2),range(diff2)]
setnames(tbl.5b,c('mpo','year','children_in_HH','CHILDHH_Group','HH_ungrouped','HH_grouped','raw_diff','sum_4+','diff_2'))

} else {
  tbl.5b <- rwt.pums.df.join |>
  filter(SPORDER == 1) |>  
  group_by(
    rpa_acr,
    year,
    children_in_HH,
    CHILDHH_Group
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup() |> 
  left_join(tbl.5,by=c('rpa_acr','year','CHILDHH_Group')) |> 
  mutate(diff = hhds.x-hhds.y) |> 
  setDT()

tbl.5b[diff!=0,unique(CHILDHH_Group)]
tbl.5b[CHILDHH_Group=='Four or More Children',c4:=sum(hhds.x),.(rpa_acr,year)]
tbl.5b[,diff2:=c4-hhds.y]
tbl.5b[!is.na(diff2),range(diff2)]
setnames(tbl.5b,c('rpa_acr','year','children_in_HH','CHILDHH_Group','HH_ungrouped','HH_grouped','raw_diff','sum_4+','diff_2'))
}

# Write output tables to K drive
write.csv(
  tbl.5,
  paste0(output_path,"/output.tbl_",scen,"_",reb,"_Households.by.NumberOfChildren.csv"),
  row.names = FALSE
)

gc()
#===============================================================================
# .6 Size of Household
if(reb=='MAPC97'){
tbl.6 <- rwt.pums.df.join  |> 
  filter(SPORDER == 1) |> 
  group_by(
    mpo,
    year,
    Size_Group
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup()
} else {
  tbl.6 <- rwt.pums.df.join  |> 
  filter(SPORDER == 1) |> 
  group_by(
    mrpa_acr,
    year,
    Size_Group
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup()
}
tbl.6.test <- tbl.6 |> 
  group_by(
    year
  ) |> 
  summarise(
    hhds = sum(hhds)
  ) |> 
  ungroup()

if (round(sum(tbl.6$hhds)) != round(sum(tbl.6.test$hhds))){
  
  print("STOP! Check table!")
  
} else {
  
  print("Proceed.")
  
}

# Write output tables to K drive
write.csv(
  tbl.6,
  paste0(output_path,"/output.tbl_",scen,"_",reb,"_Households.by.HHsize.csv"),
  row.names = FALSE
)

gc()
#===============================================================================
# .7 Tenure

if (reb=='MAPC97'){
tbl.7 <- rwt.pums.df.join  |>
  filter(SPORDER == 1) |> 
  group_by(
    mpo,
    year,
    TEN_Group
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup()
} else {
  tbl.7 <- rwt.pums.df.join  |>
  filter(SPORDER == 1) |> 
  group_by(
    rpa_acr,
    year,
    TEN_Group
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup()
}

tbl.7.test <- tbl.7 |> 
  group_by(
    year
  ) |> 
  summarise(
    hhds = sum(hhds)
  ) |> 
  ungroup()

if (round(sum(tbl.7$hhds)) != round(sum(tbl.7.test$hhds))){
  
  print("STOP! Check table!")
  
} else {
  
  print("Proceed.")
  
}

# Write output tables to K drive
write.csv(
  tbl.7,
  paste0(output_path,"/output.tbl_",scen,"_",reb,"_Households.by.Tenure.csv"),
  row.names = FALSE
)

gc()
#===============================================================================
# .8 Number of Workers in a Household
if (reb=='MAPC97'){
tbl.8 <- rwt.pums.df.join  |>
  filter(SPORDER == 1) |> 
  group_by(
    mpo,
    year,
    WRKHH_Group
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup()
}else{
  tbl.8 <- rwt.pums.df.join  |>
  filter(SPORDER == 1) |> 
  group_by(
    rpa_acr,
    year,
    WRKHH_Group
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup()
}
tbl.8.test <- tbl.8 |> 
  group_by(
    year
  ) |> 
  summarise(
    hhds = sum(hhds)
  ) |> 
  ungroup()

if (round(sum(tbl.8$hhds)) != round(sum(tbl.8.test$hhds))){
  
  print("STOP! Check table!")
  
} else {
  
  print("Proceed.")
  
}

# Write output tables to K drive
write.csv(
  tbl.8,
  paste0(output_path,"/output.tbl_",scen,"_",reb,"_Households.by.NumberOfWorkers.csv"),
  row.names = FALSE
)

gc()
#==============================================================================
# .9 Average Workers in Households with Three or More Workers

if(reb=='MAPC97'){
tbl.9 <- rwt.pums.df.join  |> 
  filter(SPORDER == 1 & WRKHH_Group == "Three or More Workers") |> 
  group_by(
    mpo,
    year
  ) |> 
  summarise(
    avg_workers = round(mean(WRKHH), 2)
  ) |> 
  ungroup()
} else {
  tbl.9 <- rwt.pums.df.join  |> 
  filter(SPORDER == 1 & WRKHH_Group == "Three or More Workers") |> 
  group_by(
    rpa_acr,
    year
  ) |> 
  summarise(
    avg_workers = round(mean(WRKHH), 2)
  ) |> 
  ungroup()
}

# Write output tables to K drive
write.csv(
  tbl.9,
  paste0(output_path,"/output.tbl_",scen,"_",reb,"_Avg.Workers.in.Households.w.ThreeOrMoreWorkers.csv"),
  row.names = FALSE
)
gc()

# Households by RPA/MPO

if(reb=='MAPC97'){
  tbl.10 <- tbl.2 |> 
    group_by(mpo,year) |> 
    summarize(total_households=round(sum(hhds))) |> 
    pivot_wider(id_cols='mpo', names_from = year, names_prefix = 'total_households_', values_from = total_households) |> 
    setDT()
}

if(reb=='MAPC101'){
  tbl.10 <- tbl.2 |> 
    group_by(rpa_acr,year) |> 
    summarize(total_households=round(sum(hhds))) |> 
    pivot_wider(id_cols='rpa_acr', names_from = year, names_prefix = 'total_households_', values_from = total_households) |> 
    setDT()
}

# Write output tables to K drive
write.csv(
  tbl.10,
  paste0(output_path,"/output.tbl_",scen,"_",reb,"_Total.Households.by.",reb,".csv"),
  row.names = FALSE
)


hhpop97 <- fread(paste0(base, "Housing/StatewideHousingPlan/04_Analysis/Data/Working/Regional_Control_Totals/", scen, "/UPDATE/pop.in.hhds.MAPC97_by.RPA.ageCAT6.csv"))

hhpop101 <- fread(paste0(base, "Housing/StatewideHousingPlan/04_Analysis/Data/Working/Regional_Control_Totals/", scen, "/UPDATE/pop.in.hhds.MAPC101_by.RPA.ageCAT6.csv"))


h97 <- hhpop97[,sum(hhpop),.(year,mpo)]
h101 <- hhpop101[,sum(hhpop),.(year,rpa_acr)]
setnames(h97,c('year','mpo','hhpop97'))
setnames(h101,c('year','mpo','hhpop101'))

w97 <- pivot_wider(h97, id_cols = 'mpo', values_from = c('hhpop97'),names_from = 'year', names_prefix = 'hhpop97_')
w101 <- pivot_wider(h101, id_cols = 'mpo', values_from = c('hhpop101'),names_from = 'year', names_prefix = 'hhpop101_')

hhpop <- left_join(w97,w101,by=c('mpo'))




```

# 3.0 Full Crosstab

```{r}
# Creates a table which is a full crosstab of variables from previously generated tables.
# Households is unrounded to avoid rounding error.
if(reb=='MAPC97'){
full.crosstab <- rwt.pums.df.join  |> 
  filter(SPORDER == 1) |>
  group_by(
    mpo,
    year,
    Income_Group,
    Age_Group,
    HHtype_Group,
    CHILDHH_Group,
    Size_Group,
    TEN_Group,
    WRKHH_Group
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup()
} else {
  full.crosstab <- rwt.pums.df.join  |> 
  filter(SPORDER == 1) |>
  group_by(
    rpa_acr,
    year,
    Income_Group,
    Age_Group,
    HHtype_Group,
    CHILDHH_Group,
    Size_Group,
    TEN_Group,
    WRKHH_Group
  ) |> 
  summarise(
    hhds = sum(new_WGTP)
  ) |> 
  ungroup()
}
#Write Output tables to K Drive
write.csv(
  full.crosstab,
  paste0(output_path,"/output.tbl_",scen,"_",reb,"_full.crosstab.fix.csv"),
  row.names = FALSE
)

gc()
```

```{r}
children_inhhd <- rwt.pums.df.join |> 
  filter(children_in_HH >= 1 & SPORDER == 1)
print(round(sum(children_inhhd$new_WGTP)))
```

# 4.0 QA/QC For Total Households and Reweighter Parameters

# 4.1 Load in Target File Data for a Scenario

```{r}
```

# 4.2 Compare Reweighter PUMS data to Reweighter Target Files

-   .1 Total Households by Year
-   .2 Total Households by RPA
-   .3 Total Households by Year and RPA
-   .4 Households by Type
-   .5 Households by Age of Householder

```{r}
#===============================================================================
# -.1 Total Households by Year
# test <- s1.df |> 
#   filter(SPORDER == 1 & TYPEHUGQ == 1) |> 
#   group_by(year) |> 
#   summarise(
#     hhd = sum(new_WGTP)
#   ) |> 
#   ungroup()
# 
# print(sum(test$new_WGTP))
#===============================================================================
# -.2 Total Households by RPA

#===============================================================================
# -.3 Total Households by Year and RPA

#===============================================================================
# -.4 Households by Household Type

#===============================================================================
# -.5 Households by Age of Householder
```

# 5.0 Mutlivariate Output Tables for UrbanSim

-   .8 Household Type by Household Income Groups
-   .9 Household Type by Householder Age Group
-   .10 Household Income Groups by Householder Age Groups
-   .11 Household Type by Household Income Groups by Householder Age Groups

```{r}
```

# 6.0 Multivariate Output Tables for Statewide Housing Plan

```{r}

p101 <- readRDS("S:/Network Shares/K Drive/DataServices/Projects/Current_Projects/Housing/StatewideHousingPlan/04_Analysis/Data/Working/Reweighter/Scenario_3/Output_Files_UPDATE/processor_corrected_PUMS_MAPC101_Scenario_3_2030.rds")

```
