---
title: "UrbanSim Data Processing for Website"
format: html
editor: visual
---

# 1 - Setup

```{r, setup, include=FALSE}

pacman::p_load(
  writexl,
  sf,
  data.table,
  tidycensus,
  mapcdatakeys,
  tidyverse,
  ggplot2,
  imputeTS,
  collapse,
  here
)

#
options(scipen = 999)
#
set.seed(351)
# root <- 'K:/DataServices/Projects/Current_Projects/Projections/Projections_2023/Data/'
root <- 'S:/Network Shares/K Drive/DataServices/Projects/Current_Projects/Projections/Projections_2023/Data/'

inpath <- paste0(root,'03_UrbanSim/UrbanSim_Outputs/Microsim_outputs')

outpath <- paste0(root,'04_Allocation_Results/Revised_Allocation_Sept2025/Tabular/')

sp.path <- paste0(root,'04_Allocation_Results/Revised_Allocation_Sept2025/Spatial/Projections_2023_revised_webmap/') 
  
years <- c(2019,2029,2039,2049)


# # REVISED UrbanSim / TDM23.2.0 data for Projections Website
# MAPCM_run='run_209'
# SWM_run='state_run_113'
# ctps_run='run_113-209'
# years=2019

# # REVISED TDM23.2.0 for Projections Website
# MAPCM_run='run_210'
# SWM_run='state_run_113'
# ctps_run='run_113-210'
# years=c(2029,2039,2049)

sun <- function(x){
  sort(unique(x))}
lun <- function(x){
  length(unique(x))}
sna <- function(x){
  sort(names(x))}

# List of RPA names
mpos <- unlist(c(mapcdatakeys::all_muni_data_keys |> 
                   select(mpo) |> 
                   unique()))

# Munis by RPA/MPO
munis <-
  mapcdatakeys::all_muni_data_keys |> 
  select(muni_id, rpa_acr, mpo) |> 
  setDT()

nonmapc <- munis[mpo!='MAPC']

#Full list of 2010 Blocks
  keys_blocks <- mapcdatakeys::geog_xw_2010 |>
    select(bl10_id, muni_name) |>
    mutate(across(where(is.double), as.character)) |>
    left_join(mapcdatakeys::all_muni_data_keys, by = c("muni_name")) |>
    select(c(bl10_id, muni_name, mpo, rpa_acr))
  
  #Rebel towns block list
  rt_bl10 <- keys_blocks |>
    filter(muni_name %in% c("Duxbury", "Hanover", "Pembroke", "Stoughton")) |>
    pull(bl10_id)
  
  #MAPC 97 town Block List
  mapc97_bl10 <- keys_blocks |>
    filter(mpo == "MAPC") |>
    pull(bl10_id)

knitr::opts_knit$set(root.dir = root)

```

# 2 - Compile Raw Data

```{r}

#2019
fread(paste0(
  inpath,
  '/urbansim_complete_population_run_113-209_2019.csv'
)) |> 
  mutate(year=2019) |> 
  fwrite(paste0(outpath, 'urbansim_complete_microsim_population_website_revised_2019.csv'))

fread(paste0(root, '03_UrbanSim/UrbanSim_Outputs/state_run_113/2019/jobs.csv')) |> fwrite(paste0(outpath,'swm_urbansim_employment_2019.csv'))

fread(paste0(root, '03_UrbanSim/UrbanSim_Outputs/run_209/2019/jobs.csv')) |> fwrite(paste0(outpath,'mapcm_urbansim_employment_2019.csv'))

#2029
fread(paste0(
  inpath,
  '/urbansim_complete_population_run_113-210_2029.csv'
)) |> 
  mutate(year=2029) |> 
  fwrite(paste0(outpath, 'urbansim_complete_microsim_population_website_revised_2029.csv'))

fread(paste0(root, '03_UrbanSim/UrbanSim_Outputs/state_run_113/2029/jobs.csv')) |> fwrite(paste0(outpath,'swm_urbansim_employment_2029.csv'))

fread(paste0(root, '03_UrbanSim/UrbanSim_Outputs/run_210/2029/jobs.csv')) |> fwrite(paste0(outpath,'mapcm_urbansim_employment_2029.csv'))

# 2039
fread(paste0(
  inpath,
  '/urbansim_complete_population_run_113-210_2039.csv'
)) |> 
  mutate(year=2039) |> 
  fwrite(paste0(outpath, 'urbansim_complete_microsim_population_website_revised_2039.csv'))

fread(paste0(root, '03_UrbanSim/UrbanSim_Outputs/state_run_113/2039/jobs.csv')) |> fwrite(paste0(outpath,'swm_urbansim_employment_2039.csv'))

fread(paste0(root, '03_UrbanSim/UrbanSim_Outputs/run_210/2039/jobs.csv')) |> fwrite(paste0(outpath,'mapcm_urbansim_employment_2039.csv'))

# 2049
fread(paste0(
  inpath,
  '/urbansim_complete_population_run_113-210_2049.csv'
)) |> 
  mutate(year=2049) |> 
  fwrite(paste0(outpath, 'urbansim_complete_microsim_population_website_revised_2049.csv'))

fread(paste0(root, '03_UrbanSim/UrbanSim_Outputs/state_run_113/2049/jobs.csv')) |> fwrite(paste0(outpath,'swm_urbansim_employment_2049.csv'))

fread(paste0(root, '03_UrbanSim/UrbanSim_Outputs/run_210/2049/jobs.csv')) |> fwrite(paste0(outpath,'mapcm_urbansim_employment_2049.csv'))



```

# 3 - Table Export for Website

```{r}

 
for (yr in years){
  
new <- fread(paste0(
  outpath,
  'urbansim_complete_microsim_population_website_revised_',
  yr,
  '.csv'
))

semp <- fread(paste0(outpath,'swm_urbansim_employment_',yr,'.csv'))
memp <- fread(paste0(outpath,'mapcm_urbansim_employment_',yr,'.csv'))

# Population in Households

hhpop <- new |>
  mutate(pop = 1) |>
  group_by(rpa_acr, mpo, muni_name, muni_id) |>
  summarise(HHPop = sum(pop)) |>
  setnames(c('RPA', 'MPO', 'Muni', 'Muni_ID', paste0('HHPop_', yr + 1)))

fwrite(
  hhpop,
  paste0(
    outpath,
    'muni_HHpop_',
    yr + 1,
    '.csv'
  )
)

# Population in Households by Age

age.grp <- data.table(
  ageCAT6 = seq(1, 18),
  age_group = c(
    '0-04',
    '05-09',
    '10-14',
    '15-19',
    '20-24',
    '25-29',
    '30-34',
    '35-39',
    '40-44',
    '45-49',
    '50-54',
    '55-59',
    '60-64',
    '65-69',
    '70-74',
    '75-79',
    '80-84',
    '85p'
  )
)

hhpop.age <- new |>
  mutate(pop = 1) |>
  group_by(rpa_acr, mpo, muni_name, muni_id, ageCAT6) |>
  summarise(HHPop = sum(pop)) |>
  ungroup() |>
  left_join(age.grp, by = 'ageCAT6') |>
  select(-ageCAT6) |>
  relocate(age_group, .before = HHPop) |>
  setorder(muni_id, age_group) |>
  setnames(c(
    'RPA',
    'MPO',
    'Muni',
    'Muni_ID',
    'Household Population Age Group',
    paste0('HHPop_', yr + 1)
  ))

fwrite(
  hhpop.age,
  paste0(
    outpath,
    'muni_HHpop_age_',
    yr + 1,
    '.csv'
  )
)

# Population in Households under the age of 18

u18 <- new |>
  filter(AGEP < 18) |>
  mutate(pop = 1) |>
  group_by(rpa_acr, mpo, muni_name, muni_id) |>
  summarise(HHPop = sum(pop)) |>
  setnames(c('RPA', 'MPO', 'Muni', 'Muni_ID', paste0('HHPop_', yr + 1)))

fwrite(
  u18,
  paste0(
    outpath,
    'muni_HHpop_u18_',
    yr + 1,
    '.csv'
  )
)

# Households by household type

hht <- data.table(
  HHT = c(1, 3, 5),
  type = c('HH with Children', '2+ Adults with No Children', 'Living Alone')
)

hh.type <- new |>
  filter(sporder == 1) |>
  mutate(hh = 1) |>
  group_by(rpa_acr, mpo, muni_name, muni_id, HHT) |>
  summarise(Households = sum(hh)) |>
  left_join(hht, by = 'HHT') |>
  setorder(muni_id, HHT) |>
  relocate(type, .before = Households) |>
  setnames(c(
    'RPA',
    'MPO',
    'Muni',
    'Muni_ID',
    'HHT',
    'Household Type',
    paste0('Households_', yr + 1)
  ))

hhtot <- new |>
  filter(sporder == 1) |>
  mutate(hh = 1) |>
  group_by(rpa_acr, mpo, muni_name, muni_id) |>
  summarise(Households = sum(hh)) |>
  mutate(HHT = 6, type = 'Total Households') |>
  relocate(Households, .after = type) |>
  setnames(c(
    'RPA',
    'MPO',
    'Muni',
    'Muni_ID',
    'HHT',
    'Household Type',
    paste0('Households_', yr + 1)
  )) |>
  ungroup()

hh.type.exp <- rbind(hh.type, hhtot) |>
  setorder(Muni_ID, HHT)

fwrite(
  hh.type.exp,
  paste0(
    outpath,
    'muni_HHType_',
    yr + 1,
    '.csv'
  )
)

# Households by age of householder

hhd.age <- new |>
  filter(sporder == 1) |>
  mutate(
    hh = 1,
    age.grp = case_when(
      ageHHder %in% 15:34 ~ '15-34',
      ageHHder %in% 35:44 ~ '35-44',
      ageHHder %in% 45:64 ~ '45-64',
      ageHHder %in% 65:74 ~ '65-74',
      ageHHder >= 75 ~ '75p'
    )
  ) |>
  group_by(rpa_acr, mpo, muni_name, muni_id, age.grp) |>
  summarise(Households = sum(hh)) |>
  setnames(c(
    'RPA',
    'MPO',
    'Muni',
    'Muni_ID',
    'Householder Age Group',
    paste0('Households_', yr + 1)
  )) |>
  ungroup()

hhdtot <- hhtot |>
  select(-HHT) |>
  rename('Householder Age Group' = 'Household Type')

hhd.age.exp <- rbind(hhd.age, hhdtot) |>
  setorder(Muni_ID, `Householder Age Group`)

fwrite(
  hhd.age.exp,
  paste0(
    outpath,
    'muni_HHDer_age_',
    yr + 1,
    '.csv'
  )
)

# Households my household income

hinc <- new |>
  filter(sporder == 1) |>
  mutate(hh = 1) |>
  group_by(rpa_acr, mpo, muni_name, muni_id, income_grp) |>
  summarise(Households = sum(hh)) |>
  setnames(c(
    'RPA',
    'MPO',
    'Muni',
    'Muni_ID',
    'Household Income Group',
    paste0('Households_', yr + 1)
  )) |>
  ungroup()

fwrite(
  hinc,
  paste0(
    outpath,
    'muni_HH_income_',
    yr + 1,
    '.csv'
  )
)

# Employment by super sector

semp[, jobs := .N, .(block_id, aggr_sector_id)]
semp <- unique(semp)
memp[, jobs := .N, .(block_id, aggr_sector_id)]
memp <- unique(memp)
semp <- semp[!block_id %in% mapc97_bl10]
memp <- memp[!block_id %in% rt_bl10]
lemp <- rbind(semp, memp)
setorder(lemp, block_id, aggr_sector_id)
lemp[, bl10_id := as.character(block_id)]
wemp <- lemp |> left_join(keys_blocks, by = 'bl10_id') |>
  select(-block_id) |> 
  filter(!is.na(muni_name))

sects <- data.table(
  aggr_sector_id = seq(1, 10),
  SuperSector = c(
    'Construction',
    'Education and Health Services',
    'Financial Activities',
    'Government',
    'Information',
    'Leisure and Hospitality',
    'Manufacturing',
    'Mining and Logging / Other Services',
    'Professional and Business Services',
    'Trade, Transportation, Utilities'
  )
)

muni.emp <- wemp |>
  group_by(muni_name, mpo, rpa_acr, aggr_sector_id) |>
  summarize(jobs = sum(jobs)) |>
  mutate(year = yr + 1) |>
  left_join(sects, by = 'aggr_sector_id')

fwrite(
  muni.emp,
  paste0(
    outpath,
    'muni_employment_',
    yr + 1,
    '.csv'
  )
)

}

# Compile all years and export as EXCEL
  
f1 <- list.files(path = outpath, pattern = 'muni_HHpop_2', full.names = T)

f2 <- list.files(path = outpath, pattern = 'muni_HHpop_age', full.names = T)
                 
f3 <- list.files(path = outpath, pattern = 'muni_HHpop_u18', full.names = T)

f4 <- list.files(path = outpath, pattern = 'muni_HHType', full.names = T) 

f5 <- list.files(path = outpath, pattern = 'muni_HHDer', full.names = T)                 
f6 <- list.files(path = outpath, pattern = 'muni_HH_income', full.names = T)

f7 <- list.files(path = outpath, pattern = 'muni_employment', full.names = T)


l1 <- lapply(f1, fread)
d1 <- l1[[1]] |> 
  left_join(l1[[2]], by=c('RPA','MPO','Muni','Muni_ID')) |>
  left_join(l1[[3]], by=c('RPA','MPO','Muni','Muni_ID')) |> 
  left_join(l1[[4]], by=c('RPA','MPO','Muni','Muni_ID'))

l2 <- lapply(f2, fread)
d2 <- l2[[1]] |> 
  left_join(l2[[2]], by=c('RPA','MPO','Muni','Muni_ID','Household Population Age Group')) |>
  left_join(l2[[3]], by=c('RPA','MPO','Muni','Muni_ID','Household Population Age Group')) |> 
  left_join(l2[[4]], by=c('RPA','MPO','Muni','Muni_ID','Household Population Age Group'))

l3 <- lapply(f3, fread)
d3 <- l3[[1]] |> 
  left_join(l3[[2]], by=c('RPA','MPO','Muni','Muni_ID')) |>
  left_join(l3[[3]], by=c('RPA','MPO','Muni','Muni_ID')) |> 
  left_join(l3[[4]], by=c('RPA','MPO','Muni','Muni_ID'))

l4 <- lapply(f4, fread)
d4 <- l4[[1]] |> 
  left_join(l4[[2]], by=c('RPA','MPO','Muni','Muni_ID','HHT','Household Type')) |>
  left_join(l4[[3]], by=c('RPA','MPO','Muni','Muni_ID','HHT','Household Type')) |>
  left_join(l4[[4]], by=c('RPA','MPO','Muni','Muni_ID','HHT','Household Type')) |> 
  setorder(Muni_ID,HHT) |> 
  select(-HHT)

l5 <- lapply(f5, fread)
d5 <- l5[[1]] |> 
  left_join(l5[[2]], by=c('RPA','MPO','Muni','Muni_ID','Householder Age Group')) |>
  left_join(l5[[3]], by=c('RPA','MPO','Muni','Muni_ID','Householder Age Group')) |>
  left_join(l5[[4]], by=c('RPA','MPO','Muni','Muni_ID','Householder Age Group'))

l6 <- lapply(f6, fread)
d6 <- l6[[1]] |> 
  left_join(l6[[2]], by=c('RPA','MPO','Muni','Muni_ID','Household Income Group')) |>
  left_join(l6[[3]], by=c('RPA','MPO','Muni','Muni_ID','Household Income Group')) |>
  left_join(l6[[4]], by=c('RPA','MPO','Muni','Muni_ID','Household Income Group')) |> 
  mutate(inc = case_when(
    `Household Income Group` == '<=35000' ~ 1,
    `Household Income Group` == '35000-75000' ~ 2,
    `Household Income Group` == '75000-125000' ~ 3,
    `Household Income Group` == '125000-225000' ~ 4,
    `Household Income Group` == '>225000' ~ 5
  )) |> 
  setorder(Muni_ID,inc) |> 
  select(-inc)


l7 <- lapply(f7, fread)
d7 <- rbindlist(l7)
setnames(d7,'aggr_sector_id','SuperSector_ID')

dw7 <- pivot_wider(
  d7,
  id_cols = c('rpa_acr', 'mpo', 'muni_name', 'SuperSector_ID', 'SuperSector'),
  names_prefix = 'Employment_',
  names_from = 'year',
  values_from = 'jobs'
)

emptot <- d7 |>
  group_by(rpa_acr, mpo, muni_name, year) |>
  summarize(jobs = sum(jobs)) |>
  mutate(SuperSector_ID = 100, SuperSector = 'Total Employment') |>
  pivot_wider(
    id_cols = c('rpa_acr', 'mpo', 'muni_name', 'SuperSector_ID', 'SuperSector'),
    names_prefix = 'Employment_',
    names_from = 'year',
    values_from = 'jobs'
  ) |> rbind(dw7) |>
  setorder(muni_name, SuperSector_ID)

xl.list <- list(
  'Household Population' = d1,
  'Household Population by Age' = d2,
  'Household Population Under 18' = d3,
  'Households by Type' = d4,
  'Households by Age of HHder' = d5,
  'Households by HH Income Group' = d6,
  'Employment by Super Sector' = emptot
)

write_xlsx(xl.list, paste0(outpath,'LRTP2023_Municipal_Data_revised_20250929.xlsx'))

```

# 4 - Shapefile Export for Website

```{r}
# Export shapefiles for web map

sfm <- muni_sf(2010)

sf.hhpop <- sfm |>
  left_join(d1, by = c('muni_id' = 'Muni_ID')) |>
  st_write(paste0(sp.path, 'muni_HHpop.shp'), append = F)

hhtot <- d4 |> filter(`Household Type` == 'Total Households') |> select(-`Household Type`) |> rename(
  HH2020 = Households_2020,
  HH2030 = Households_2030,
  HH2040 = Households_2040,
  HH2050 = Households_2050
)
sf.hhds <- sfm |>
  left_join(hhtot, by = c('muni_id' = 'Muni_ID')) |>
  st_write(paste0(sp.path, 'muni_HHds.shp'), append = F)

et <- emptot |> filter(SuperSector_ID == 100) |> select(-SuperSector_ID, -SuperSector) |>
  rename(
    RPA = rpa_acr,
    MPO = mpo,
    Muni = muni_name,
    jobs2020 = Employment_2020,
    jobs2030 = Employment_2030,
    jobs2040 = Employment_2040,
    jobs2050 = Employment_2050
  )
sf.hhds <- sfm |>
  left_join(et, by = c('muni_name' = 'Muni')) |>
  st_write(paste0(sp.path, 'muni_employment.shp'), append = F)

# Massbuilds shapefiles

vintage <- '20250930'

mb.in.lrtp <- fread(paste0(sp.path, 'all_devprojects_9_30_2025_SWM.csv')) |>
  select(identification) |> 
  unique() |> 
  pull()

res <- fread(
  paste0(
    root,
    '03_UrbanSim/MassBuilds/Massbuilds_inputs_',
    vintage,
    '/Statewide/residential.csv'
  )
)
com <- fread(
  paste0(
    root,
    '03_UrbanSim/MassBuilds/Massbuilds_inputs_',
    vintage,
    '/Statewide/non_residential.csv'
  )
)
mb <- rbind(res, com, fill = T) |>
  filter(identification %in% mb.in.lrtp) |>
  select(
    identification,
    x,
    y,
    name,
    start_year,
    status,
    street,
    city,
    state,
    zip,
    residential_units,
    employment_capacity
  ) |>
  group_by(identification, x, y, name, status, street, city, state, zip) |>
  summarize(
    res_units = sum(residential_units, na.rm = T),
    emp_cap = sum(employment_capacity, na.rm = T),
    year = max(start_year)
  ) |>
  setDT()

mb[is.na(res_units), res_units := 0]
mb[is.na(emp_cap), emp_cap := 0]

dupe <- mb[duplicated(identification), unique(identification)]

mb1 <- mb[!identification %in% dupe]
mb2 <- mb[identification %in% dupe & res_units == 0] |>
  select(identification, emp_cap)
mb3 <- mb[identification %in% dupe & emp_cap == 0] |>
  select(-emp_cap) |>
  left_join(mb2, by = 'identification') |>
  relocate(year, .before = status) |>
  setDT()
mb3[status == 'Proposed', status := 'Planned']

mbsf <- st_as_sf(mb3, coords = c('x', 'y'))
st_crs(mbsf) <- 4326
mexp <- st_transform(mbsf, 26986)
st_write(
  mexp,
  paste0(
    sp.path,'massbuilds_projects_lrtp_revised.shp'
  ),
  append = F
)

```

# 5.0 - Plots and Charts
# 5.1 Load 2010 Census Data
```{r}
#Municipal GEOIDs for 2010 Decennial Census
GEOID.10 <- mapcdatakeys::all_muni_data_keys |>  
  select(muni_id, cosub_cn10) |> 
  mutate(
    GEOID = as.character(cosub_cn10)
  ) |> 
  select(GEOID, muni_id)

# 
vars.dec10 <- load_variables(2010, dataset = c("sf1"))

#Decennial 2010 Population
#Creates a list of sex by age population variables to use in the API query
pop.vars <- dec.vars(vars.dec10, "SEX BY AGE")
pop.vars <- data.frame(pop.vars) |> dplyr::slice(1:49) |> pull(pop.vars)

#Assigns appropriate labels to the Decennial Census variables
pop.labels <- dec.labels("2010", "SEX BY AGE") |> slice(1:46)

#Queries the Census API via {tidycensus} to retrieve data about the population
pop.values <- decennial_call(pop.vars, geog = "county subdivision", 2010)

#Clean the Population data.
pop.10 <- pop.values |> 
  filter(GEOID %in% GEOID.10$GEOID) |> 
  right_join(
    pop.labels,
    by = c("variable"="name")
  ) |> 
  #Change the Census age category levels to match the ageCAT6 labels
  mutate(
    ageCAT6 = case_when(
      label == 'pop_Under_5_years' ~ '1',
      label == "pop_5_to_9_years" ~ '2',
      label == "pop_10_to_14_years" ~ '3',
      label %in% c("pop_15_to_17_years","pop_18_and_19_years") ~ '4',
      label %in% c("pop_20_years","pop_21_years","pop_22_to_24_years") ~ '5',
      label == "pop_25_to_29_years" ~ '6',
      label == "pop_30_to_34_years" ~ '7',
      label == "pop_35_to_39_years" ~ '8',
      label == "pop_40_to_44_years" ~ '9',
      label == "pop_45_to_49_years" ~ '10',
      label == "pop_50_to_54_years" ~ '11',
      label == "pop_55_to_59_years" ~ '12',
      label %in% c("pop_60_and_61_years","pop_62_to_64_years") ~ '13',
      label %in% c("pop_65_and_66_years","pop_67_to_69_years") ~ '14',
      label == "pop_70_to_74_years" ~ '15',
      label == "pop_75_to_79_years" ~ '16',
      label == "pop_80_to_84_years" ~ '17',
      label == "pop_85_years_and_over" ~ '18'
    )
  ) |> 
  filter(!is.na(ageCAT6)) |>  
  right_join(
    GEOID.10,
    by = c('GEOID')
  ) |>  
  mutate(
    muni_id = as.character(muni_id)
  ) |> 
  dplyr::rename(
    Population = value,
    Sex = sex
  ) |>  
  select(muni_id, Sex, ageCAT6, Population) |>  
  group_by(
    muni_id,
    Sex,
    ageCAT6
  ) |> 
  summarise(
    Population = sum(Population)
  ) |>  
  ungroup()

#Remove dataframes no longer needed in the process.
rm(pop.labels, pop.values, pop.vars)

#===============================================================================
#Decennial 2010 Population in Households
#Creates a list of sex by age population variables to use in the API query
pa.vars <- dec.vars(vars.dec10, "SEX BY AGE FOR THE POPULATION IN HOUSEHOLDS")

#Assigns appropriate labels to the Decennial Census variables
pa.labels <- dec.labels("2010", "SEX BY AGE FOR THE POPULATION IN HOUSEHOLDS")

#Queries the Census API via {tidycensus} to retrieve data about the population
#in households
pa.values <- decennial_call(pa.vars, geog = "county subdivision", 2010)

#Clean the Household Population data.
hhpop.10 <- pa.values |> 
  filter(GEOID %in% GEOID.10$GEOID) |> 
  right_join(
    pa.labels,
    by = c("variable"="name")
  ) |> 
  #Change the Census age category levels to match the ageCAT6 labels
  mutate(
    ageCAT6 = case_when(
      label == 'pop_Under_5_years' ~ '1',
      label == "pop_5_to_9_years" ~ '2',
      label == "pop_10_to_14_years" ~ '3',
      label %in% c("pop_15_to_17_years","pop_18_and_19_years") ~ '4',
      label %in% c("pop_20_years","pop_21_years","pop_22_to_24_years") ~ '5',
      label == "pop_25_to_29_years" ~ '6',
      label == "pop_30_to_34_years" ~ '7',
      label == "pop_35_to_39_years" ~ '8',
      label == "pop_40_to_44_years" ~ '9',
      label == "pop_45_to_49_years" ~ '10',
      label == "pop_50_to_54_years" ~ '11',
      label == "pop_55_to_59_years" ~ '12',
      label %in% c("pop_60_and_61_years","pop_62_to_64_years") ~ '13',
      label %in% c("pop_65_and_66_years","pop_67_to_69_years") ~ '14',
      label == "pop_70_to_74_years" ~ '15',
      label == "pop_75_to_79_years" ~ '16',
      label == "pop_80_to_84_years" ~ '17',
      label == "pop_85_years_and_over" ~ '18'
    )
  ) |> 
  filter(!is.na(ageCAT6)) |>  
  right_join(
    GEOID.10,
    by = c('GEOID')
  ) |>  
  mutate(
    muni_id = as.character(muni_id)
  ) |> 
  dplyr::rename(
    HHPopulation = value,
    Sex = sex
  ) |>  
  select(muni_id, Sex, ageCAT6, HHPopulation) |>  
  group_by(
    muni_id,
    Sex,
    ageCAT6
  ) |> 
  summarise(
    HHPopulation = sum(HHPopulation)
  ) |>  
  ungroup()

#Remove dataframes no longer needed in the process.
rm(pa.labels, pa.values, pa.vars)
```
# 5.2 Load 
```{r}
# Functions created for the Urbansim/Projections workflow
source(here("urbansim_model_update/projections_functions.R"))

# Set the path to the data to load in the for loop below.
input.path <- "K:/DataServices/Projects/Current_Projects/Projections/Projections_2023/Data/04_Allocation_Results/Revised_Allocation_Sept2025/Tabular/"
output_path <- "K:/DataServices/Projects/Current_Projects/Projections/Projections_2023/Final_Release/Visualizations"

# Load data from workbook posted to Projections webiste.
# Household Population
pop <- readxl::read_xlsx(paste0(input.path, "LRTP2023_Municipal_Data_revised_20250929.xlsx"), sheet = 2)
pop.age <- readxl::read_xlsx(paste0(input.path, "LRTP2023_Municipal_Data_revised_20250929.xlsx"), sheet = 3)
pop.child <- readxl::read_xlsx(paste0(input.path, "LRTP2023_Municipal_Data_revised_20250929.xlsx"), sheet = 4)

# Households  
hhds.type <- readxl::read_xlsx(paste0(input.path, "LRTP2023_Municipal_Data_revised_20250929.xlsx"), sheet = 5)
hhds.age <- readxl::read_xlsx(paste0(input.path, "LRTP2023_Municipal_Data_revised_20250929.xlsx"), sheet = 6)
hhds.income <- readxl::read_xlsx(paste0(input.path, "LRTP2023_Municipal_Data_revised_20250929.xlsx"), sheet = 7)

# Employment
emp <- readxl::read_xlsx(paste0(input.path, "LRTP2023_Municipal_Data_revised_20250929.xlsx"), sheet = 8)

```
# 5.3 - Level Demographic Components of Projected Households (Stacked Bar Charts)
```{r}
# Overall
# Household Type

# Household Size

# Householder Age

# RPA
# Household Type

# Household Size

# Householder Age

```

# 5.4 Level Components of Projected Employment (Stacked Bar Charts)
```{r}
# Overall
# Supersectors

# RPA
# Supersectors

```

# 5.5 Level Projected Population (Stacked Bar Charts)
```{r}
# 5.5.1 Overall Population
# State
pop.bc.1 <- pop |> 
  mutate(state = 1) |> 
  group_by(state) |> 
  summarise(
    HHPop_2020 = sum(HHPop_2020),
    HHPop_2030 = sum(HHPop_2030),
    HHPop_2040 = sum(HHPop_2040),
    HHPop_2050 = sum(HHPop_2050),
  ) |>
  ungroup() |> 
  dplyr::rename(
    `2020` = HHPop_2020,
    `2030` = HHPop_2030,
    `2040` = HHPop_2040,
    `2050` = HHPop_2050,
  ) |> 
  pivot_longer(
    !state,
    names_to = "year",
    values_to = "hh.population"
  ) |> 
  select(-c(state))


graph <-
  sv.bar.chart_(
    df = "pop.bc.1",
    x_var = "year",
    y_var = "hh.population",
    x_label = "Year",
    y_label = "Population",
    geo = "Massachusetts"
  )

graph

ggsave(
  filename = paste0("pop_overall.jpeg"),
  plot = get_last_plot(),
  path = output_path,
  dpi = 320,
  bg = NULL
)

# RPA
pop.bc.2 <- pop |>
  group_by(RPA) |> 
  summarise(
    HHPop_2020 = sum(HHPop_2020),
    HHPop_2030 = sum(HHPop_2030),
    HHPop_2040 = sum(HHPop_2040),
    HHPop_2050 = sum(HHPop_2050),
  ) |>
  ungroup() |> 
  pivot_longer(
    !RPA,
    names_to = "year",
    values_to = "hh.population"
  ) 

# Create a plot of total population for each RPA.
for (r in RPA) {
  # Create graph object for each RPA
  tmp <- pop.bc.2 |>
    filter(RPA == r) |>
    ggplot(aes(x = year, y = hh.population)) +
    geom_col()
  
  # Display graph object in the terminal.
  tmp
  
  # Save graph to disk.
  ggsave(
    filename = paste0("pop_",r,".jpeg"),
    plot = get_last_plot(),
    path = output_path,
    dpi = 320,
    bg = NULL
  )
  
  # Remove tmp graph object
  rm(tmp)
  
  # Print statement
  print(paste0("Population bar chart for ", r, " saved on disk!"))
  
}

# 5.5.2 Population by Age
# State

# RPA

# 5.5.3 Population under 18
# State

# RPA

```

# 5.6 Percent Change Demographic Components of Projected Households (Stacked Bar Charts)
```{r}
# Overall
# Household Type

# Household Size

# Householder Age

# RPA
# Household Type

# Household Size

# Householder Age

```

# 5.7 Percent Change Components of Projected Employment (Stacked Bar Charts)
```{r}
# Overall
# Supersectors

# RPA
# Supersectors

```

# 5.8 Percent Change Projected Population (Stacked Bar Charts)
```{r}
# Overall
# Age 

# RPA
# Age

```
