---
title: "Collection of all old projections code"
format: html
editor: visual
---

# 3.6.1 - Custom exports - PCAP

```{r}
#| eval: false


out.dir <- 'S:/Network Shares/K Drive/DataServices/Projects/Current_Projects/Projections/Projections_2023/Associated_Projects/PCAP/'

mlist <- list()
#for (yr in c(2039)) {
for (yr in c(2029,2039,2049)){
  myear = yr
  syear = yr
  
  hh <- readRDS(
    paste0(
      root,
      SWM_run,
      '/urbansim_',
      SWM_run,
      '_microhouseholds_',
      syear,
      '_pop_corrected_v3.rds'
    )
  )
  mh <- readRDS(
    paste0(
      root,
      MAPCM_run,
      '/urbansim_',
      MAPCM_run,
      '_microhouseholds_',
      myear,
      '_pop_corrected_v3.rds'
    )
  )
  
  munis <- mapcdatakeys::all_muni_data_keys |> select(muni_id, muni_name, mpo)
  
  # Keeps rebel towns in MAPC region
  #hp <- hh[!block_id %in% mp$block_id]
  
  #Full list of 2010 Blocks
  keys_blocks <- mapcdatakeys::geog_xw_2010 |>
    select(bl10_id, muni_name) |>
    mutate(across(where(is.double), as.character)) |> left_join(mapcdatakeys::all_muni_data_keys, by = c("muni_name")) |>
    select(c(bl10_id, muni_name, mpo, rpa_acr))
  
  #Rebel towns block list
  rt_bl10 <- keys_blocks |>
    filter(muni_name %in% c("Duxbury", "Hanover", "Pembroke", "Stoughton")) |>
    pull(bl10_id)
  
  #MAPC 97 town Block List
  mapc97_bl10 <- keys_blocks |>
    filter(mpo == "MAPC") |>
    pull(bl10_id)
  
  #Subset MAPC to 97 towns
  mp <- mh |>
    mutate(bl10_id = as.character(bl10_id)) |>
    dplyr::filter(bl10_id %in% mapc97_bl10)
  
  #Removes MAPC97 from statewide data
  hp <- hh |>
    mutate(bl10_id = as.character(bl10_id)) |> 
    filter(!bl10_id %in% mapc97_bl10)
  
  new <- rbind(mp, hp)
  new <- new[block_id != -1]
  setnames(new, 'inc2010', 'wage_inc')
  setnames(new, 'hhinc2010', 'hh_inc')
  
  blk.hh <- new[, lun(hid), block_id]
  setnames(blk.hh, 'V1', 'total_households')
  setkey(blk.hh, block_id)
  
  new[person_num==1, hh:=1]
  new[,pop:=1]
  new[,state:='MA']
  muni.sum <- new[,lapply(.SD,sum,na.rm=T),.(muni_name,state),.SDcols=c('pop','hh','child','adult','is_worker')]
  ns <- c('pop','hh','children','adults','workers')
  setnames(muni.sum,c('muni_name','state',paste0(ns,'_',yr+1)))
  
  mlist[[yr]] <- muni.sum
}

pop.muni <- mlist[[2029]] |> 
  left_join(mlist[[2039]], by=c('muni_name','state')) |> 
  left_join(mlist[[2049]], by=c('muni_name','state'))
rm(mlist);gc()
  
# NH + RI

taz <- fread(paste0(out.dir, 'taz_muni_join.csv')) |>
  filter(EXTERNAL == 0) |>
  select(tazID, muni_name, state)

nh20 <- fread(
  'S:/Network Shares/K Drive/DataServices/Datasets/Transportation/CTPS/TDM23/tdm23.2.0/Inputs/inputs/zonal/nhri_population_2020_v20230518.csv'
) |>
  left_join(taz, by = c('block_id' = 'tazID')) |>
  mutate(
    pop = 1,
    hh = case_when(person_num == 1 ~ 1, .default = 0),
    year = 2020,
    child = case_when(age < 18 ~ 1, .default = 0),
    adult = case_when(age >= 18 ~ 1, .default = 0)
  ) |>
  group_by(year, muni_name, state) |>
  summarize(
    pop = sum(pop),
    hh = sum(hh),
    children = sum(child),
    adults = sum(adult),
    workers = sum(is_worker)
  ) |>
  ungroup() |>
  pivot_wider(names_from = year,
              values_from = c(pop, hh, adults, children, workers))

nh50 <- fread(
  'S:/Network Shares/K Drive/DataServices/Datasets/Transportation/CTPS/TDM23/tdm23.2.0/Inputs/inputs/zonal/nhri_population_2050_v20230518.csv'
) |>
  left_join(taz, by = c('block_id' = 'tazID')) |>
  mutate(
    pop = 1,
    hh = case_when(person_num == 1 ~ 1, .default = 0),
    year = 2050,
    child = case_when(age < 18 ~ 1, .default = 0),
    adult = case_when(age >= 18 ~ 1, .default = 0)
  ) |>
  group_by(year, muni_name, state) |>
  summarize(
    pop = sum(pop),
    hh = sum(hh),
    children = sum(child),
    adults = sum(adult),
    workers = sum(is_worker)
  ) |>
  ungroup() |>
  pivot_wider(names_from = year,
              values_from = c(pop, hh, adults, children, workers)) |>
  left_join(nh20, by = c('muni_name', 'state')) |>
  mutate(
    pop_2030 = round((pop_2050 - pop_2020) / 3 + pop_2020),
    pop_2040 = round(2 * ((pop_2050 - pop_2020) / 3) + pop_2020),
    hh_2030 = round((hh_2050 - hh_2020) / 3 + hh_2020),
    hh_2040 = round(2 * ((hh_2050 - hh_2020) / 3) + hh_2020),
    children_2030 = round((children_2050 - children_2020) / 3 + children_2020),
    children_2040 = round(2 * ((
      children_2050 - children_2020
    ) / 3) + children_2020),
    adults_2030 = round((adults_2050 - adults_2020) / 3 + adults_2020),
    adults_2040 = round(2 * ((
      adults_2050 - adults_2020
    ) / 3) + adults_2020),
    workers_2030 = round((workers_2050 - workers_2020) / 3 + workers_2020),
    workers_2040 = round(2 * ((
      workers_2050 - workers_2020
    ) / 3) + workers_2020)
  ) |> 
  select(-pop_2020,-hh_2020,-adults_2020,-children_2020,-workers_2020)

pop_exp <- rbind(pop.muni,nh50)
pop_exp[,key:=paste(muni_name,state)]

msa <- fread(paste0(out.dir,'msa_munis.csv'))
msa[,key:=paste(muni_name,state)]

pop_exp_df <- pop_exp[key %in% msa$key]  

stamp <- gsub('-', '', lubridate::today())
fwrite(
    pop_exp_df,
    paste0(
    out.dir,
      '/msa_population_',
      ctps_run,
      '_2030_2040_2050',
      '_v',
      stamp,
      '.csv'
    )
  )

  semp <- fread(paste0(root, SWM_run, '/', syear, '/jobs.csv'))
  sew <- dcast.data.table(semp,
                          block_id ~ aggr_sector_id,
                          value.var = 'aggr_sector_id',
                          fun.aggregate = length)
  setnames(sew, c('block_id', paste0('emp_naics_', 1:10)))
  sew[, total_emp := emp_naics_1 + emp_naics_2 + emp_naics_3 + emp_naics_4 +
        emp_naics_5 + emp_naics_6 + emp_naics_7 + emp_naics_8 + emp_naics_9 + emp_naics_10]
  
  memp <- fread(paste0(root, MAPCM_run, '/', myear, '/jobs.csv'))
  mew <- dcast.data.table(memp,
                          block_id ~ aggr_sector_id,
                          value.var = 'aggr_sector_id',
                          fun.aggregate = length)
  setnames(mew, c('block_id', paste0('emp_naics_', 1:10)))
  mew[, total_emp := emp_naics_1 + emp_naics_2 + emp_naics_3 + emp_naics_4 +
        emp_naics_5 + emp_naics_6 + emp_naics_7 + emp_naics_8 + emp_naics_9 + emp_naics_10]
  
  sew <- sew |>
    mutate(bl10_id = as.character(block_id)) |>
    left_join(keys_blocks, by = "bl10_id")
  
  mew <- mew |>
    mutate(bl10_id = as.character(block_id)) |>
    left_join(keys_blocks, by = "bl10_id")
  
  sew <- sew[!bl10_id %in% mapc97_bl10]
  mew <- mew[!bl10_id %in% rt_bl10]
  sew <- rbind(sew, mew)
  setnames(sew, 'total_emp', 'total_jobs')
  setkey(sew, block_id)
  setkey(blk.hh, block_id)
  sew.exp <- merge(sew, blk.hh, all = T, by = 'block_id') |>
    select(
      emp_naics_1,
      emp_naics_2,
      emp_naics_3,
      emp_naics_4,
      emp_naics_5,
      emp_naics_6,
      emp_naics_7,
      emp_naics_8,
      emp_naics_9,
      emp_naics_10,
      total_jobs,
      total_households,
      block_id
    )
  sew.exp[is.na(sew.exp)] <- 0
  
  setnames(sew.exp, (
    c(
      '1_constr',
      '2_eduhlth',
      '3_finance',
      '4_public',
      '5_info',
      '6_ret_leis',
      '7_manu',
      '8_other',
      '9_profbus',
      '10_ttu',
      'total_jobs',
      'total_households',
      'block_id'
    )
  ))
  
  fwrite(sew.exp,
         paste0(
           root,
           dir,
           '/ma_employment_',
           ctps_run,
           '_',
           yr,
           '_v',
           stamp,
           '_v2.csv'
         ))
  
  
  
  # blk.pop <- new[, .N, block_id]
  # setnames(blk.pop, 'N', paste0('pop_', syear + 1))
  # setkey(blk.pop, block_id)
  # blk.hh <- new[sporder == 1, .N, block_id]
  # setnames(blk.hh, 'N', paste0('households_', syear + 1))
  # setkey(blk.hh, block_id)
  # blk.sum <- blk.hh[blk.pop]
  # rm(hh, mp, hp)
  # gc()
  # 
# semp[,jobs:=.N,.(block_id,aggr_sector_id)]
# semp <- unique(semp)
# memp[,jobs:=.N,.(block_id,aggr_sector_id)]
# memp <- unique(memp)
# semp <- semp[!block_id %in% mapc97_bl10]
# memp <- memp[!block_id %in% rt_bl10]
# lemp <- rbind(semp,memp)
# setorder(lemp,block_id,aggr_sector_id)
# fwrite(lemp,paste0(root,dir,'/urbansim_',SWM_run,'_',MAPCM_run,'_emp_by_supernaics_',syear,'_long.csv'))
}


```

# 3.6.2 - Custom exports - Massbuilds LRTP

```{r}
dir <- 'Microsim_outputs'

# Household trends
hh19 <- fread(paste0(root,dir,'/urbansim_',SWM_run,'_2019_mapc_',MAPCM_run,'_2019_combined_microsim_households.csv'))
hh19 <- hh19[,lun(hid),.(bl10_id)] |> 
  rename(households_2020=V1)
hh29 <- fread(paste0(root,dir,'/urbansim_',SWM_run,'_2029_mapc_',MAPCM_run,'_2029_combined_microsim_households.csv'))
hh29 <- hh29[,lun(hid),.(bl10_id)] |> 
  rename(households_2030=V1) |> 
  full_join(hh19, by='bl10_id',keep=T) |> 
  setDT()
hh <- hh29 |> mutate(bl10_id=coalesce(bl10_id.x,bl10_id.y),.keep='unused') |> mutate(bl10_id = as.character(bl10_id))
hh[is.na(hh)] <- 0
hh[,household_change_2020_2030:=households_2030-households_2020]
setkey(hh,bl10_id)

# Employment trends
emp19 <- fread(paste0(root,dir,'/urbansim_',SWM_run,'_',MAPCM_run,'_emp_by_supernaics_2019_wide.csv')) |> select(bl10_id,total_emp) |> 
  rename(employment_2020 = total_emp)
emp <- fread(paste0(root,dir,'/urbansim_',SWM_run,'_',MAPCM_run,'_emp_by_supernaics_2029_wide.csv')) |> select(bl10_id,total_emp) |> 
  rename(employment_2030 = total_emp) |> 
  full_join(emp19,keep=T) |> 
  mutate(bl10_id = coalesce(bl10_id.x,bl10_id.y),.keep='unused') |> 
  mutate(bl10_id = as.character(bl10_id)) |> 
  setDT()
emp[is.na(emp)] <- 0
emp[,employment_change_2020_2030:=employment_2030-employment_2020]

# Units trends
sun19 <- fread(paste0(root,SWM_run,'/2019/residential_units.csv')) |> 
  filter(!block_id %in% mapc97_bl10) |> 
  mutate(units = 1) |> 
  group_by(block_id) |> 
  summarise(units2019 = sum(units)) |> 
  ungroup()
mun19 <- fread(paste0(root,MAPCM_run,'/2019/residential_units.csv')) |> 
  filter(block_id %in% mapc97_bl10) |> 
  mutate(units = 1) |> 
  group_by(block_id) |> 
  summarise(units2019 = sum(units)) |> 
  ungroup() |> 
  rbind(sun19)
sun29 <- fread(paste0(root,SWM_run,'/2029/residential_units.csv')) |> 
  filter(!block_id %in% mapc97_bl10) |> 
  mutate(units = 1) |> 
  group_by(block_id) |> 
  summarise(units2029 = sum(units)) |> 
  ungroup()
units <- fread(paste0(root,MAPCM_run,'/2029/residential_units.csv')) |> 
  filter(block_id %in% mapc97_bl10) |> 
  mutate(units = 1) |> 
  group_by(block_id) |> 
  summarise(units2029 = sum(units)) |> 
  ungroup() |> 
  rbind(sun29) |> 
  full_join(mun19,by='block_id',keep=T) |> 
  mutate(
    bl10_id = coalesce(block_id.x,block_id.y),.keep='unused') |> 
  setnames(c('units2030','units2020','bl10_id')) |> 
  mutate(bl10_id = as.character(bl10_id)) |> 
  setDT()
units[is.na(units2020),units2020:=0]
units[is.na(units2030),units2030:=0]
units[,units_added_2020_2030 := units2030-units2020]

comp <- keys_blocks |> 
  left_join(units, by='bl10_id') |> 
  left_join(emp, by='bl10_id') |> 
  left_join(hh, by='bl10_id') |> 
  filter(rpa_acr=='MAPC') |> 
  setDT()
comp[is.na(comp)] <- 0

# Massbuilds
m4 <- fread(paste0('massbuilds-20230401.csv'))
m4 <- m4[,.(id,loc_id,status,longitude,latitude,year_compl,hu,commsf,ret_sqft,ofcmd_sqft,indmf_sqft,whs_sqft,rnd_sqft,ei_sqft,other_sqft,hotel_sqft,gqpop)]
setnames(m4,'id','identification')
setkey(m4,identification)

mb <- fread('all_devprojects_4_4_2025.csv') |> 
  group_by(block_id,identification,name) |> 
  summarise(
    res_units = sum(residential_units,na.rm=T),
    emp_cap = sum(employment_capacity,na.rm=T)
    ) |>
  ungroup() |> 
  left_join(m4, by='identification') |> 
  filter(year_compl %in% 2021:2029) |> 
  select(-hu,-gqpop) |> 
  setDT()

fwrite(mb,paste0(root,dir,'lrtp_all_developments_summary.csv'))
cs <- names(mb)[c(4,5,11:19)]
mb.blk <- mb[,lapply(.SD,sum,na.rm=T),block_id,.SDcols=c(cs)]
setnames(mb.blk,'block_id','bl10_id')
mb.blk[,bl10_id:=as.character(bl10_id)]
setkey(mb.blk,bl10_id)
comp2 <- mb.blk[comp]
comp2[is.na(comp2)] <- 0

fwrite(mb.blk,paste0(root,dir,'lrtp_massbuilds_block_2020_2030.csv'))
fwrite(comp2,paste0(root,dir,'lrtp_2030_compare.csv'))

```

# 3.6.3 - Custom exports - Fenway TAP

```{r}

d2 <- paste0(r2,'Associated_Projects/Fenway_TAP_data_request/')

tzs <- c(26,87,92,93,108,109,253,254,255,256,257,258,259,260,261,262,263,264,265,266,267,268,269,270,271,272,273,274,275,276,277,278,279,280,281,282,284,285,288,289,293,729,735)

t19 <- fread(paste0(d2,'results_mapc_run_198_zone_indicators_2019.csv')) |> 
  select(zone_id,
         total_households,
         total_residential_units,
         sum_residential_unit_capacity,
         sum_vacant_residential_units,
         total_jobs,
         sum_employment_capacity) |> 
  filter(zone_id %in% tzs) |> 
  mutate(sum_residential_unit_capacity=round(sum_residential_unit_capacity)) |> 
  setDT()
  setnames(t19, paste0(names(t19),'_2019'))
  t19[,taz_id:=zone_id_2019][,zone_id_2019:=NULL] 


t30 <- fread(paste0(d2,'results_mapc_run_198_zone_indicators_2029.csv')) |> 
  select(zone_id,
         total_households,
         total_residential_units,
         sum_residential_unit_capacity,
         sum_vacant_residential_units,
         total_jobs,
         sum_employment_capacity) |> 
  filter(zone_id %in% tzs) |> 
  mutate(sum_residential_unit_capacity=round(sum_residential_unit_capacity)) |> 
  setDT()
  setnames(t30,paste0(names(t30),'_2030'))
  t30[,taz_id:=zone_id_2030][,zone_id_2030:=NULL] 
  
m4 <- fread(paste0(d2,'massbuilds20230401_taz_inter.csv')) |> 
  select(taz_id,id,loc_id,status,longitude,latitude,year_compl,hu,commsf,ret_sqft,ofcmd_sqft,indmf_sqft,whs_sqft,rnd_sqft,ei_sqft,other_sqft,hotel_sqft,gqpop) |> 
  setDT()

setnames(m4,'id','identification')
setkey(m4,identification)
# 
# taz.assign <- fread(paste0(r2,'Associated_Projects/taz_2010block_assignment_20230314.csv'))

mb <- fread(paste0(d2,'all_devprojects_4_4_2025.csv')) |> 
  group_by(identification,name) |> 
  summarise(
    res_units = sum(residential_units,na.rm=T),
    emp_cap = sum(employment_capacity,na.rm=T)
    ) |>
  ungroup() |> 
  left_join(m4, by='identification') |> 
  filter(year_compl %in% 2021:2029 & taz_id %in% tzs) |> 
  select(-hu,-gqpop) |> 
  setDT()
cs <- c('res_units','emp_cap','commsf','ret_sqft','ofcmd_sqft','indmf_sqft','whs_sqft','rnd_sqft','ei_sqft','other_sqft','hotel_sqft')

mb.taz <- mb[,lapply(.SD,sum,na.rm=T),taz_id,.SDcols=c(cs)]
fwrite(mb.taz,paste0(d2,'fenway_massbuilds_taz_2019_2030.csv'))

setkey(t19,taz_id)
setkey(t30,taz_id)
setkey(mb.taz, taz_id)
dt <- mb.taz[t19][t30]
dt[is.na(dt)] <- 0
  fwrite(dt,paste0(d2,'fenway_2019_2030_taz_comparison.csv'))


#   df <- list()
# for (yr in 2019:2029){
#   i=2030-yr
#   df[[i]] <- fread(paste0('results_mapc_run_198_muni_indicators_',yr,'.csv')) |>
#     summarize(total_emp = sum(total_jobs)) |> 
#     mutate(year=yr)
# }
  
```

# 3.7 - Project-scale land use adjustments

```{r}

# Set project name (name of project folder in Current_Projects/Projections/Projections_2023/Associated_Projects)

project <- 'Fenway_TAP_data_request'
proj.dir <- paste0(r2,'Associated_Projects/', project, '/')

# Project target file from consultant
targets.raw <- readxl::read_xlsx(paste0(proj.dir,'MAPC_LandUse_Adjustment_FTAP 2030_091924.xlsx'))

#target year
yr <- 2029

# export timestamp
stamp <- gsub('-', '', lubridate::today())

# UrbanSim run codes for TDM23.2.0
mrun <- 'run_198'
srun <- 'state_run_113'

# Select whether project is in MAPC or not
run <- mrun

urbsim_runs <- 'run_113-198_'

# TAZ crosswalks for TDM23

taz.assign <- fread(paste0(r2,'Associated_Projects/taz_2010block_assignment_20230314.csv'))
taz.alloc <- fread(paste0(r2,'Associated_Projects/taz_2010block_allocation_20230314.csv'))

# Load TDM23.2.0 land use for year in question
rundate <- ifelse(yr %in% c(2019,2049), '20250212', '20250507')

pop <- fread(paste0(root,'Microsim_outputs/ma_population_',urbsim_runs,yr,'_v',rundate,'.csv'))
emp <- fread(paste0(root,'Microsim_outputs/ma_employment_',urbsim_runs,yr,'_v',rundate,'.csv'))
pop.blk <- pop[person_num==1,.N,block_id]


# If employment adjustments are missing sector level data, then the existing distributions of sector-level data are used, unless the zone has zero existing employment. 
# ~~~~~~~~~~ !!! If the latter is the case, the consultant MUST specify sector-level distributions. !!! ~~~~~~~~~~~~~~~~~~

targets <- fread(paste0(proj.dir,'results_mapc_',run,'_zone_indicators_',yr,'.csv')) |> 
  select(zone_id,
         sum_jobs_aggr_sector_1,
         sum_jobs_aggr_sector_2,
         sum_jobs_aggr_sector_3,
         sum_jobs_aggr_sector_4,
         sum_jobs_aggr_sector_5,
         sum_jobs_aggr_sector_6,
         sum_jobs_aggr_sector_7,
         sum_jobs_aggr_sector_8,
         sum_jobs_aggr_sector_9,
         sum_jobs_aggr_sector_10,
         total_jobs) |> 
  rename(taz_id=zone_id) |> 
  right_join(targets.raw, by='taz_id') |> 
  mutate(emp1 = round(sum_jobs_aggr_sector_1 / total_jobs * emp_tot),
         emp2 = round(sum_jobs_aggr_sector_2 / total_jobs * emp_tot),
         emp3 = round(sum_jobs_aggr_sector_3 / total_jobs * emp_tot),
         emp4 = round(sum_jobs_aggr_sector_4 / total_jobs * emp_tot),
         emp5 = round(sum_jobs_aggr_sector_5 / total_jobs * emp_tot),
         emp6 = round(sum_jobs_aggr_sector_6 / total_jobs * emp_tot),
         emp7 = round(sum_jobs_aggr_sector_7 / total_jobs * emp_tot),
         emp8 = round(sum_jobs_aggr_sector_8 / total_jobs * emp_tot),
         emp9 = round(sum_jobs_aggr_sector_9 / total_jobs * emp_tot),
         emp10 = round(sum_jobs_aggr_sector_10 / total_jobs * emp_tot),
         emp_chk = emp1+emp2+emp3+emp4+emp5+emp6+emp7+emp8+emp9+emp10,
         emp_adj = emp_tot - emp_chk) |> 
  setDT()

targets[emp_tot > 0 & emp_chk == 0, emp8 := emp_tot]
targets[, emp_chk := emp1 + emp2 + emp3 + emp4 + emp5 + emp6 + emp7 + emp8 + emp9 + emp10]
targets[, emp_adj := emp_tot - emp_chk]

# Fix rounding errors
# Adjust employment in the largest target employment sector up or down 

col.chk <- .c(emp1, emp2, emp3, emp4, emp5, emp6, emp7, emp8, emp9, emp10)

ecols <- colnames(targets) %in% col.chk

targets$maxcol <- colnames(targets)[ecols][apply(targets[, ..col.chk], 1, which.max)]

for (j in 1:nrow(targets)) {
  rev <- targets$maxcol[j]
  targets[j, eval(rev) := get(eval(maxcol)) + emp_adj]
}

targets <- targets |> 
  select(taz_id,
         year,
         hh,
         emp_tot,
         emp1,
         emp2,
         emp3,
         emp4,
         emp5,
         emp6,
         emp7,
         emp8,
         emp9,
         emp10) |> 
  setDT()

# tlist <- fread(paste0(r2,'Associated_Projects/Western_Ave/Western_Ave_TAZ_sample_list.csv')) |> pull()

# Subset of TAZ from which replication samples will be taken
tlist <- unique(targets$taz)
blist <- taz.alloc[taz_id %in% tlist, block_id]
pop.sub <- pop[block_id %in% blist]

# Sample HH to add on
add.hid <- data.table()
sub.hid <- data.table()
for (taz in targets$taz_id) {
  if (targets[taz_id == taz, hh] > 0){
  x <- data.table(hid = sample(pop.sub$hid, targets[taz_id == taz, hh], replace = TRUE), new_block_id = sample(taz.alloc[taz_id == taz & area_fct==1, block_id], targets[taz_id == taz, hh], replace = TRUE))
  add.hid <- rbind(x, add.hid)
  }
  if (targets[taz_id == taz, hh] < 0){
  y <- sample(pop.sub[person_num==1, hid], targets[taz_id == taz, abs(hh)], replace = FALSE)
  sub.hid <- rbind(y, sub.hid)
  }
}
if(dim(sub.hid)[1] > 0){
  pop.sub <- pop.sub[!hid %in% sub.hid$x]
}

setorder(add.hid, hid)
add.hid[, cnt := rowid(hid)]

# Join full household/person records to replicated hids
add.pop <- add.hid |>
  left_join(pop.sub, by = 'hid', relationship = 'many-to-many')

# Append to original household table and recode hid value for replicated households
new.pop <- rbind(pop, add.pop, fill = TRUE) |> 
  mutate(
    hid = case_when(
      !is.na(cnt) ~ paste0(hid, '_', cnt),
      .default = hid
      ),
    block_id = case_when(
      !is.na(new_block_id) ~ new_block_id,
      .default = block_id
      )
    ) |> 
  select(-cnt, -new_block_id)

bpop <- new.pop[person_num==1,.N,block_id]

# Adjust employment in first block listed for each TAZ with a 100% area overlap
# Adjust total households column in employment file as needed as well

emp.adj <- emp |>
  left_join(taz.alloc, by='block_id') |> 
  left_join(targets, by='taz_id')

el <- data.table()

for(tz in targets$taz_id){
  x <- emp.adj[taz_id == tz & area_fct==1][1]|> 
    mutate(
      total_jobs = total_jobs + emp_tot,
      `1_constr` = `1_constr` + emp1,
      `2_eduhlth` = `2_eduhlth` + emp2,
      `3_finance` = `3_finance` + emp3,
      `4_public` = `4_public` + emp4,
      `5_info` = `5_info` + emp5,
      `6_ret_leis` = `6_ret_leis` + emp6,
      `7_manu` = `7_manu` + emp7,
      `8_other` = `8_other` + emp8,
      `9_profbus` = `9_profbus` + emp9,
      `10_ttu` = `10_ttu` + emp10
    )
  el <- rbind(el,x)
  }

e2 <- emp[!block_id %in% el$block_id]

ns <- names(emp)
new.emp <- el |> 
  select(all_of(ns)) |> 
  rbind(e2) |> 
  left_join(bpop, by='block_id') |> 
  mutate(total_households = case_when(
    !is.na(N) ~ N,
    .default=total_households)) |> 
  select(-N)
    
chk <- taz.alloc |> 
  left_join(new.emp, by='block_id') |> 
  left_join(bpop, by='block_id') |> 
  mutate(jobs = total_jobs * area_fct,
         hhds = N * area_fct) |> 
  group_by(taz_id) |> 
  summarize(emp.taz=round(sum(jobs, na.rm=T)),
            hh.taz=round(sum(hhds, na.rm=T))) |> 
  ungroup() |> 
  right_join(targets, by='taz_id')

chk2 <- taz.assign |> 
  left_join(new.emp, by='block_id') |> 
  left_join(bpop, by='block_id') |> 
  group_by(taz_id) |> 
  summarize(emp.taz=round(sum(total_jobs, na.rm=T)),
            hh.taz=round(sum(N, na.rm=T))) |> 
  ungroup() |> 
  right_join(targets, by='taz_id')

og.alloc <- taz.alloc |> 
  left_join(emp, by='block_id') |> 
  left_join(pop.blk, by='block_id') |> 
  mutate(jobs = total_jobs * area_fct,
         hhds = N * area_fct) |> 
  group_by(taz_id) |> 
  summarize(tdm23.emp.taz=round(sum(jobs, na.rm=T)),
            tdm23.hh.taz=round(sum(hhds, na.rm=T))) |> 
  right_join(chk, by='taz_id') |> 
  mutate(hh.chk=tdm23.hh.taz + hh,
         emp.chk=tdm23.emp.taz+emp_tot) |> 
  relocate(emp.chk,hh.chk, .before=year) |> 
  setDT()

og.assign <- taz.assign |> 
  left_join(emp, by='block_id') |> 
  left_join(pop.blk, by='block_id') |> 
  group_by(taz_id) |> 
  summarize(tdm23.emp.taz=round(sum(total_jobs, na.rm=T)),
            tdm23.hh.taz=round(sum(total_households, na.rm=T))) |> 
  ungroup() |> 
  right_join(chk2, by='taz_id') |> 
  mutate(hh.chk=tdm23.hh.taz + hh,
         emp.chk=tdm23.emp.taz+emp_tot) |> 
  relocate(emp.chk,hh.chk, .before=year) |> 
  setDT()

fwrite(og.assign,paste0(proj.dir,'QA_taz_assign.csv'))
fwrite(og.alloc,paste0(proj.dir,'QA_taz_alloc.csv'))

new.pop[,block_id:=as.character(block_id)]
new.emp[,block_id:=as.character(block_id)]

# Export
fwrite(new.pop,
  paste0(
    proj.dir,
    'ma_population_',urbsim_runs,
    yr,
    '_v',
    stamp,
    '_adjusted_',project,'_Project.csv'
  )
)
# Export
fwrite(new.emp,
  paste0(
    proj.dir,
    'ma_employment_',urbsim_runs,
    yr,
    '_v',
    stamp,
    '_adjusted_',project,'_Project.csv'
  )
)


```

# 8.1 - Load Municipal Indicators

```{r}


year <- 2049

new.run <- 'run_195'
old.run <- 'run_196'

new.urb <- fread(paste0(new.run,'/results_mapc_',new.run,'_muni_indicators_',year,'.csv')) |> 
  select(muni_id,
         total_households,
         total_jobs)
setnames(new.urb,'total_households',paste0('hh_',new.run))
setnames(new.urb,'total_jobs',paste0('emp_',new.run))

old.urb <- fread(paste0(old.run,'/results_mapc_',old.run,'_muni_indicators_',year,'.csv'))|> 
  select(muni_id,
         total_households,
         total_jobs)
setnames(old.urb,'total_households',paste0('hh_',old.run))
setnames(old.urb,'total_jobs',paste0('emp_',old.run))

comp.muni <- left_join(old.urb,new.urb, by='muni_id') |> 
  mutate(hh_diff = get(paste0('hh_',new.run)) - get(paste0('hh_',old.run)),
         emp_diff = get(paste0('emp_',new.run)) - get(paste0('emp_',old.run)))
```

# 8.2 - Load Census Block Indicators

```{r}
new.b <- fread(paste0(new.run,'/results_mapc_',new.run,'_block_indicators_',year,'.csv')) |> 
  select(block_id,
         total_households,
         total_jobs)
setnames(new.b,'total_households',paste0('hh_',new.run))
setnames(new.b,'total_jobs',paste0('emp_',new.run))

old.b <- fread(paste0(old.run,'/results_mapc_',old.run,'_block_indicators_',year,'.csv'))|> 
  select(block_id,
         total_households,
         total_jobs)
setnames(old.b,'total_households',paste0('hh_',old.run))
setnames(old.b,'total_jobs',paste0('emp_',old.run))

comp.blk <- left_join(old.b,new.b, by='block_id') |> 
  mutate(hh_diff = get(paste0('hh_',new.run)) - get(paste0('hh_',old.run)),
         emp_diff = get(paste0('emp_',new.run)) - get(paste0('emp_',old.run)),
         block_id = as.character(block_id))
```

# 8.3 - Spatial data for mapping

```{r}

#| echo: false
########## Spatial data for mapping ###### 
######################################### 

# Need to use 2020 census for munis because mapc datakeys has GEOIDs for 2020 munis
sfm <-
  get_decennial(
    year = 2020,
    state = 'MA',
    geography = 'county subdivision',
    variables = 'H1_001N',
    geometry = T
  ) |> 
  mutate(GEOID=as.numeric(GEOID)) |> 
  select(GEOID,geometry)

ms <- mapcdatakeys::all_muni_data_keys |>
  select(muni_id,muni_name,cosub_cn20) |> 
  rename(GEOID = cosub_cn20) |> 
  left_join(comp.muni, by='muni_id')

sf.muni <- left_join(sfm,ms,by='GEOID') |> 
  filter(!is.na(hh_diff))

sfb <-
  get_decennial(
    year = 2010,
    state = 'MA',
    geography = 'block',
    variables = 'P001001',
    geometry = T
  ) |> 
  select(GEOID,geometry)

sf.blk <- left_join(sfb, comp.blk, by=c('GEOID' = 'block_id')) |> filter(!is.na(hh_diff))

```

# 8.4 - Visualize muni differences

```{r}
#| echo: false
# Palette for household difference
bins_hh_diff<-
  c(-Inf, -2000, -1000, -500, -250, -1, 1, 250, 500, 1000, 2000, Inf)
pal_hh_diff <-
  colorBin(c("blue", "white", "red"),
           bins = bins_hh_diff,
           na.color = "#808080")

# Palette for employment difference
bins_emp_diff<-
  c(-Inf, -2000, -1000, -500, -250, -1, 1, 250, 500, 1000, 2000, Inf)
pal_emp_diff <-
  colorBin(c("blue", "white", "red"),
           bins = bins_emp_diff,
           na.color = "#808080")

muni_diff <- sf.muni |>
  st_transform(st_crs(4326)) |>
  leaflet() |>
  addProviderTiles(provider = "CartoDB.Positron") |>
  addPolygons(
    popup = ~ paste(
      "Muni: ",
      sf.muni$muni_name,
      "<br>",
      paste0('Households ',new.run, ' = '),
      sf.muni[[paste0('hh_',new.run)]],
      "<br>",
      paste0('Households ',old.run, ' = '),
      sf.muni[[paste0('hh_',old.run)]],
      "<br>",
      'Household Difference = ',
      sf.muni$hh_diff
    ),
    label=~muni_name,
    fillOpacity = 0.6,
    fillColor = ~ pal_hh_diff(sf.muni$hh_diff),
    weight = 1,
    stroke=T,
    col='black',
    group = "Households"
  ) |>
  addLegend(
    position = "bottomleft",
    pal = pal_hh_diff,
    values = ~ bins_hh_diff,
    title = paste(
      "Households difference between <br> ",new.run,'and',old.run),
    opacity = 1,
    group = "Households",
    className = "info legend Households",
    layerId = 'Households'
  ) |>
  addPolygons(
    popup = ~ paste(
      "Muni: ",
      sf.muni$muni_name,
      "<br>",
      paste0('Employment ',new.run, ' = '),
      sf.muni[[paste0('emp_',new.run)]],
      "<br>",
      paste0('Employment ',old.run, ' = '),
      sf.muni[[paste0('emp_',old.run)]],
      "<br>",
      "Employment Difference = ",
      sf.muni$emp_diff
    ),
    label=~muni_name,
    fillOpacity = 0.6,
    fillColor = ~ pal_emp_diff(sf.muni$emp_diff),
    weight = 1,
    stroke=T,
    col='black',
    group = "Employment"
  ) |>
  addLegend(
    position = "bottomleft",
    pal = pal_emp_diff,
    values = ~ bins_emp_diff,
    title = paste(
      "Employment difference between <br> ",new.run,'and',old.run),
    opacity = 1,
    group = "Employment",
    className = "info legend Employment",
    layerId = 'Employment'
  ) |>
  addSearchFeatures(
    targetGroups = "Households",
    options = searchFeaturesOptions(
      zoom = 10,
      position = 'topright',
      textPlaceholder = "Search by Town Name..."
    )
  ) |> 
  # Layer controls
  addLayersControl(
    baseGroups = c(
      "Households",
      "Employment"
    )
  ) |> 
  htmlwidgets::onRender(
    "
      function(el, x) {
         var updateLegend = function () {
            var selectedGroup = document.querySelectorAll('input:checked')[0].nextSibling.innerText.substr(1);
            var selectedClass = selectedGroup.replace(' ', '');
            document.querySelectorAll('.legend').forEach(a => a.hidden=true);
            document.querySelectorAll('.legend').forEach(l => {
               if (l.classList.contains(selectedClass)) l.hidden=false;
            });
         };
         updateLegend();
         this.on('baselayerchange', el => updateLegend());
      }"
  )

```

# 8.5 - Visualize block differences

```{r}

block_diff <- sf.blk |>
  st_transform(st_crs(4326)) |>
  leaflet() |>
  addProviderTiles(provider = "CartoDB.Positron") |>
  addPolygons(
    popup = ~ paste(
      "Muni: ",
      sf.blk$GEOID,
      "<br>",
      paste0('Households ',new.run, ' = '),
      sf.blk[[paste0('hh_',new.run)]],
      "<br>",
      paste0('Households ',old.run, ' = '),
      sf.blk[[paste0('hh_',old.run)]],
      "<br>",
      'Household Difference = ',
      sf.blk$hh_diff
    ),
    label=~GEOID,
    fillOpacity = 0.4,
    fillColor = ~ pal_hh_diff(sf.blk$hh_diff),
    weight = 0.4,
    stroke=T,
    col='grey80',
    group = "Households"
  ) |>
  addLegend(
    position = "bottomleft",
    pal = pal_hh_diff,
    values = ~ bins_hh_diff,
    title = paste(
      "Households difference between <br> ",new.run,'and',old.run),
    opacity = 1,
    group = "Households",
    className = "info legend Households"
  ) |>
  addPolygons(
    popup = ~ paste(
      "Block: ",
      sf.blk$GEOID,
      "<br>",
      paste0('Employment ',new.run, ' = '),
      sf.blk[[paste0('emp_',new.run)]],
      "<br>",
      paste0('Employment ',old.run, ' = '),
      sf.blk[[paste0('emp_',old.run)]],
      "<br>",
      "Employment Difference = ",
      sf.blk$emp_diff
    ),
    label=~GEOID,
    fillOpacity = 0.4,
    fillColor = ~ pal_emp_diff(sf.blk$emp_diff),
    weight = 0.4,
    stroke=T,
    col='grey80',
    group = "Employment"
  ) |>
  addLegend(
    position = "bottomleft",
    pal = pal_emp_diff,
    values = ~ bins_emp_diff,
    title = paste(
      "Employment difference between <br> ",new.run,'and',old.run),
    opacity = 1,
    group = "Employment",
    className = "info legend Employment"
  ) |>
  addSearchFeatures(
    targetGroups = "Households",
    options = searchFeaturesOptions(
      zoom = 10,
      position = 'topright',
      textPlaceholder = "Search by Census Block ID..."
    )
  ) |> 
  # Layer controls
  addLayersControl(
    baseGroups = c(
      "Households",
      "Employment"
    )
  ) |> 
  htmlwidgets::onRender(
    "
      function(el, x) {
         var updateLegend = function () {
            var selectedGroup = document.querySelectorAll('input:checked')[0].nextSibling.innerText.substr(1);
            var selectedClass = selectedGroup.replace(' ', '');
            document.querySelectorAll('.legend').forEach(a => a.hidden=true);
            document.querySelectorAll('.legend').forEach(l => {
               if (l.classList.contains(selectedClass)) l.hidden=false;
            });
         };
         updateLegend();
         this.on('baselayerchange', el => updateLegend());
      }"
)

```

# 8.6 - Difference at municipal level:

```{r}
muni_diff

saveRDS(muni_diff,paste0(comp.path,'muni_comparison_map_y',year,'_',new.run,'_',old.run,'.rds'))
# saveWidget(
#   widget = muni_diff,
#   selfcontained = F,
#   file = paste0(
#     comp.path,new.run,'_',old.run,
#     "/muni_comparison_y",year,"_tdm19_tdm23_",
#     Sys.Date(),
#     ".html"
#   )
#   )
```

# 8.7 - Difference at Census Block level:

```{r}
block_diff
saveRDS(block_diff,paste0(comp.path,'block_comparison_map_y',year,'_',new.run,'_',old.run,'.csv'))

# saveWidget(
#   widget = block_diff,
#   file = paste0(
#     comp.path,new.run,'_',old.run,
#     "/block_comparison_y",year,"_tdm19_tdm23_",
#     Sys.Date(),
#     ".html"
#   )
# )
```

# 8.8 - Compare TAZ destination SOV average costs

```{r}
taz <- st_read('../Skims/ctps_taz_statewide.shp')
park <- fread('../Skims/TDM23/tdm23.1.0/parking_v20221007.csv')

nskim <- fread('../Skims/TDM23/tdm23.1.0/mapc_2050plan_skims_freepark_tdm23.1.0.csv') |>
  select(to_zone_id,sov_am_totcost) |> 
  rename(new_sov_cost=sov_am_totcost) |> 
  summarize(new_sov_cost = mean(new_sov_cost),.by='to_zone_id')

oskim <- fread('../Skims/travel_model_skims_2040_id50.csv') |>
  select(to_zone_id,sov_am_totcost) |> 
  rename(old_sov_cost=sov_am_totcost) |> 
  summarize(old_sov_cost = mean(old_sov_cost),.by='to_zone_id') |> 
  left_join(nskim, by='to_zone_id') |> 
  left_join(park, by=c('to_zone_id' = 'taz_id'))

sov <- inner_join(taz,oskim,by=c('tazID'='to_zone_id')) |> 
  mutate(cost_diff = new_sov_cost - old_sov_cost)

bins_cost_diff<-
  c(-Inf, -20, -10, -5, -1, 1, 5, 10, 20, Inf)
pal_cost_diff <-
  colorBin(c("blue", "white", "red"),
           bins = bins_cost_diff,
           na.color = "#808080")

cost_diff <- sov |>
  st_transform(st_crs(4326)) |>
  leaflet() |>
  addProviderTiles(provider = "CartoDB.Positron") |>
  addPolygons(
    popup = ~ paste(
      "TAZ: ",
      sov$to_zone_id,
      "<br>",
      paste0('New Skim - SOV cost = '),
      sov$new.sov,
      "<br>",
      paste0('Old Skim - SOV cost = '),
      sov$old.sov,
      "<br>",
      'SOV cost difference = ',
      sov$cost_diff
    ),
    label=~ID,
    fillOpacity = 0.4,
    fillColor = ~ pal_cost_diff(sov$cost_diff),
    weight = 0.4,
    stroke=T,
    col='grey80',
    group = "New Skims"
  ) |>
  addLegend(
    position = "bottomleft",
    pal = pal_cost_diff,
    values = ~ bins_cost_diff,
    title = paste(
      "SOV total cost difference"),
    opacity = 1,
    group = "New Skims",
    className = "info legend New Skims"
  ) |>
  # # Layer controls
  # addLayersControl(
  #   baseGroups = c(
  #     "Old Skims",
  #     "New Skims"
  #   )
  # ) |> 
  htmlwidgets::onRender(
    "
      function(el, x) {
         var updateLegend = function () {
            var selectedGroup = document.querySelectorAll('input:checked')[0].nextSibling.innerText.substr(1);
            var selectedClass = selectedGroup.replace(' ', '');
            document.querySelectorAll('.legend').forEach(a => a.hidden=true);
            document.querySelectorAll('.legend').forEach(l => {
               if (l.classList.contains(selectedClass)) l.hidden=false;
            });
         };
         updateLegend();
         this.on('baselayerchange', el => updateLegend());
      }"
  )

```

```{r}

for(year in c(2019,2049)){

m19 <- fread(paste0('run_196/results_mapc_run_196_block_indicators_',year,'.csv'))
s19 <- fread(paste0('state_run_114/results_mapc_state_run_114_block_indicators_',year,'.csv'))

m23 <- fread(paste0('run_195/results_mapc_run_195_block_indicators_',year,'.csv'))
s23 <- fread(paste0('state_run_113/results_mapc_state_run_113_block_indicators_',year,'.csv'))

ms <- unique(m19$muni_id)
s19 <- s19[!muni_id%in%ms]
s23 <- s23[!muni_id%in%ms]

blk19 <- rbind(m19,s19,fill=T)
blk23 <- rbind(m23,s23,fill=T)

ns <- names(blk19)[!grepl('sz_',names(blk19))]
blk19 <- blk19 |> select(all_of(ns)) |> relocate(c(total_households,total_jobs),.after=muni_id)
setorder(blk19,muni_id,block_id)

blk23 <- blk23 |> select(all_of(ns)) |> relocate(c(total_households,total_jobs),.after=muni_id)
setorder(blk23,muni_id,block_id)

mn19 <- blk19[,lapply(.SD,sum,na.rm=T),muni_id]
mn19 <- mn19 |> 
  select(all_of(ns)) |>
  select(-block_id,-density_jobs,-density_households,density_residential_units,-median_value,-median_income,-ru_vacancy_rate) 
mmm <- sort(names(mn19))
mn19 <- mn19[,(mmm),with=F]
mn19 <- mn19 |> relocate(c(muni_id,total_households,total_jobs),.before=acres)
setorder(mn19,muni_id)

mn23 <- blk23[,lapply(.SD,sum,na.rm=T),muni_id]
mn23 <- mn23 |> 
  select(all_of(ns)) |>
  select(-block_id,-density_jobs,-density_households,density_residential_units,-median_value,-median_income,-ru_vacancy_rate) 
mmm <- sort(names(mn23))
mn23 <- mn23[,(mmm),with=F]
mn23 <- mn23 |> relocate(c(muni_id,total_households,total_jobs),.before=acres)
setorder(mn23,muni_id)

fwrite(blk19,paste0('urbansim_block_indicators_TDM19_year',year,'.csv'))
fwrite(blk23,paste0('urbansim_block_indicators_TDM23_year',year,'.csv'))
fwrite(mn19,paste0('urbansim_muni_indicators_TDM19_year',year,'.csv'))
fwrite(mn23,paste0('urbansim_muni_indicators_TDM23_year',year,'.csv'))

}


```

# 3.3 - Post-processor 03 - person shifts

```{r}

adiff <- fread(paste0(root,run,'/urbansim_',run,'_umdi_pop_age_difference_v3.csv'))

for (yr in years) {
  dt <- readRDS(
    paste0(root,
      run,
      '/urbansim_',
      run,
      '_microhouseholds_',
      yr,
      '_unrevised_v2.rds'
    )
  )

  dt_shift <- data.table()
  
  
  if(run==SWM_run){
    msub <- mpos[mpos!='MAPC']
  }
  if(run==MAPCM_run){
    msub <- 'MAPC'
  }
    for (m in msub) {
    diff <- adiff[mpo == m & year == yr]
    swm <- dt[mpo == m]
    
    # ageCAT 2 and 3
    for (i in 2:3) {
      dlag <- diff[ageCAT6 == i - 1, diff]
      da <- diff[ageCAT6 == i, diff]
      dlead <- diff[ageCAT6 == i + 1, diff]
      
      if (da > 0 & dlag < 0 & dlead < 0) {
        if ((abs(dlag) + abs(dlead)) > da) {
          subN <- abs(round(da * (dlag / (dlag + dlead))))
          addN <- abs(round(da * (dlead / (dlag + dlead))))
        }
        else if ((abs(dlag) + abs(dlead)) <= da) {
          subN <- abs(dlag)
          addN <- abs(dlead)
        }
        tosub <-
          swm[ageCAT6 == i &
                person_num > 1, sample(unique(pid), subN)]
        toadd <-
          swm[ageCAT6 == i &
                person_num > 1, sample(unique(pid), addN)]
        swm[pid %in% tosub, `:=` (age = age - 5, ageCAT6 = ageCAT6 - 1)]
        swm[pid %in% toadd, `:=` (age = age + 5, ageCAT6 = ageCAT6 + 1)]
      }
      if (da > 0 & dlag < 0 & dlead > 0) {
        if (abs(dlag) > da) {
          subN <- abs(da)
        }
        else if (abs(dlag) <= da) {
          subN <- abs(dlag)
        }
        tosub <-
          swm[ageCAT6 == i &
                person_num > 1, sample(unique(pid), subN)]
        swm[pid %in% tosub, `:=` (age = age - 5, ageCAT6 = ageCAT6 - 1)]
      }
      if (da > 0 & dlag > 0 & dlead < 0) {
        if (abs(dlead) > da) {
          addN <- abs(da)
        }
        else if (abs(dlead) <= da) {
          addN <- abs(dlead)
        }
        toadd <-
          swm[ageCAT6 == i &
                person_num > 1, sample(unique(pid), addN)]
        swm[pid %in% toadd, `:=` (age = age + 5, ageCAT6 = ageCAT6 + 1)]
      }
      # Update age difference table
      newlag <- swm[ageCAT6 == i - 1, .N]
      newpop <- swm[ageCAT6 == i, .N]
      newlead <- swm[ageCAT6 == i + 1, .N]
      diff[ageCAT6 == i - 1, `:=`(urbansim_hh_pop = newlag, diff = newlag - umdi_hh_pop)]
      diff[ageCAT6 == i, `:=`(urbansim_hh_pop = newpop, diff = newpop - umdi_hh_pop)]
      diff[ageCAT6 == i + 1, `:=`(urbansim_hh_pop = newlead, diff = newlead - umdi_hh_pop)]
      suppressWarnings(rm(newlag,newpop,newlead,addN,subN,da,dlag,dlead,tosub,toadd))
      gc()
    }
    
    # ageCAT = 1
    da <- diff[ageCAT6 == 1, diff]
    dlead <- diff[ageCAT6 == 2, diff]
    if (da > 0 & dlead < 0) {
      if (abs(dlead) > da) {
        addN <- abs(da)
      }
      else if (abs(dlead) <= da) {
        addN <- abs(dlead)
      }
      toadd <-
        swm[ageCAT6 == 1 &
              person_num > 1, sample(unique(pid), addN)]
      swm[pid %in% toadd, `:=` (age = age + 5, ageCAT6 = ageCAT6 + 1)]
    }
    # Update age difference table
    newpop <- swm[ageCAT6 == 1, .N]
    newlead <- swm[ageCAT6 == 2, .N]
    diff[ageCAT6 == 1, `:=`(urbansim_hh_pop = newpop, diff = newpop - umdi_hh_pop)]
    diff[ageCAT6 == 2, `:=`(urbansim_hh_pop = newlead, diff = newlead - umdi_hh_pop)]
    suppressWarnings(rm(newlag,newpop,newlead,addN,subN,da,dlag,dlead,tosub,toadd))
    gc()
    
    # ageCAT = 4
    dlag <- diff[ageCAT6 == 3, diff]
    da <- diff[ageCAT6 == 4, diff]
    dlead <- diff[ageCAT6 == 5, diff]
    if (da > 0 & dlag < 0) {
      if (abs(dlag) > da) {
        subN <- abs(da)
      }
      else if (abs(dlag) <= da) {
        subN <- abs(dlag)
      }
      tosub <-
        swm[age %in% 15:18 &
              person_num > 1, sample(unique(pid), subN)]
      swm[pid %in% tosub, `:=` (age = age - 5, ageCAT6 = ageCAT6 - 1)]
    }
    # Update age difference table
    newpop <- swm[ageCAT6 == 4, .N]
    newlag <- swm[ageCAT6 == 3, .N]
    diff[ageCAT6 == 4, `:=`(urbansim_hh_pop = newpop, diff = newpop - umdi_hh_pop)]
    diff[ageCAT6 == 3, `:=`(urbansim_hh_pop = newlag, diff = newlag - umdi_hh_pop)]
    suppressWarnings(rm(newlag,newpop,newlead,addN,subN,da,dlag,dlead,tosub,toadd))
    gc()
    
    for (i in 6:17) {
      dlag <- diff[ageCAT6 == i - 1, diff]
      da <- diff[ageCAT6 == i, diff]
      dlead <- diff[ageCAT6 == i + 1, diff]
      
      if (da > 0 & dlag < 0 & dlead < 0) {
        if ((abs(dlag) + abs(dlead)) > da) {
          subN <- abs(round(da * (dlag / (dlag + dlead))))
          addN <- abs(round(da * (dlead / (dlag + dlead))))
        }
        else if ((abs(dlag) + abs(dlead)) <= da) {
          subN <- abs(dlag)
          addN <- abs(dlead)
        }
        tosub <-
          swm[ageCAT6 == i &
                person_num > 1, sample(unique(pid), subN)]
        toadd <-
          swm[ageCAT6 == i &
                person_num > 1, sample(unique(pid), addN)]
        swm[pid %in% tosub, `:=` (age = age - 5, ageCAT6 = ageCAT6 - 1)]
        swm[pid %in% toadd, `:=` (age = age + 5, ageCAT6 = ageCAT6 + 1)]
      }
      if (da > 0 & dlag < 0 & dlead > 0) {
        if (abs(dlag) > da) {
          subN <- abs(da)
        }
        else if (abs(dlag) <= da) {
          subN <- abs(dlag)
        }
        tosub <-
          swm[ageCAT6 == i &
                person_num > 1, sample(unique(pid), subN)]
        swm[pid %in% tosub, `:=` (age = age - 5, ageCAT6 = ageCAT6 - 1)]
      }
      # Update age difference table
      newpop <- swm[ageCAT6 == i, .N]
      newlag <- swm[ageCAT6 == i - 1, .N]
      newlead <- swm[ageCAT6 == i + 1, .N]
      diff[ageCAT6 == i - 1, `:=`(urbansim_hh_pop = newlag, diff = newlag - umdi_hh_pop)]
      diff[ageCAT6 == i, `:=`(urbansim_hh_pop = newpop, diff = newpop - umdi_hh_pop)]
      diff[ageCAT6 == i + 1, `:=`(urbansim_hh_pop = newlead, diff = newlead - umdi_hh_pop)]
      suppressWarnings(rm(newlag,newpop,newlead,addN,subN,da,dlag,dlead,tosub,toadd))
      gc()
      
    }
    
    # ageCAT = 5
    da <- diff[ageCAT6 == 5, diff]
    dlead <- diff[ageCAT6 == 6, diff]
    if (da > 0 & dlead < 0) {
      if (abs(dlead) > da) {
        addN <- abs(da)
      }
      else if (abs(dlead) <= da) {
        addN <- abs(dlead)
      }
      toadd <-
        swm[ageCAT6 == 5 &
              person_num > 1, sample(unique(pid), addN)]
      swm[pid %in% toadd, `:=` (age = age + 5, ageCAT6 = ageCAT6 + 1)]
    }
    # Update age difference table
    newpop <- swm[ageCAT6 == 5, .N]
    newlead <- swm[ageCAT6 == 6, .N]
    diff[ageCAT6 == 5, `:=`(urbansim_hh_pop = newpop, diff = newpop - umdi_hh_pop)]
    diff[ageCAT6 == 6, `:=`(urbansim_hh_pop = newlead, diff = newlead - umdi_hh_pop)]
    
    suppressWarnings(rm(newlag,newpop,newlead,addN,subN,da,dlag,dlead,tosub,toadd))
    gc()
    
    # ageCAT = 18
    da <- diff[ageCAT6 == 18, diff]
    dlag <- diff[ageCAT6 == 17, diff]
    if (da > 0 & dlag < 0) {
      if (abs(dlag) > da) {
        subN <- abs(da)
      }
      else if (abs(dlag) <= da) {
        subN <- abs(dlag)
      }
      avail <-  swm[ageCAT6 == 18 &
                      person_num > 1, .N]
      tosub <-
        swm[ageCAT6 == 18 &
              person_num > 1, sample(unique(pid), min(avail, subN))]
      swm[pid %in% tosub, `:=` (age = age - 5, ageCAT6 = ageCAT6 - 1)]
    }
    if (da < 0 & dlag > 0) {
      if (abs(da) > dlag) {
        addN <- abs(dlag)
      }
      else if (abs(da) <= dlag) {
        addN <- abs(da)
      }
      toadd <-
        swm[ageCAT6 == 17 &
              person_num > 1, sample(unique(pid), addN)]
      swm[pid %in% toadd, `:=` (age = age + 5, ageCAT6 = ageCAT6 + 1)]
    }
    # Update age difference table
    newpop <- swm[ageCAT6 == 18, .N]
    newlag <- swm[ageCAT6 == 17, .N]
    diff[ageCAT6 == 18, `:=`(urbansim_hh_pop = newpop, diff = newpop - umdi_hh_pop)]
    diff[ageCAT6 == 17, `:=`(urbansim_hh_pop = newlag, diff = newlag - umdi_hh_pop)]
    
    suppressWarnings(rm(newlag,newpop,newlead,addN,subN,da,dlag,dlead,tosub,toadd))
    gc()
    
    # Recalculate household characteristics
    # Persons by 5-yr age categories
    swm[age < 5 , ageCAT6 := 1]
    swm[age %in% 5:9 , ageCAT6 := 2]
    swm[age %in% 10:14 , ageCAT6 := 3]
    swm[age %in% 15:19 , ageCAT6 := 4]
    swm[age %in% 20:24 , ageCAT6 := 5]
    swm[age %in% 25:29 , ageCAT6 := 6]
    swm[age %in% 30:34 , ageCAT6 := 7]
    swm[age %in% 35:39 , ageCAT6 := 8]
    swm[age %in% 40:44 , ageCAT6 := 9]
    swm[age %in% 45:49 , ageCAT6 := 10]
    swm[age %in% 50:54 , ageCAT6 := 11]
    swm[age %in% 55:59 , ageCAT6 := 12]
    swm[age %in% 60:64 , ageCAT6 := 13]
    swm[age %in% 65:69 , ageCAT6 := 14]
    swm[age %in% 70:74 , ageCAT6 := 15]
    swm[age %in% 75:79 , ageCAT6 := 16]
    swm[age %in% 80:84 , ageCAT6 := 17]
    swm[age >= 85 , ageCAT6 := 18]
    # Count of persons by ageCAT6 in each HHd
    swm[ageCAT6 == 1, age1 := .N, hid]
    swm[ageCAT6 == 2, age2 := .N, hid]
    swm[ageCAT6 == 3, age3 := .N, hid]
    swm[ageCAT6 == 4, age4 := .N, hid]
    swm[ageCAT6 == 5, age5 := .N, hid]
    swm[ageCAT6 == 6, age6 := .N, hid]
    swm[ageCAT6 == 7, age7 := .N, hid]
    swm[ageCAT6 == 8, age8 := .N, hid]
    swm[ageCAT6 == 9, age9 := .N, hid]
    swm[ageCAT6 == 10, age10 := .N, hid]
    swm[ageCAT6 == 11, age11 := .N, hid]
    swm[ageCAT6 == 12, age12 := .N, hid]
    swm[ageCAT6 == 13, age13 := .N, hid]
    swm[ageCAT6 == 14, age14 := .N, hid]
    swm[ageCAT6 == 15, age15 := .N, hid]
    swm[ageCAT6 == 16, age16 := .N, hid]
    swm[ageCAT6 == 17, age17 := .N, hid]
    swm[ageCAT6 == 18, age18 := .N, hid]
    dt_shift <- rbind(dt_shift, swm)
    message(paste0(m, ' complete ', yr))
  }
  saveRDS(dt_shift,paste0(root, run,'/urbansim_pop_shifted_',run,'_',yr,'_v3.rds'))
}

# Recalculate age differences to UMDI
fs <-
  list.files(path = paste0(root, run),
             pattern = glob2rx("urbansim_pop_shifted*v3*"),
             full.names = TRUE)

if (run == SWM_run) {
  readComp <- function(x) {
    dt <- readRDS(x)
    dt[, pop := 1]
    dt <-
      dt[, lapply(.SD, sum), .(mpo, year, ageCAT6), .SDcols = 'pop']
  }
  urbsim <- lapply(fs, readComp)
  uspop <- rbindlist(urbsim)
  uspop <- uspop[!is.na(mpo)]
  setkey(uspop, mpo, year, ageCAT6)
  
  adiff <- umdi[uspop]
  adiff[, diff := pop - umdi_hh_pop]
  setnames(adiff, 'pop', 'urbansim_hh_pop')
  fwrite(adiff,
         paste0(root,
           run,
           '/urbansim_',
           run,
           '_post_shift_umdi_pop_age_difference_v3.csv'
         ))
}
if (run == MAPCM_run) {
  readComp <- function(x) {
    dt <- readRDS(x)
    dt[, pop := 1]
    dt <-
      dt[, lapply(.SD, sum), .(mpo, year, ageCAT6), .SDcols = 'pop']
  }
  urbsim <- lapply(fs, readComp)
  uspop <- rbindlist(urbsim)
  uspop <- uspop[!is.na(mpo)]
  setkey(uspop, mpo, year, ageCAT6)
  
  adiff <- umdi[uspop]
  adiff[, diff := pop - umdi_hh_pop]
  setnames(adiff, 'pop', 'urbansim_hh_pop')
  fwrite(adiff,
         paste0(root,
           run,
           '/urbansim_',
           run,
           '_post_shift_umdi_pop_age_difference_v3.csv'
         ))
}

```

# 2- Unpack the h5 files in python

# The following packages must be installed in your python environment:

pandas h5py tables

#\\Data-Sync\users$\cgately\Documents\.virtualenvs\r-reticulate\Scripts\python.exe -m pip install --upgrade pip #\\Data-Sync\users$\cgately\Documents.virtualenvs\r-reticulate\Scripts\python.exe -m pip install pandas #\\Data-Sync\users$\cgately\Documents\.virtualenvs\r-reticulate\Scripts\python.exe -m pip install h5py #\\Data-Sync\users$\cgately\Documents.virtualenvs\r-reticulate\Scripts\python.exe -m pip install tables

```{r}
use_python('C:\\Users\\gately\\anaconda3')
repl_python()

```

```{python}
#| eval: false
# C:\Users\cgately\AppData\Local\anaconda3
# C:\Users\gately\anaconda3
# use_python('C:/Users/gately/Documents/.virtualenvs/r-reticulate/Scripts')

import os
#import numpy as np
import pandas as pd
import h5py

# Set Name of UrbanSim Model Run

run = r.MAPCM_run


# In-Office path
# os.chdir('K:\\DataServices\\Projects\\Current_Projects\\Projections\\Projections_2023\\Data\\03_UrbanSim\\UrbanSim_Outputs\\' + run)
#os.chdir('/mnt/k/DataServices/Projects/Current_Projects/Projections/Data/03_UrbanSim/UrbanSim_Outputs/' + run)

# Remote path
dir = 'S:\\Network Shares\\K Drive\\DataServices\\Projects\\Current_Projects\\Projections\\Projections_2023\\Data\\03_UrbanSim\\UrbanSim_Outputs\\' + run + '\\'
# os.chdir('/mnt/s/Network Shares/K Drive/DataServices/Projects/Current_Projects/Projections/Projections_2023/Data/03_UrbanSim/UrbanSim_Outputs/' + run)


# Years in the H5 file
# years = ['2010','2019','2029','2034']
years = ['2010','2019','2029','2039','2049']
for yr in years:
  if not os.path.exists(dir + yr):
    os.makedirs(dir + yr)

h5=pd.HDFStore(dir + 'results_mapc_' + run + '_run_results.h5')

keys = h5.keys()

for key in keys:
  out = h5[key]
  out.to_csv('.' + key + ".csv", header = True, index = False)
  print("export of " + key + " complete")

exit
```
