---
title: "Urbansim Model Update Exploratory Analysis"
format: html
editor: visual
---

# 0.0 Environment and Script Set-up

# 0.1 Load Libraries and Functions

```{r}
# {here} library to initiate conditional filepaths
library(here)
# Libraries used in Urbansim/Projections workflow
source(here("urbansim_model_update/projections_libraries.R"))
# Functions created for the Urbansim/Projections workflow
source(here("urbansim_model_update/projections_functions.R"))

```

# 0.2 Set Environment Variables for this Script

```{r}
# Set random seed
set.seed(351)

# Remove scientific notation
options(scipen = 999)

# Set the PUMS API call variables for this analysis. Used in the 
# Variables from the PUMS data to query (data dictionary is here: https://www.census.gov/programs-surveys/acs/microdata/documentation.2023.html#list-tab-1370939201)
pums.vars <- c(
               # General PUMS variables
               'PUMA', #PUMA designation
               'ADJINC', # Inflation adjustment factor for income variables
               'TYPEHUGQ', # Type of household (Household, institutional, non-institutional group quarters)
  
               # Demographic variables
               'AGEP', # Age of person
               "HHLDRAGEP", #Householder age 
               "SEX", # Sex
               'NP', # Number of persons in a household
               'TEN', # Tenure
               'RAC1P', # Race
               'HISP', # Recoded detailed Hispanic origin (Ethnicity)
               'R18', #Presence of persons under 18 years in household
               'R60', # Presence of persons over 60 years in household
               'R65', # Presence of persons over 65 years in household
               'ESR', #Employment status recode
  
               # Income variables
               "FINCP", # Family Income
               "HINCP", # Household Income
               "INTP", # Interest, dividends, and net rental income
               "PINCP", # Total person's income
               "WAGP", # Wages or salary income past 12 months
               "SSIP", # Supplementary Security Income
               "SSP", # Social Security income
               "SEMP", # Self employment income
               "RETP", # Retirement income
               "PAP", # Public assistance income
               "OIP", #All other income
               
               # Overcrowding and Household Composition variables
               'BDSP', # Number of bedrooms
               'RMSP', # Number of rooms
               'CPLT', # Couple Type
               'RELSHIPP', # Relationship to the householder
               'PSF', # Presence of subfamily
               'SFN', # Subfamily number
               'SFR', # Subfamily relationship
               'NR', # Presence of non-relative in household
               'NRC', # Number of related children in a household
               'MULTG', # Multigenerational household flag
               'HHT', # Household/family type
               'HHT2', # Household family type (including cohabiting)
               "PARTNER", # Unmarried partner household
               'NPP', # Grandparent headed household with no parent present
               'NPF', # Number of persons in family
               'HUGCL', # Household with grandparent living with grandchildren
               'NOC', # Number of own children
               'HUPAC', # Presence and age of children
               'HUPAOC', # Presence and age of own children
               'HUPARC', # Presence and age of related children
               "MSP", #Married, spouse present/absent (B, 1, 2, 3, 4, 5, 6)
               
               # Employment Variables
              'ESR', #
              'WKHP', #
               
               # Disability variables
               "DIS",   # Disability status (general)
               "DEAR",  # Hearing difficulty Boolean
               "DEYE",  # Vision difficulty Boolean
               "DOUT",  # Independent living difficulty (B, 1, 2)
               "DREM",  # Cognitive difficulty (B, 1, 2)
               "DPHY",  # Difficulty walking, climbing stairs OR ambulatory difficulty (B, 1, 2)
               "DDRS",   # Self care difficulty (B, 1, 2)
               "DRATX",  # Veterans service connected disability rating (B, 1, 2)
               "HINS3",  # Medicare for 65 or older or people with certain disabilities Boolean
               "HINS4"  # Medicaid for low income or disability Boolean
               
            )

# Terminal year in the vintage
pums.vintage = 2023

# Survey from which the data is requested
survey = "acs5"
```

# 1.0 Load Data

# 1.1 Load PUMS Data from {tidycensus}

```{r}
# Load PUMA to RPA Crosswalk
# For this analysis we will group Martha's Vineyard Commission, Nantucket Planning and Economic Development Commission and Cape Cod Commission into one group -- "CCC" -- which will refer to the Cape and Islands.
puma.rpa_xwalk <- read.csv(here(paste("urbansim_model_update/inputs/puma2020_rpa_crosswalk.csv"))) |> 
  mutate(PUMACE20 = as.character(str_pad(PUMACE20, width = 5, side = "left", pad = 0)))

# Check we have 11 RPAs int the crosswalk
# NPEDC and MVC are cannot be separated from CCC since they share PUMA 1301.
print(length(unique(puma.rpa_xwalk$rpa_acr)))

# Load PUMS data
# Vintage: 2019-2023
# Variables: Described above
# Survey: ACS 5-year PUMS
# Geography: Massachusetts
pums_test <- pums_query(var.list = pums.vars, yr = 2022, srvy = "acs1")

pums <- pums_query(var.list = pums.vars, yr = pums.vintage, srvy = survey)

```

# 1.2 Clean PUMS Data

```{r}
pums_cleaned <- pums |> 
  # Filter for households
  filter(TYPEHUGQ == "1" & SPORDER == "1") |> 
  # Join RPA acronyms to PUMS dataframe by PUMA
  left_join(puma.rpa_xwalk, by = c("PUMA" = "PUMACE20")) |> 
  # Clean data and create new variables we need for analysis 
  mutate(
    # Generate five-year age groupings to match UMDI population projections data.
    ageCAT6 = cut(AGEP, breaks = c(-Inf, seq(4, 84, 5), Inf), labels = 1:18),
    # Adjust Household Income variables to 2023 USD
    HINCP_adj = HINCP*as.numeric(ADJINC),
    # Urbansim Household Income Groupings
    income.1 = case_when(
      HINCP_adj <= 35000 ~ "<=$35,000",
      HINCP_adj > 35000 & HINCP_adj <= 75000 ~ "$35,001-$75,000",
      HINCP_adj > 75000 & HINCP_adj <= 125000 ~ "$75,001-$125,000",
      HINCP_adj > 125000 & HINCP_adj <= 225000 ~ "$125,001-$225,000",
      HINCP_adj > 225000 ~ ">$225,000",
      .default = "Check"
    ),
    # Census Household Income Groupings Type 1
    income.2 = case_when(
      HINCP_adj < 20000 ~ "<$20,000",
      HINCP_adj < 40000 & HINCP_adj >= 20000 ~ "$20,000-$39,999",
      HINCP_adj < 60000 & HINCP_adj >= 40000 ~ "$40,000-$59,999",
      HINCP_adj < 75000 & HINCP_adj >= 60000 ~ "$60,000-$74,999",
      HINCP_adj < 100000 & HINCP_adj >= 75000 ~ "$75,000-$99,999",
      HINCP_adj >= 100000 ~ ">=$100,000",
      .default = "Check"
    ),
    # Census Household Income Groupings Type 2
    income.3 = case_when(
      HINCP_adj < 15000 ~ "<$15,000",
      HINCP_adj < 35000 & HINCP_adj >= 15000 ~ "$15,000-$34,999",
      HINCP_adj < 75000 & HINCP_adj >= 35000 ~ "$35,000-$74,999",
      HINCP_adj < 150000 & HINCP_adj >= 75000 ~ "$75,000-$149,999",
      HINCP_adj >= 150000 ~ ">=$150,000",
      .default = "Check"
    ),
    # Census Household Income Groupings Type 3
    income.4 = case_when(
      HINCP_adj < 10000 ~ "<$10,000",
      HINCP_adj < 15000 & HINCP_adj >= 10000 ~ "$10,000-$14,999",
      HINCP_adj < 20000 & HINCP_adj >= 15000 ~ "$15,000-$19,999",
      HINCP_adj < 25000 & HINCP_adj >= 20000 ~ "$20,000-$24,999",
      HINCP_adj < 30000 & HINCP_adj >= 25000 ~ "$25,000-$29,999",
      HINCP_adj < 35000 & HINCP_adj >= 30000 ~ "$30,000-$34,999",
      HINCP_adj < 40000 & HINCP_adj >= 35000 ~ "$35,000-$39,999",
      HINCP_adj < 45000 & HINCP_adj >= 40000 ~ "$40,000-$44,999",
      HINCP_adj < 50000 & HINCP_adj >= 45000 ~ "$45,000-$49,999",
      HINCP_adj < 60000 & HINCP_adj >= 50000 ~ "$50,000-$59,999",
      HINCP_adj < 75000 & HINCP_adj >= 60000 ~ "$60,000-$74,999",
      HINCP_adj < 100000 & HINCP_adj >= 75000 ~ "$75,000-$99,999",
      HINCP_adj < 125000 & HINCP_adj >= 100000 ~ "$100,000-$124,999",
      HINCP_adj < 150000 & HINCP_adj >= 125000 ~ "$125,000-$149,999",
      HINCP_adj < 200000 & HINCP_adj >= 150000 ~ "$150,000-$199,999",
      HINCP_adj >= 200000 ~ ">=$200,000",
      .default = "Check"
    ),
    # High Income Group Stratification
    income.5 = case_when(
      HINCP_adj < 25000 ~ "<$25,000",
      HINCP_adj < 50000 & HINCP_adj >= 25000 ~ "$25,000-$49,999",
      HINCP_adj < 75000 & HINCP_adj >= 50000 ~ "$50,000-$74,999",
      HINCP_adj < 100000 & HINCP_adj >= 75000 ~ "$75,000-$99,999",
      HINCP_adj < 125000 & HINCP_adj >= 100000 ~ "$100,000-$124,999",
      HINCP_adj < 150000 & HINCP_adj >= 125000 ~ "$125,000-$149,999",
      HINCP_adj < 175000 & HINCP_adj >= 150000 ~ "$150,000-$174,999",
      HINCP_adj < 200000 & HINCP_adj >= 175000 ~ "$175,000-$199,999",
      HINCP_adj < 225000 & HINCP_adj >= 200000 ~ "$200,000-$224,999",
      HINCP_adj < 250000 & HINCP_adj >= 225000 ~ "$225,000-$249,999",
      HINCP_adj < 275000 & HINCP_adj >= 250000 ~ "$250,000-$274,999",
      HINCP_adj < 300000 & HINCP_adj >= 275000 ~ "$275,000-$299,999",
      HINCP_adj < 400000 & HINCP_adj >= 300000 ~ "$300,000-$399,999",
      HINCP_adj < 500000 & HINCP_adj >= 400000 ~ "$400,000-$499,999",
      HINCP_adj >= 500000 ~ ">=$500,000",
      .default = "Check"
    ),
    # Generate ten-year age groupings
    # ageCAT4 is the 2020 decennial census age breaks for age of householder - used for 2020 base year control total reweighter targets
    ageCAT4 = case_when(
      AGEP %in% 0:14 ~ 0,
      AGEP %in% 15:24 ~ 1,
      AGEP %in% 25:34 ~ 2,
      AGEP %in% 35:44 ~ 3,
      AGEP %in% 45:54 ~ 4,
      AGEP %in% 55:59 ~ 5,
      AGEP %in% 60:64 ~ 6,
      AGEP %in% 65:74 ~ 7,
      AGEP %in% 75:84 ~ 8,
      AGEP >= 85 ~ 9,
      .default = 0
    ), 
    # Counts as Worker
    # NOTE: Categories are:
    # (0) under 16 or not in labor force
    # (1) in labor force or non-civilian worker
    worker = case_when(
      # (b) indicates a person under the age of 16.
      # (6) indicates someone not in the laborforce
      (ESR == "b" | ESR == 6) ~ 0,
      # (1) indicates an employed civilian at work.
      # (2) indicates an employed civilian with a job but not at work.
      # (3) indicates civilian who is unemployed.
      # (4) Indicates an employed person in the armed forces. 
      # (5) indicates an employed person in the armed forces with a job but not at work.      
      (ESR == 1 | ESR == 2 | ESR == 3 | ESR == 4 | ESR == 5) ~ 1,
      # When ESR is NA, default to not in the laborforce.
      is.na(ESR) ~ 0,
      # If no information is present default to not in the laborforce.
      .default = 0
    ),
    # Categorizing Civilian Labor Force or not
    # NOTE: Categories are (0) under 16 or non-civilian, (1) in labor force, (2) not in labor force
    lf = case_when(
      ESR == "b" | ESR == 4 | ESR == 5  ~ 0,
      ESR == 1 | ESR == 2 | ESR == 3  ~ 1,
      ESR == 6 ~ 2,
      is.na(ESR) ~ 0,
      .default = 0
    ),
    # Categorizing full or part-time work
    wrkfull = case_when(
      WKHP < 35 ~ "Part time (less than 35 hrs)",
      WKHP >= 35 ~ "Full time (35 hrs or more)",
      is.na(WKHP) ~ "Under 16, unemployed, or not in labor force",
      .default = "Under 16, unemployed, or not in labor force"
    ),
    # Create a grouping variable for types of relationships to the householder.
    # NOTE: 
    # (0) indicates a person in a household that is related in some way to the householder.
    # (1) indicates a person in a household unrelated to the householder.
    RELSHIPP_Group = case_when(
      RELSHIPP %in% c(21,22,23,24,25,26,27,28,29,30,31,32,33) ~ 0,
      RELSHIPP %in% c(34,35,36) ~ 1,
      # Set a default flag for householders.
      .default = 0
    ),
    # Create a grouping variable for owner and renter households.
    TEN_Group = case_when(
      # Owned with a mortage or free and clear -> "Own"
      TEN %in% c("1","2") ~ "Own",
      # Rented or occupied without payment of rent -> "Rent"
      TEN %in% c("3","4") ~ "Rent",
      .default = "Unknown"
    ),
    # Create a "Studio/Non-Defined Bedroom" variable to denote where households do
    # not have a bedroom number.
    studio_flag = case_when(
      BDSP == 0 ~ "Studio/No Defined Bedroom",
      BDSP != 0 ~ "Defined Bedroom",
      .default = "Unknown"
    ),
    # Redefine BDSP to label studios/non-defined bedroom units as One-Bedrooms
    # For purpose of calculating the number of people per bedrooms.
    BDSP = if_else(BDSP == 0, 1, BDSP),
    # Calculate the number of people per bedroom in a housing unit.
    PPB = NP/BDSP,
    # Categorize each housing unit as overcrowded or not based on PPB.
    overcrowded.ppb = case_when(
      PPB > 2.0 ~ "1",
      PPB <= 2.0 ~ "0"
    ),
    # Convert SFN into a numeric column. Introduces NAs for people not in subfamilies
    SFN = as.numeric(SFN),
    # Replace the NA for non-subfamily members with 0.
    SFN = replace_na(SFN, 0),
    # Create a flag for whether a person in a household is a roommate or housemate
    roommate_flag = case_when(
      RELSHIPP == 34 ~ 1,
      RELSHIPP != 34 ~ 0,
      .default = 0
    ),
    # Determine if case counts as child
    child = case_when(
      AGEP < 18 ~ 1,
      AGEP >= 18 ~ 0
      )
  ) |> 
  # Generating Household Aggregations Variables
  group_by(SERIALNO) |> 
  mutate(
    # Total number of workers in a household
    Worker_Total = sum(worker),
    # Total number of children in a household (irrespective of relationship to householder)
    Child_Total = sum(child)
  ) |> 
  ungroup() |> 
  mutate(
    # Categorizing households by presence of workers:
    # (0) no workers
    # (1) one worker
    # (2) two workers
    # (3) three or more workers
    WRKHH = case_when(
      Worker_Total == 0 ~ 0,
      Worker_Total == 1 ~ 1,
      Worker_Total == 2 ~ 2,
      Worker_Total >= 3 ~ 3
      ),
    # Determine what type of household the person is heading or not heading:
    # (1) Householder with children in the household,
    # (2) Non-householder with children in the household
    # (3) Householder with no children in the household
    # (4) Non-householder with no children in the household
    # (5) Household living alone
    HHtype = case_when(
      Child_Total >= 1 & NP > 1 & SPORDER == 1 ~ 1,
      Child_Total >= 1 & NP > 1 & SPORDER != 1 ~ 2,
      Child_Total == 0 & NP > 1 & SPORDER == 1 ~ 3,
      Child_Total == 0 & NP > 1 & SPORDER != 1 ~ 4,
      NP == 1 & SPORDER == 1 ~ 5,
      .default = 99
    )
  )

# Check if we've created any NA values unintentionally.
print(sapply(X = pums_cleaned, FUN =  function(x) sum(is.na(x))))
```

# 2.0 Household Type Analysis
# 2.1 Analyze Additional Household Type Crosstabs
```{r}
# Roommates

# Related Subfamilies

# Grandparents with Children

# Single Parents

# Same-sex Households
```

# 2.2 Understand How Household Types Relate to the Housing Units they Occupy
```{r}
```

# 3.0 Income Distribution Analysis

# 3.1 Distribution Households by Different Income Categories
```{r}
# Standard Urbansim Income Thresholds
income.1 <- pums_cleaned |> 
  group_by(income.1) |> 
  summarise(
    hhds = sum(WGTP)
  ) |> 
  ungroup() |> 
  mutate(
    prop_hhds = round((hhds/sum(hhds))*100, 1)
  )

print(sum(income.1$hhds))

# Census Income Groupings Type 1
income.2 <- pums_cleaned |>
  group_by(income.2) |> 
  summarise(
    hhds = sum(WGTP)
  ) |> 
  ungroup() |> 
  mutate(
    prop_hhds = round((hhds/sum(hhds))*100, 1)
  )

# Census Income Groupings Type 2
income.3 <- pums_cleaned |> 
  group_by(income.3) |> 
  summarise(
    hhds = sum(WGTP)
  ) |> 
  ungroup() |> 
  mutate(
    prop_hhds = round((hhds/sum(hhds))*100, 1)
  )

# Census Income Groupings Type 3
income.4 <- pums_cleaned |>  
  group_by(income.4) |> 
  summarise(
    hhds = sum(WGTP)
  ) |> 
  ungroup() |> 
  mutate(
    prop_hhds = round((hhds/sum(hhds))*100, 1)
  )

# Brandon Thinks Its Important to know more about High Income Households 1\
income.5 <- pums_cleaned |> 
  group_by(income.5) |> 
  summarise(
    hhds = sum(WGTP)
  ) |> 
  ungroup() |> 
  mutate(
    prop_hhds = round((hhds/sum(hhds))*100, 1)
  )

# Commentary: Income Groups
# Urbansim income groupings are relatively well balanced. Each of the five groups accounts for at most 24% and at least 17% of households in MA. The highest and lowest income groups represent 17 and 19 percent of households in MA while the middle income categories each represent 20 percent of households. Would it be useful to tweak the edges of some of the groups to further balance the distribution - attempt to achieve 20 percent in each?

# Census Income groups type 1 show a balance in the low income groupings, about 10 percent in each grouping. Curious why they deviate from the $20k buckets in those middle income categories? More balanced with those $20k groupings (maximum deviation between highest and lowest bucket is 2.0 percentage points v. ~3 percentage points with the 15 and 25k groupings at the top end of the below 100k groups.) Doesn't account for potential vartiation in high household income groups (50 percent of households earn above 100k).

# Distribution balance is poor. Only 8 percent of households earn less than $15k.High income categories not granular enough and carry outsized weight.

# The 5K income groupings from 15K to $50K are decently balanced at ~2.5 percent of total households (with ~70k households in each bin). Below 15K we have nearly 200k households (~10 percent of households).

# $100k-$200k ~ 28.5 percent (100-150 ~ 16 percent, 150-200 ~ 11 percent). $200k-$300K ~ 11 percent. $300+ about ~10 percent of households. 

# Commentary: General
# Do we want to align any of the income groups with federal or state cut offs for income assistance legislated by different programs (SNAP, RAFT, etc.)?

# Income groups by household size? How far can that income actually go? Disposable income? Income spent on housing? 
  
```

# 3.2 Income Group Spatial Distribution (By RPA)

```{r}
# Standard Urbansim Income Thresholds
income.1.geo <- pums_cleaned |> 
  group_by(rpa_acr, income.1) |> 
  summarise(
    hhds = sum(WGTP)
  ) |> 
  ungroup() |> 
  group_by(rpa_acr) |> 
  mutate(
    prop_hhds = round((hhds/sum(hhds))*100, 1)
  ) |> 
  ungroup()

print(sum(income.1$hhds))

# Census Income Groupings Type 1
income.2.geo <- pums_cleaned |>
  group_by(rpa_acr, income.2) |> 
  summarise(
    hhds = sum(WGTP)
  ) |> 
  ungroup() |> 
  group_by(rpa_acr) |> 
  mutate(
    prop_hhds = round((hhds/sum(hhds))*100, 1)
  ) |> 
  ungroup()

# Census Income Groupings Type 2
income.3.geo <- pums_cleaned |> 
  group_by(rpa_acr, income.3) |> 
  summarise(
    hhds = sum(WGTP)
  ) |> 
  ungroup() |> 
  group_by(rpa_acr) |> 
  mutate(
    prop_hhds = round((hhds/sum(hhds))*100, 1)
  ) |> 
  ungroup()

# Census Income Groupings Type 3
income.4.geo <- pums_cleaned |>  
  group_by(rpa_acr, income.4) |> 
  summarise(
    hhds = sum(WGTP)
  ) |> 
  ungroup() |> 
  group_by(rpa_acr) |> 
  mutate(
    prop_hhds = round((hhds/sum(hhds))*100, 1)
  ) |> 
  ungroup()

# Brandon Thinks Its Important to know more about High Income Households 1\
income.5.geo <- pums_cleaned |> 
  group_by(rpa_acr, income.5) |> 
  summarise(
    hhds = sum(WGTP)
  ) |> 
  ungroup() |> 
  group_by(rpa_acr) |> 
  mutate(
    prop_hhds = round((hhds/sum(hhds))*100, 1)
  ) |> 
  ungroup()

# Commentary
```

# 3.3 Analysis of Income and Other Household Variables

```{r}
# By Household Type
```

# 4.0 Older Household Analysis

```{r}
#Define Older Adults - two types of variables - more than 60 and more than 65 years old 
pums_cleaned <- pums_test |> 
  # Join RPA acronyms to PUMS dataframe by PUMA
  left_join(puma.rpa_xwalk, by = c("PUMA" = "PUMACE20")) |> 
  # Clean data and create new variables we need for analysis 
  mutate(
    older60 = case_when(
     (AGEP >= 60) ~ 1,
      # If AGEP is NA, default to not being older60.
      is.na(AGEP) ~ 0,
      # If under 60 years of age then default to 0.
      .default = 0
    ),

    older65 = case_when(
     (AGEP >= 65) ~ 1,
      # If AGEP is NA, default to not being older65.
      is.na(AGEP) ~ 0,
      # If under 65 years of age then default to 0.
      .default = 0
    )
  )

```


```{r}
#Filter Older Adults with disability
pums_disold60 <- filter(pums_cleaned, DIS=="1" & older60==1)
pums_disold65 <- filter(pums_cleaned, DIS=="1" & older65==1)

print(summary(pums_disold60$PINCP))
print(summary(pums_disold65$PINCP))

print (older60_dis_count/older60_count) #% of older adults that are disabled
print (older65_dis_count/older65_count) #% of older adults that are disabled



```
```{r}
#In 2022, the federal poverty level for a single person was $13,590, and for a family of four it was $29,678 - BLS

bpl_count <- count(filter(pums_cleaned, PINCP<= 13590))
older60_count <- count(filter(pums_cleaned, older60==1))
older65_count <- count(filter(pums_cleaned, older65==1))

older60_dis_count <- count(pums_disold60)
older65_dis_count <- count(pums_disold65)

bpl_older60_count <-count(filter(pums_cleaned,older60==1 & PINCP <= 13590))
bpl_older65_count <-count(filter(pums_cleaned,older65==1 & PINCP <= 13590))

bpl_older60_dis_count <-count(filter(pums_disold60, PINCP <= 13590))
bpl_older65_dis_count <-count(filter(pums_disold65, PINCP <= 13590))

print (bpl_older60_count/older60_count) #% of older adults that are bpl
print (bpl_older65_count/older65_count) #% of older adults that are bpl

print(bpl_older60_dis_count/older60_count) #% of older adults that are bpl and disabled
print(bpl_older65_dis_count/older65_count) #% of older adults that are bpl and disabled

```
```{r}
# Older adults (60+ and 65+) receiving Social Security income (SSP > 0)

older60_ssp_count <- count(filter(pums_cleaned, older60 == 1 & !is.na(SSP) & SSP > 0))
older65_ssp_count <- count(filter(pums_cleaned, older65 == 1 & !is.na(SSP) & SSP > 0))

# % of older adults receiving Social Security income
print(older60_ssp_count / older60_count)
print(older65_ssp_count / older65_count)

#Older adults (60+ and 65+) receiving Retirement income (RETP > 0)

older60_retp_count <- count(filter(pums_cleaned, older60 == 1 & !is.na(RETP) & RETP > 0))
older65_retp_count <- count(filter(pums_cleaned, older65 == 1 & !is.na(RETP) & RETP > 0))

print(older60_retp_count / older60_count) # % of adults 60+ with Retirement income
print(older65_retp_count / older65_count) # % of adults 65+ with Retirement income

```
```{r}
# Older adults living in group quarters vs households

# Count of institutional and noninstitutional group quarters residents
older60_gq_count <- count(filter(pums_cleaned, older60 == 1 & TYPEHUGQ != 1))
older65_gq_count <- count(filter(pums_cleaned, older65 == 1 & TYPEHUGQ != 1))

# Count of household residents
older60_household_count <- count(filter(pums_cleaned, older60 == 1 & TYPEHUGQ == 1))
older65_household_count <- count(filter(pums_cleaned, older65 == 1 & TYPEHUGQ == 1))

# Print proportions
print(older60_gq_count / older60_count)        # % of adults 60+ living in group quarters
print(older65_gq_count / older65_count)        # % of adults 65+ living in group quarters

print(older60_household_count / older60_count) # % of adults 60+ living in households
print(older65_household_count / older65_count) # % of adults 65+ living in households

```

# 5.0 Household Size Analysis

```{r}
```

# 6.0 Overcrowding Analysis

```{r}
```

# 7.0 Laborforce Participation Rate: PUMS v. CPS

```{r}
#ACS 2016-2020 5-YEAR PUMS Testing

#Labor Force Comparison Data (from BLS/FRED)

lfpr_PUMS <- pums_DATA %>%
  filter(TYPE == 1 & lf >= 1) %>%
  group_by(year) %>%
  count(var = lf, wt = PWGTP) %>%
  mutate(
    var = case_when(
      var == 1 ~ "in LF",
      var == 2 ~ "not in LF"),
    freq = (n/sum(n))*100) %>%
  filter(var == "in LF")

fredr::fredr_set_key("8bdd1423b21c9e0bc762d1d259e7fecc")

lfpr_FRED <- fredr::fredr(
  series_id = "LBSSA25",
  observation_start = as.Date("1994-01-01"),
  observation_end = as.Date("2022-03-01")
)


#Not too happy with this. I trust the BLS data more so than the PUMS data, especially since the start of the pandemic.
#Not a fan of the rate of change over the 2012-2019 period in PUMS. LFPR was not decreasing. It was growing substantially and levels were commensurate with pre-GR numbers.

#Pull in CPS data to create labor force participation rates for the state by age sex educational attainment to produce adjustment factors for PUMS LFPR data.
library(cpsR)

cps_years_1 <- 1994:2021
cps_years_2 <- 1994:2022

cps_varlist <- c("pwsswgt", "pwcmpwgt", "pemlr", "peeduca", "pesex", "prtage", "gestfips")

CPS_data_Jan <- map_dfr(cps_years_2, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 1
  )
},  .id = "year")

CPS_data_Feb <- map_dfr(cps_years_2, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 2
  )
},  .id = "year")

cps_mar_years_1 <- 1994:2018
cps_mar_years_2 <- 2019:2022

CPS_data_Mar_1 <- map_dfr(cps_mar_years_1, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 3
  )
},  .id = "year")

CPS_data_Mar_2 <- map_dfr(cps_mar_years_2, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 3
  )
},  .id = "year")

CPS_data_Apr <- map_dfr(cps_years_1, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 4
  )
},  .id = "year")

CPS_data_May <- map_dfr(cps_years_1, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 5
  )
},  .id = "year")

CPS_data_Jun <- map_dfr(cps_years_1, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 6
  )
},  .id = "year")

CPS_data_Jul <- map_dfr(cps_years_1, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 7
  )
},  .id = "year")

CPS_data_Aug <- map_dfr(cps_years_1, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 8
  )
},  .id = "year")

cps_sep_years_1 <- 1994:2015
cps_sep_years_2 <- 2016:2021

CPS_data_Sep_1 <- map_dfr(cps_sep_years_1, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 9
  )
},  .id = "year")

CPS_data_Sep_2 <- map_dfr(cps_sep_years_2, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 9
  )
},  .id = "year")

CPS_data_Oct <- map_dfr(cps_years_1, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 10
  )
},  .id = "year")

CPS_data_Nov <- map_dfr(cps_years_1, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 11
  )
},  .id = "year")

CPS_data_Dec <- map_dfr(cps_years_1, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 12
  )
},  .id = "year")

CPS_data_Jan <- CPS_data_Jan %>%
  mutate(month = "01") %>%
  filter(gestfips == "25")

CPS_data_Feb <- CPS_data_Feb %>%
  mutate(month = "02") %>%
  filter(gestfips == "25")

CPS_data_Mar_1 <- CPS_data_Mar_1 %>%
  mutate(month = "03",
         pemlr = as.character(pemlr)) %>%
  filter(gestfips == "25")

CPS_data_Mar_2 <- CPS_data_Mar_2 %>%
  mutate(month = "03") %>%
  filter(gestfips == "25")

CPS_data_Apr <- CPS_data_Apr %>%
  mutate(month = "04") %>%
  filter(gestfips == "25")

CPS_data_May <- CPS_data_May %>%
  mutate(month = "05") %>%
  filter(gestfips == "25")

CPS_data_Jun <- CPS_data_Jun %>%
  mutate(month = "06") %>%
  filter(gestfips == "25")

CPS_data_Jul <- CPS_data_Jul %>%
  mutate(month = "07") %>%
  filter(gestfips == "25")

CPS_data_Aug <- CPS_data_Aug %>%
  mutate(month = "08") %>%
  filter(gestfips == "25")

CPS_data_Sep_1 <- CPS_data_Sep_1 %>%
  mutate(month = "09",
         pemlr = as.character(pemlr)) %>%
  filter(gestfips == "25")

CPS_data_Sep_2 <- CPS_data_Sep_2 %>%
  mutate(month = "09") %>%
  filter(gestfips == "25")

CPS_data_Oct <- CPS_data_Oct %>%
  mutate(month = "10") %>%
  filter(gestfips == "25")

CPS_data_Nov <- CPS_data_Nov %>%
  mutate(month = "11") %>%
  filter(gestfips == "25")

CPS_data_Dec <- CPS_data_Dec %>%
  mutate(month = "12") %>%
  filter(gestfips == "25")

CPS_data_Mar_2 <- CPS_data_Mar_2 %>%
  mutate(
    year = case_when(
      year == "1" ~ "26",
      year == "2" ~ "27",
      year == "3" ~ "28",
      year == "4" ~ "29"
    )
  )

CPS_data_Sep_2 <- CPS_data_Sep_2 %>%
  mutate(
    year = case_when(
      year == "1" ~ "24",
      year == "2" ~ "25",
      year == "3" ~ "26",
      year == "4" ~ "27",
      year == "5" ~ "28",
      year == "6" ~ "29"
    )
  )

CPS_data <- rbind(CPS_data_Jan, CPS_data_Feb, CPS_data_Mar_1, CPS_data_Mar_2, CPS_data_Apr, CPS_data_May, CPS_data_Jun,
                  CPS_data_Jul, CPS_data_Aug, CPS_data_Sep_1, CPS_data_Sep_2, CPS_data_Oct, CPS_data_Nov, CPS_data_Dec)

CPS_data <- CPS_data %>%
  mutate(
    year = case_when(
      year == "1" ~ 1994,
      year == "2" ~ 1995,
      year == "3" ~ 1996,
      year == "4" ~ 1997,
      year == "5" ~ 1998,
      year == "6" ~ 1999,
      year == "7" ~ 2000,
      year == "8" ~ 2001,
      year == "9" ~ 2002,
      year == "10" ~ 2003,
      year == "11" ~ 2004,
      year == "12" ~ 2005,
      year == "13" ~ 2006,
      year == "14" ~ 2007,
      year == "15" ~ 2008,
      year == "16" ~ 2009,
      year == "17" ~ 2010,
      year == "18" ~ 2011,
      year == "19" ~ 2012,
      year == "20" ~ 2013,
      year == "21" ~ 2014,
      year == "22" ~ 2015,
      year == "23" ~ 2016,
      year == "24" ~ 2017,
      year == "25" ~ 2018,
      year == "26" ~ 2019,
      year == "27" ~ 2020,
      year == "28" ~ 2021,
      year == "29" ~ 2022),
    educ_attn = case_when(
      peeduca <= 39 ~ 1,
      peeduca %in% 40:42 ~ 2,
      peeduca == 43 ~ 3,
      peeduca >= 44 ~ 4,
      peeduca == -1 ~ 1,
      is.na(peeduca) ~ 1),
    PAGEC =
      cut(prtage,
          breaks = c(-Inf, seq(4, 84, 5), Inf),
          labels = 1:18),
    lf = case_when(
      pemlr == 1 | pemlr == 2 | pemlr == 4 ~ 1,
      pemlr == 3 | pemlr == 5 | pemlr == 6 | pemlr == 7 ~ 2,
      pemlr == -1 ~ 2,
      is.na(pemlr) ~ 2
    )
  )

lfpr_CPS <- CPS_data %>%
  filter(lf >= 1 & prtage > 16) %>%
  mutate(date_1 = paste(year, month, "01", sep = "-")) %>%
  group_by(date_1) %>%
  count(var = lf, wt = pwcmpwgt) %>%
  mutate(
    var = case_when(
      var == 1 ~ "in LF",
      var == 2 ~ "not in LF"),
    freq = (n/sum(n))*100) %>%
  filter(var == "in LF")

##
lfpr_FRED_post09 <- lfpr_FRED[186:314,]
##

lfpr_FRED %>%
  ggplot(aes(x=date, y=value, color = "lfpr_MA")) +
  geom_line(lwd = 1.0) +
  geom_line(data = lfpr_PUMS,
            mapping = aes(x=as.Date(year, format = "%Y"), y = freq, color = "lfpr_PUMS"),
            lwd = 1.0) +
  geom_line(data = lfpr_CPS,
            mapping = aes(x = as.Date(date_1, format = "%Y-%m-%d"), y = freq, color = "lfpr_CPS"),
            lwd = 1.0) +
  geom_hline(yintercept = max(lfpr_FRED_post09$value), color = "blue", linetype = 2) +
  geom_hline(yintercept = min(lfpr_FRED_post09$value), color = "blue", linetype = 2) +
  scale_color_manual(name = "Data Series",
                     breaks = c("lfpr_FRED", "lfpr_PUMS", "lfpr_CPS"),
                     values = c("lfpr_FRED" = "black", "lfpr_PUMS" = "red", "lfpr_CPS" = "green")) +
  labs(title = "") +
  ylab("Percent (%)") +
  xlab("Year") +
  theme(legend.title = element_text(size = 16),
        legend.text = element_text(size = 12),
        panel.background = element_blank(),
        panel.border = element_rect(color = "black",  fill = NA, size = 1),
        panel.grid = element_line(color = "gray", size = 0.5),
        axis.title.x = element_text(color = "black", size = 16),
        axis.title.y = element_text(color = "black", size = 16),
        axis.text.x = element_text(color = "black", size = 14),
        axis.text.y = element_text(color = "black", size = 14))

```
