---
title: "Urbansim Model Update Exploratory Analysis"
format: html
editor: visual
---
# 0.0 Environment and Script Set-up
# 0.1 Load Libraries and Functions
```{r}
# {here} library to initiate conditional filepaths
library(here)
# Libraries used in Urbansim/Projections workflow
source(here("urbansim_model_update/projections_libraries.R"))
# Functions created for the Urbansim/Projections workflow
source(here("urbansim_model_update/projections_functions.R"))
```

# 0.2 Set Environment Variables for this Script
```{r}
# Set random seed
set.seed(351)

# Remove scientific notation
options(scipen = 999)

# Set the PUMS API call variables for this analysis. Used in the 
# Variables from the PUMS data to query (data dictionary is here: https://www.census.gov/programs-surveys/acs/microdata/documentation.2023.html#list-tab-1370939201)
pums.vars <- c(
               # General PUMS variables
               'PUMA', #PUMA designation
               'ADJINC', # Inflation adjustment factor for income variables
               'TYPEHUGQ', # Type of household (Household or group quarters)
  
               #Demographic variables
               'AGEP', # Age of person
               "SEX", # Sex
               'NP', # Number of persons in a household
               'TEN', # Tenure
               'RAC1P', # Race
               'HISP', # Recoded detailed Hispanic origin (Ethnicity)
               'R18', #Presence of persons under 18 years in household
               'R60', # Presence of persons over 60 years in household
               'R65', # Presence of persons over 65 years in household
  
               # Income variables
               "FINCP", # Family Income
               "HINCP", # Household Income
               "INTP", # Interest, dividends, and net rental income
               "PINCP", # Total person's income
               "WAGP", # Wages or salary income past 12 months
               "SSIP", # Supplementary Security Income
               "SSP", # Social Security income
               "SEMP", # Self employment income
               "RETP", # Retirement income
               "PAP", # Public assistance income
               "OIP", #All other income
               
               # Overcrowding and Household Composition variables
               'BDSP', # Number of bedrooms
               'RMSP', # Number of rooms
               'CPLT', # Couple Type
               'RELSHIPP', # Relationship to the householder
               'PSF', # Presence of subfamily
               'SFN', # Subfamily number
               'SFR', # Subfamily relationship
               'NR', # Presence of non-relative in household
               'NRC', # Number of related children in a household
               'MULTG', # Multigenerational household flag
               'HHT', # Household/family type
               'HHT2', # Household family type (including cohabiting)
               "PARTNER", # Unmarried partner household
               'NPP', # Grandparent headed household with no parent present
               'NPF', # Number of persons in family
               'HUGCL', # Household with grandparent living with grandchildren
               'NOC', # Number of own children
               'HUPAC', # Presence and age of children
               'HUPAOC', # Presence and age of own children
               'HUPARC' # Presence and age of related children
               
               # Disability variables
                    
            )

# Terminal year in the vintage
pums.vintage = 2023

# Survey from which the data is requested
survey = "acs5"
```
# 1.0 Load Data
# 1.1 Load PUMS Data from {tidycensus}
```{r}
# Load PUMA to RPA Crosswalk
# For this analysis we will group Martha's Vineyard Commission, Nantucket Planning and Economic Development Commission and Cape Cod Commission into one group -- "CCC" -- which will refer to the Cape and Islands.
puma.rpa_xwalk <- read.csv(here(paste("urbansim_model_update/inputs/puma2020_rpa_crosswalk.csv"))) |> 
  mutate(PUMACE20 = as.character(str_pad(PUMACE20, width = 5, side = "left", pad = 0)))

# Check we have 11 RPAs int the crosswalk
# NPEDC and MVC are cannot be separated from CCC since they share PUMA 1301.
print(length(unique(puma.rpa_xwalk$rpa_acr)))

# Load PUMS data
# Vintage: 2019-2023
# Variables: Described above
# Survey: ACS 5-year PUMS
# Geography: Massachusetts
pums <- pums_query(var.list = pums.vars, yr = pums.vintage, srvy = survey) 

```

# 1.2 Clean PUMS Data
```{r}
pums_cleaned <- pums |> 
  # Filter for households
  filter(TYPEHUGQ == "1" & SPORDER == "1") |> 
  # Join RPA acronyms to PUMS dataframe by PUMA
  left_join(puma.rpa_xwalk, by = c("PUMA" = "PUMACE20")) |> 
  # Clean data and create new variables we need for analysis 
  mutate(
    # Generate five-year age groupings to match UMDI population projections data.
    ageCAT6 = cut(AGEP, breaks = c(-Inf, seq(4, 84, 5), Inf), labels = 1:18),
    # Adjust Household Income variables to 2023 USD
    HINCP_adj = HINCP*as.numeric(ADJINC),
    # Generate ten-year age groupings
    # ageCAT4 is the 2020 decennial census age breaks for age of householder - used for 2020 base year control total reweighter targets
    ageCAT4 = case_when(
      AGEP %in% 0:14 ~ 0,
      AGEP %in% 15:24 ~ 1,
      AGEP %in% 25:34 ~ 2,
      AGEP %in% 35:44 ~ 3,
      AGEP %in% 45:54 ~ 4,
      AGEP %in% 55:59 ~ 5,
      AGEP %in% 60:64 ~ 6,
      AGEP %in% 65:74 ~ 7,
      AGEP %in% 75:84 ~ 8,
      AGEP >= 85 ~ 9,
      .default = 0
    ), 
    # Counts as Worker
    # NOTE: Categories are:
    # (0) under 16 or not in labor force
    # (1) in labor force or non-civilian worker
    worker = case_when(
      # (b) indicates a person under the age of 16.
      # (6) indicates someone not in the laborforce
      (ESR == "b" | ESR == 6) ~ 0,
      # (1) indicates an employed civilian at work.
      # (2) indicates an employed civilian with a job but not at work.
      # (3) indicates civilian who is unemployed.
      # (4) Indicates an employed person in the armed forces. 
      # (5) indicates an employed person in the armed forces with a job but not at work.      
      (ESR == 1 | ESR == 2 | ESR == 3 | ESR == 4 | ESR == 5) ~ 1,
      # When ESR is NA, default to not in the laborforce.
      is.na(ESR) ~ 0,
      # If no information is present default to not in the laborforce.
      .default = 0
    ),
    # Create a grouping variable for types of relationships to the householder.
    # NOTE: 
    # (0) indicates a person in a household that is related in some way to the householder.
    # (1) indicates a person in a household unrelated to the householder.
    RELSHIPP_Group = case_when(
      RELSHIPP %in% c(21,22,23,24,25,26,27,28,29,30,31,32,33) ~ 0,
      RELSHIPP %in% c(34,35,36) ~ 1,
      # Set a default flag for householders.
      .default = 0
    ),
    # Create a grouping variable for owner and renter households.
    TEN_Group = case_when(
      # Owned with a mortage or free and clear -> "Own"
      TEN %in% c("1","2") ~ "Own",
      # Rented or occupied without payment of rent -> "Rent"
      TEN %in% c("3","4") ~ "Rent",
      .default = "Unknown"
    ),
    # Create a "Studio/Non-Defined Bedroom" variable to denote where households do
    # not have a bedroom number.
    studio_flag = case_when(
      BDSP == 0 ~ "Studio/No Defined Bedroom",
      BDSP != 0 ~ "Defined Bedroom",
      .default = "Unknown"
    ),
    # Redefine BDSP to label studios/non-defined bedroom units as One-Bedrooms
    # For purpose of calculating the number of people per bedrooms.
    BDSP = if_else(BDSP == 0, 1, BDSP),
    # Calculate the number of people per bedroom in a housing unit.
    PPB = NP/BDSP,
    # Categorize each housing unit as overcrowded or not based on PPB.
    overcrowded.ppb = case_when(
      PPB > 2.0 ~ "1",
      PPB <= 2.0 ~ "0"
    ),
    # Convert SFN into a numeric column. Introduces NAs for people not in subfamilies
    SFN = as.numeric(SFN),
    # Replace the NA for non-subfamily members with 0.
    SFN = replace_na(SFN, 0),
    # Create a flag for whether a person in a household is a roommate or housemate
    roommate_flag = case_when(
      RELSHIPP == 34 ~ 1,
      RELSHIPP != 34 ~ 0,
      .default = 0
    )
  )


# Check if we've created any NA values unintentionally.
print(sapply(X = pums_cleaned, FUN =  function(x) sum(is.na(x))))
```

# 2.0 Household Type Analysis
```{r}
```

# 3.0 Income Distribution Analysis
# 3.1 Evaluation of Census Income Thresholds
```{r}
# Standard Urbansim Income Thresholds
income.1 <- pums_cleaned |>
  mutate(
    income.1 = case_when(
      HINCP_adj <= 35000 ~ "<=$35,000",
      HINCP_adj > 35000 & HINCP_adj <= 75000 ~ "$35,001-$75,000",
      HINCP_adj > 75000 & HINCP_adj <= 125000 ~ "$75,001-$125,000",
      HINCP_adj > 125000 & HINCP_adj <= 225000 ~ "$125,001-$225,000",
      HINCP_adj > 225000 ~ ">$225,000",
      .default = "Check"
    )
  ) |> 
  group_by(income.1) |> 
  summarise(
    hhds = sum(WGTP)
  ) |> 
  ungroup() |> 
  mutate(
    prop_hhds = round((hhds/sum(hhds))*100, 1)
  )

print(sum(income.1$hhds))

# Census Income Groupings Type 1
income.2 <- pums_cleaned |> 
  mutate(
    income.2 = case_when(
      HINCP_adj < 20000 ~ "<$20,000",
      HINCP_adj < 40000 & HINCP_adj >= 20000 ~ "$20,000-$39,999",
      HINCP_adj < 60000 & HINCP_adj >= 40000 ~ "$40,000-$59,999",
      HINCP_adj < 75000 & HINCP_adj >= 60000 ~ "$60,000-$74,999",
      HINCP_adj < 100000 & HINCP_adj >= 75000 ~ "$75,000-$99,999",
      HINCP_adj >= 100000 ~ ">=$100,000",
      .default = "Check"
    )
  ) |> 
  group_by(income.2) |> 
  summarise(
    hhds = sum(WGTP)
  ) |> 
  ungroup() |> 
  mutate(
    prop_hhds = round((hhds/sum(hhds))*100, 1)
  )

# Census Income Groupings Type 2
income.3 <- pums_cleaned |>
  mutate(
    income.3 = case_when(
      HINCP_adj < 15000 ~ "<$15,000",
      HINCP_adj < 35000 & HINCP_adj >= 15000 ~ "$15,000-$34,999",
      HINCP_adj < 75000 & HINCP_adj >= 35000 ~ "$35,000-$74,999",
      HINCP_adj < 150000 & HINCP_adj >= 75000 ~ "$75,000-$149,999",
      HINCP_adj >= 150000 ~ ">=$150,000",
      .default = "Check"
    )
  ) |> 
  group_by(income.3) |> 
  summarise(
    hhds = sum(WGTP)
  ) |> 
  ungroup() |> 
  mutate(
    prop_hhds = round((hhds/sum(hhds))*100, 1)
  )

# Census Income Groupings Type 3
income.4 <- pums_cleaned |> 
  mutate(
    income.4 = case_when(
      HINCP_adj < 10000 ~ "<$10,000",
      HINCP_adj < 15000 & HINCP_adj >= 10000 ~ "$10,000-$14,999",
      HINCP_adj < 20000 & HINCP_adj >= 15000 ~ "$15,000-$19,999",
      HINCP_adj < 25000 & HINCP_adj >= 20000 ~ "$20,000-$24,999",
      HINCP_adj < 30000 & HINCP_adj >= 25000 ~ "$25,000-$29,999",
      HINCP_adj < 35000 & HINCP_adj >= 30000 ~ "$30,000-$34,999",
      HINCP_adj < 40000 & HINCP_adj >= 35000 ~ "$35,000-$39,999",
      HINCP_adj < 45000 & HINCP_adj >= 40000 ~ "$40,000-$44,999",
      HINCP_adj < 50000 & HINCP_adj >= 45000 ~ "$45,000-$49,999",
      HINCP_adj < 60000 & HINCP_adj >= 50000 ~ "$50,000-$59,999",
      HINCP_adj < 75000 & HINCP_adj >= 60000 ~ "$60,000-$74,999",
      HINCP_adj < 100000 & HINCP_adj >= 75000 ~ "$75,000-$99,999",
      HINCP_adj < 125000 & HINCP_adj >= 100000 ~ "$100,000-$124,999",
      HINCP_adj < 150000 & HINCP_adj >= 125000 ~ "$125,000-$149,999",
      HINCP_adj < 200000 & HINCP_adj >= 150000 ~ "$150,000-$199,999",
      HINCP_adj >= 200000 ~ ">=$200,000",
      .default = "Check"
    )
  ) |> 
  group_by(income.4) |> 
  summarise(
    hhds = sum(WGTP)
  ) |> 
  ungroup() |> 
  mutate(
    prop_hhds = round((hhds/sum(hhds))*100, 1)
  )

# Brandon Thinks Its Important to know more about High Income Households 1\
income.5 <- pums_cleaned |> 
  mutate(
    income.5 = case_when(
      HINCP_adj < 25000 ~ "<$25,000",
      HINCP_adj < 50000 & HINCP_adj >= 25000 ~ "$25,000-$49,999",
      HINCP_adj < 75000 & HINCP_adj >= 50000 ~ "$50,000-$74,999",
      HINCP_adj < 100000 & HINCP_adj >= 75000 ~ "$75,000-$99,999",
      HINCP_adj < 125000 & HINCP_adj >= 100000 ~ "$100,000-$124,999",
      HINCP_adj < 150000 & HINCP_adj >= 125000 ~ "$125,000-$149,999",
      HINCP_adj < 175000 & HINCP_adj >= 150000 ~ "$150,000-$174,999",
      HINCP_adj < 200000 & HINCP_adj >= 175000 ~ "$175,000-$199,999",
      HINCP_adj < 225000 & HINCP_adj >= 200000 ~ "$200,000-$224,999",
      HINCP_adj < 250000 & HINCP_adj >= 225000 ~ "$225,000-$249,999",
      HINCP_adj < 275000 & HINCP_adj >= 250000 ~ "$250,000-$274,999",
      HINCP_adj < 300000 & HINCP_adj >= 275000 ~ "$275,000-$299,999",
      HINCP_adj < 400000 & HINCP_adj >= 300000 ~ "$300,000-$399,999",
      HINCP_adj < 500000 & HINCP_adj >= 400000 ~ "$400,000-$499,999",
      HINCP_adj >= 500000 ~ ">=$500,000",
      .default = "Check"
    )
  ) |> 
  group_by(income.5) |> 
  summarise(
    hhds = sum(WGTP)
  ) |> 
  ungroup() |> 
  mutate(
    prop_hhds = round((hhds/sum(hhds))*100, 1)
  )

# Commentary
  
```
# 3.2 Income Group Spatial Distribution (By RPA)
```{r}
```

# 3.3 Analysis of Income and Other Household Variables
```{r}
```

# 4.0  Older Household Analysis
```{r}
```

# 5.0 Household Size Analysis
```{r}
```

# 6.0 Overcrowding Analysis
```{r}
```

# 7.0 Laborforce Participation Rate: PUMS v. CPS
```{r}
#ACS 2016-2020 5-YEAR PUMS Testing

#Labor Force Comparison Data (from BLS/FRED)

lfpr_PUMS <- PUMS_data %>%
  filter(TYPE == 1 & lf >= 1) %>%
  group_by(year) %>%
  count(var = lf, wt = PWGTP) %>%
  mutate(
    var = case_when(
      var == 1 ~ "in LF",
      var == 2 ~ "not in LF"),
    freq = (n/sum(n))*100) %>%
  filter(var == "in LF")

fredr::fredr_set_key("8bdd1423b21c9e0bc762d1d259e7fecc")

lfpr_FRED <- fredr::fredr(
  series_id = "LBSSA25",
  observation_start = as.Date("1994-01-01"),
  observation_end = as.Date("2022-03-01")
)


#Not too happy with this. I trust the BLS data more so than the PUMS data, especially since the start of the pandemic.
#Not a fan of the rate of change over the 2012-2019 period in PUMS. LFPR was not decreasing. It was growing substantially and levels were commensurate with pre-GR numbers.

#Pull in CPS data to create labor force participation rates for the state by age sex educational attainment to produce adjustment factors for PUMS LFPR data.
library(cpsR)

cps_years_1 <- 1994:2021
cps_years_2 <- 1994:2022

cps_varlist <- c("pwsswgt", "pwcmpwgt", "pemlr", "peeduca", "pesex", "prtage", "gestfips")

CPS_data_Jan <- map_dfr(cps_years_2, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 1
  )
},  .id = "year")

CPS_data_Feb <- map_dfr(cps_years_2, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 2
  )
},  .id = "year")

cps_mar_years_1 <- 1994:2018
cps_mar_years_2 <- 2019:2022

CPS_data_Mar_1 <- map_dfr(cps_mar_years_1, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 3
  )
},  .id = "year")

CPS_data_Mar_2 <- map_dfr(cps_mar_years_2, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 3
  )
},  .id = "year")

CPS_data_Apr <- map_dfr(cps_years_1, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 4
  )
},  .id = "year")

CPS_data_May <- map_dfr(cps_years_1, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 5
  )
},  .id = "year")

CPS_data_Jun <- map_dfr(cps_years_1, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 6
  )
},  .id = "year")

CPS_data_Jul <- map_dfr(cps_years_1, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 7
  )
},  .id = "year")

CPS_data_Aug <- map_dfr(cps_years_1, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 8
  )
},  .id = "year")

cps_sep_years_1 <- 1994:2015
cps_sep_years_2 <- 2016:2021

CPS_data_Sep_1 <- map_dfr(cps_sep_years_1, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 9
  )
},  .id = "year")

CPS_data_Sep_2 <- map_dfr(cps_sep_years_2, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 9
  )
},  .id = "year")

CPS_data_Oct <- map_dfr(cps_years_1, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 10
  )
},  .id = "year")

CPS_data_Nov <- map_dfr(cps_years_1, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 11
  )
},  .id = "year")

CPS_data_Dec <- map_dfr(cps_years_1, ~{
  get_basic(
    vars = cps_varlist,
    year = .x,
    month = 12
  )
},  .id = "year")

CPS_data_Jan <- CPS_data_Jan %>%
  mutate(month = "01") %>%
  filter(gestfips == "25")

CPS_data_Feb <- CPS_data_Feb %>%
  mutate(month = "02") %>%
  filter(gestfips == "25")

CPS_data_Mar_1 <- CPS_data_Mar_1 %>%
  mutate(month = "03",
         pemlr = as.character(pemlr)) %>%
  filter(gestfips == "25")

CPS_data_Mar_2 <- CPS_data_Mar_2 %>%
  mutate(month = "03") %>%
  filter(gestfips == "25")

CPS_data_Apr <- CPS_data_Apr %>%
  mutate(month = "04") %>%
  filter(gestfips == "25")

CPS_data_May <- CPS_data_May %>%
  mutate(month = "05") %>%
  filter(gestfips == "25")

CPS_data_Jun <- CPS_data_Jun %>%
  mutate(month = "06") %>%
  filter(gestfips == "25")

CPS_data_Jul <- CPS_data_Jul %>%
  mutate(month = "07") %>%
  filter(gestfips == "25")

CPS_data_Aug <- CPS_data_Aug %>%
  mutate(month = "08") %>%
  filter(gestfips == "25")

CPS_data_Sep_1 <- CPS_data_Sep_1 %>%
  mutate(month = "09",
         pemlr = as.character(pemlr)) %>%
  filter(gestfips == "25")

CPS_data_Sep_2 <- CPS_data_Sep_2 %>%
  mutate(month = "09") %>%
  filter(gestfips == "25")

CPS_data_Oct <- CPS_data_Oct %>%
  mutate(month = "10") %>%
  filter(gestfips == "25")

CPS_data_Nov <- CPS_data_Nov %>%
  mutate(month = "11") %>%
  filter(gestfips == "25")

CPS_data_Dec <- CPS_data_Dec %>%
  mutate(month = "12") %>%
  filter(gestfips == "25")

CPS_data_Mar_2 <- CPS_data_Mar_2 %>%
  mutate(
    year = case_when(
      year == "1" ~ "26",
      year == "2" ~ "27",
      year == "3" ~ "28",
      year == "4" ~ "29"
    )
  )

CPS_data_Sep_2 <- CPS_data_Sep_2 %>%
  mutate(
    year = case_when(
      year == "1" ~ "24",
      year == "2" ~ "25",
      year == "3" ~ "26",
      year == "4" ~ "27",
      year == "5" ~ "28",
      year == "6" ~ "29"
    )
  )

CPS_data <- rbind(CPS_data_Jan, CPS_data_Feb, CPS_data_Mar_1, CPS_data_Mar_2, CPS_data_Apr, CPS_data_May, CPS_data_Jun,
                  CPS_data_Jul, CPS_data_Aug, CPS_data_Sep_1, CPS_data_Sep_2, CPS_data_Oct, CPS_data_Nov, CPS_data_Dec)

CPS_data <- CPS_data %>%
  mutate(
    year = case_when(
      year == "1" ~ 1994,
      year == "2" ~ 1995,
      year == "3" ~ 1996,
      year == "4" ~ 1997,
      year == "5" ~ 1998,
      year == "6" ~ 1999,
      year == "7" ~ 2000,
      year == "8" ~ 2001,
      year == "9" ~ 2002,
      year == "10" ~ 2003,
      year == "11" ~ 2004,
      year == "12" ~ 2005,
      year == "13" ~ 2006,
      year == "14" ~ 2007,
      year == "15" ~ 2008,
      year == "16" ~ 2009,
      year == "17" ~ 2010,
      year == "18" ~ 2011,
      year == "19" ~ 2012,
      year == "20" ~ 2013,
      year == "21" ~ 2014,
      year == "22" ~ 2015,
      year == "23" ~ 2016,
      year == "24" ~ 2017,
      year == "25" ~ 2018,
      year == "26" ~ 2019,
      year == "27" ~ 2020,
      year == "28" ~ 2021,
      year == "29" ~ 2022),
    educ_attn = case_when(
      peeduca <= 39 ~ 1,
      peeduca %in% 40:42 ~ 2,
      peeduca == 43 ~ 3,
      peeduca >= 44 ~ 4,
      peeduca == -1 ~ 1,
      is.na(peeduca) ~ 1),
    PAGEC =
      cut(prtage,
          breaks = c(-Inf, seq(4, 84, 5), Inf),
          labels = 1:18),
    lf = case_when(
      pemlr == 1 | pemlr == 2 | pemlr == 4 ~ 1,
      pemlr == 3 | pemlr == 5 | pemlr == 6 | pemlr == 7 ~ 2,
      pemlr == -1 ~ 2,
      is.na(pemlr) ~ 2
    )
  )

lfpr_CPS <- CPS_data %>%
  filter(lf >= 1 & prtage > 16) %>%
  mutate(date_1 = paste(year, month, "01", sep = "-")) %>%
  group_by(date_1) %>%
  count(var = lf, wt = pwcmpwgt) %>%
  mutate(
    var = case_when(
      var == 1 ~ "in LF",
      var == 2 ~ "not in LF"),
    freq = (n/sum(n))*100) %>%
  filter(var == "in LF")

##
lfpr_FRED_post09 <- lfpr_FRED[186:314,]
##

lfpr_FRED %>%
  ggplot(aes(x=date, y=value, color = "lfpr_MA")) +
  geom_line(lwd = 1.0) +
  geom_line(data = lfpr_PUMS,
            mapping = aes(x=as.Date(year, format = "%Y"), y = freq, color = "lfpr_PUMS"),
            lwd = 1.0) +
  geom_line(data = lfpr_CPS,
            mapping = aes(x = as.Date(date_1, format = "%Y-%m-%d"), y = freq, color = "lfpr_CPS"),
            lwd = 1.0) +
  geom_hline(yintercept = max(lfpr_FRED_post09$value), color = "blue", linetype = 2) +
  geom_hline(yintercept = min(lfpr_FRED_post09$value), color = "blue", linetype = 2) +
  scale_color_manual(name = "Data Series",
                     breaks = c("lfpr_FRED", "lfpr_PUMS", "lfpr_CPS"),
                     values = c("lfpr_FRED" = "black", "lfpr_PUMS" = "red", "lfpr_CPS" = "green")) +
  labs(title = "") +
  ylab("Percent (%)") +
  xlab("Year") +
  theme(legend.title = element_text(size = 16),
        legend.text = element_text(size = 12),
        panel.background = element_blank(),
        panel.border = element_rect(color = "black",  fill = NA, size = 1),
        panel.grid = element_line(color = "gray", size = 0.5),
        axis.title.x = element_text(color = "black", size = 16),
        axis.title.y = element_text(color = "black", size = 16),
        axis.text.x = element_text(color = "black", size = 14),
        axis.text.y = element_text(color = "black", size = 14))

```